---
title: "Strandings"
author: "Florian Mayer"
date: "20/02/2017"
output: html_document
---

This document illustrates an example workflow of using AnimalEncounter data 
(strandings, taggings) from the WAStD API.

```{r r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/projects/turtle-scripts/wastd")
# https://cran.r-project.org/web/packages/tidyjson/vignettes/introduction-to-tidyjson.html
# install.packages("tidyjson") # a better way to flatten JSON to CSV?
library(httr)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyjson)
library(DT)
library(leaflet)
library(ggplot2)
library(rgdal)
# install.packages("../data/packfor_0.0-8.tar.gz", repos = NULL, type = "source")
library(packfor)
if (file.exists("../config/setup.R")) source("../config/setup.R")
```

After loading some libraries and configs, the WAStD API is called, and the response
is read into a JSON dictionary.

```{r call_api}
api <- "https://strandings.dpaw.wa.gov.au/api/1/"
parms <- "animal-encounters/?encounter_type=stranding&format=json&limit=5000"
hdr <- c(Authorization = APITOKEN)
res <- httr::GET(paste0(api, parms), add_headers(.headers = hdr))
res_json <- res %>% content("text", encoding="utf-8") %>% as.tbl_json()
data_json <- attr(res_json, "JSON")[[1]]$results
```

To flatten the nested JSON structure in to a CSV-like table, we need to define
an algorithm to pull the individual fields from the depths of the nested objects.

```{r prepare_data}
flatten_observations <- function(li){
  datetime <- li$when
  lon <- li$where$longitude
  lat <- li$where$latitude
  taxon <- li$taxon
  species <- li$species
  health <- li$health
  sex <- li$sex
  encounter_type <- li$encounter_type
  cause_of_death <- li$cause_of_death
  out <- c(datetime, lon, lat, taxon, species, health, sex, encounter_type, cause_of_death)
  out
}


d <- ldply(data_json, flatten_observations) %>% tbl_df()
colnames(d) <- c("datetime", "longitude", "latitude", "taxon", "species",
                 "health", "sex", "encounter_type", "cause_of_death")
```

Finally, the data type of some columns (date, numbers) is set, and the data 
object is stored as a save point.

```{r}
utc <- "UTC"
gmt08 <- "Australia/Perth"
ord <- c("YmdHMSz", "adbYHMS")

d <- mutate(d,
            datetime=with_tz(parse_date_time(datetime, orders=c("YmdHMSz"), tz=utc), tzone=gmt08),
            longitude=as.numeric(longitude),
            latitude=as.numeric(latitude)
            ) %>%
  mutate(date=as_date(datetime))
save(d, file="data/animalencounters.Rda")
load("data/animalencounters.Rda")
glimpse(d)
```

Getting some metadata.

```{r}
# Metadata on data currency
data_retrieved_on <- with_tz(parse_date_time(res$headers$date, orders=ord, tz=utc), tzone=gmt08)
no_observations <- nrow(d)
earliest_observation <- with_tz(min(d$datetime), tzone=gmt08)
latest_observation <- with_tz(max(d$datetime), tzone=gmt08)
```

**`r no_observations` records** retrieved from the WAStD API on 
  `r data_retrieved_on` GMT+08 when compiling this workbook.

The observations span **`r earliest_observation` - `r latest_observation` GMT+08**.

## Interactive map
Hover-scroll to zoom, click features to inspect.
```{r}
data <- d %>% 
    filter(taxon=="Cecacea") %>%
    # filter(cause_of_death=="boat-strike")
    
leaflet(data) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  setView(lng=120, lat=-25, zoom=5) %>%
  addAwesomeMarkers(~longitude, ~latitude, label=~species,
                    popup=~paste(species, health, cause_of_death))
```


## Data summaries

```{r}
tally_data <- d %>%
  group_by(species, cause_of_death) %>% tally(sort=F) %>% ungroup()
DT::datatable(tally_data)
```
