---
title: "Turtle tracks"
author: "Florian Mayer, Dept Parks & Wildlife WA"
date: "12/02/2017"
output: html_document
---

```{r r setup, include=FALSE, message=FALSE, warning=FALSE}
setwd("~/projects/turtle-scripts/wastd")
knitr::opts_chunk$set(echo = TRUE)
# https://cran.r-project.org/web/packages/tidyjson/vignettes/introduction-to-tidyjson.html
# install.packages("tidyjson")
library(httr)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyjson)
library(DT)
library(leaflet)
library(ggplot2)
if (file.exists("../config/setup.R")) source("../config/setup.R")
api <- "https://strandings.dpaw.wa.gov.au/api/1/"
thv <- "turtle-nest-encounters/?where=17&limit=10000&format=json"
hdr <- c(Authorization = APITOKEN)
res <- httr::GET(paste0(api, thv), add_headers(.headers = hdr))
res_json <- res %>% content("text") %>% as.tbl_json()
data_json <- attr(res_json, "JSON")[[1]]$results

flatten_track_observations <- function(li){
  datetime <- li$when
  lon <- li$where$longitude
  lat <- li$where$latitude
  species <- li$species
  nest_age <- li$nest_age
  nest_type <- li$nest_type
  out <- c(datetime, lon, lat, species, nest_age, nest_type)
  out
}

utc <- "UTC"
gmt08 <- "Australia/Perth"
ord <- c("YmdHMSz", "adbYHMS")

d <- ldply(data_json, flatten_track_observations) %>% tbl_df()
colnames(d) <- c("datetime", "longitude", "latitude", "species", "nest_age", "nest_type")
d <- mutate(d,
            datetime=with_tz(parse_date_time(datetime, orders=c("YmdHMSz"), tz=utc), tzone=gmt08),
            longitude=as.numeric(longitude),
            latitude=as.numeric(latitude)
            ) %>%
  mutate(date=as_date(datetime))
# glimpse(d)
save(d, file="data/tracks.Rda")
load("data/tracks.Rda")

# Metadata on data currency
data_retrieved_on <- with_tz(parse_date_time(res$headers$date, orders=ord, tz=utc), tzone=gmt08)
no_observations <- nrow(d)
earliest_observation <- with_tz(min(d$datetime), tzone=gmt08)
latest_observation <- with_tz(max(d$datetime), tzone=gmt08)

# Tally all tracks
tally_data <- d %>%
  group_by(date, species, nest_age) %>% tally(sort=F) %>% ungroup()

# Tally only "fresh" tracks
tally_fresh <- d %>% filter(nest_age=="fresh") %>%
  group_by(date, species, nest_type) %>% tally(sort=F) %>% ungroup()
```
**`r no_observations` observations 
spanning from `r earliest_observation` to `r latest_observation` GMT+08 
were retrieved from WAStD via the API on `r data_retrieved_on` GMT+08.**

## Data preview
The first 200 records are shown.
```{r}
DT::datatable(head(d, n=200))
```

## Interactive map
All data are shown on an interactive map.
```{r}
leaflet(d) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  setView(lng=115.0, lat=-21.45, zoom=12) %>%
  addAwesomeMarkers(~longitude, ~latitude, label=~species,
                    popup=~paste(nest_age, species, nest_type, datetime))
```

## Fresh tracks by species
```{r}
ggplot(tally_fresh, aes(x=date, y=n, colour=nest_type)) +
  geom_point() + facet_wrap("species", ncol=1)
```

**Note** Track and nest counts within the tagging sector (Jetty to Chevron) 
from the night of 20/11/2016 to 21/11/2016 were not recorded
in the ODK Collect app the morning after, as wind and foot traffic during the night
rendered the tracks unreadable. The tracks and nests were recorded on paper data
sheets and will turn up in the tagging database WAMTRAM2.

November 20-22 were windy, and tracks on the south side of the western end were
likely blown over. This will result in false absence of tracks on 20/11/2016-22/11/2016.
Wind conditions after 22/11/2016 are unknown and may also decrease the number of 
detected tracks.

Durign the field trip in Feb 2017 (exact dates to be confirmed), a cyclone 
prevented track counts on some days and may have blown away many tracks on other 
days, resulting in days without track counts as well as days with falsely low 
track counts.


# Insight

* spatial distribution of species within location per location
* temporal distribution: beginning, peak, end of nesting season per species at location
* spatial and temporal patterns of species abundance (daily species tallies) (PCNM)


```{r, echo=FALSE}
setwd("~/projects/turtle-scripts")
```
