---
title: "Turtle nesting effort"
author: "Turtle Conservation Program, Dept Parks & Wildlife WA"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 7
    fig_height: 5
---

```{r r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/projects/turtle-scripts/wastd")
if (file.exists("../config/setup.R")) source("../config/setup.R")
if (file.exists("../config/load.R")) source("../config/load.R")
```

# Data

## Source

Track count: 

* Tracks were recorded on Android tablets using digital data capture software ODK Collect
* Forms "TrackCount 0.10" (Nov-Dec 2016), "Track or Treat 0.26" (Feb 2017) and v0.31 (Feb 2017)
* Data warehouse ODK Aggregate at [dpaw-data.appspot.com](https://dpaw-data.appspot.com/)
* Data exported (through GUI) from dpaw-data as JSON and ingested (scripted) into
  [WAStD](https://strandings.dpaw.wa.gov.au/) after each field trip

Tagging: 

* Nesting turtles recorded on paper data sheets, entered and QA'd in WAMTRAM
* Data exported from WAMTRAM master db (scripted) into CSV and saved to 
  internal data catalogue
* Data ingested (scripted) from CSV into WAStD

Using some [api tricks](https://www.rstudio.com/resources/videos/using-web-apis-from-r/),
a wrapper function `wastd_api` encapsulates the WAStD API:

## Access

* From a given serializer name (e.g. `animal-encounters` for taggings or 
  `turtle-nest-encounters` for tracks and nests) and some sensible defaults, 
  the API call is constructed and executed.
* On successful completion (which, depending on result size, can take a while),
  the response is decoded, an informative message is printed, and the nested JSON 
  structure is flattened into a data.table.

```{r api}
track_records <- wastd_api("turtle-nest-encounters")
tag_records <- wastd_api("animal-encounters")
nest_json <- wastd_api("turtle-nest-encounters", 
                             query=list(taxon="Cheloniidae", 
                                        limit=10000, 
                                        format="json", 
                                        nest_type="hatched-nest"),
                            simplify=FALSE)
# listviewer::jsonedit(nest_json$content)
# area_json <- wastd_api("areas", simplify=FALSE)
grd <- readOGR("thv_grid.geojson", "OGRGeoJSON")
save(grd, file = "data/grid.Rda")
load("data/grid.Rda")

tagging_area <- readOGR("thv_tagging.geojson", "OGRGeoJSON")
save(tagging_area, file = "data/tagging_area.Rda")
load("data/tagging_area.Rda")
```

## Transformation

Using [purrr magic](https://www.rstudio.com/resources/videos/happy-r-users-purrr/)

* data is parsed into native types (e.g. dates and numbers),
* records are filtered to a location (THV) and date range (from Dec 20), and
* the resulting R objects are saved locally to form a restore point.

```{r data_munging}
#' Return a UTC HTTP date as GMT+08
utc_as_gmt08 <- . %>%
  parse_date_time(orders = c("YmdHMSz", "adbYHMS")) %>%
  with_tz(tzone="Australia/Perth")

#' Calculate the "turtle date" from a given datetime
as_turtle_date <- . %>% utc_as_gmt08 %>% -hours(12) %>% as_date

#' Convert between url-safe dash-separated-names and Human Readable Title Case
humanize <- . %>% stringr::str_to_title() %>% stringr::str_replace("-", " ")
urlize <- . %>% stringr::str_to_lower() %>% stringr::str_replace(" ", "-")

#' Filter a data.frame to records collected at THV after 19 Dec 2016
#' TODO replace with filtering for THV 2016/2017 field trips at API
thv_filter <- . %>% filter(
    latitude < -21.43,
    latitude > -21.48,
    longitude > 114.96,
    longitude < 115.05,
    date > dmy("17/11/2016"))

#' O Lord of Darkness, accept this our functional sacrifice of three days and three nights
#' From a list of unnamed lists, or the data.frame equivalent, extract a field or -1
extract_possibly <- possibly(extract, otherwise = -1)
get_f <- function(lol, field) lol %>% unlist %>% extract_possibly(field) %>% unlist %>% as.numeric

# Flatten nested JSON, parse field formats and rename columns
tracks <- track_records$content %>% 
  mutate(
    datetime = properties.when %>% utc_as_gmt08,
    longitude = properties.longitude %>% as.numeric,
    latitude = properties.latitude %>% as.numeric,
    date = properties.when %>% as_turtle_date,
    species = properties.species,
    nest_age = properties.nest_age,
    nest_type = properties.nest_type,
    obs = properties.observation_set,
    hatching_success =  obs %>% map(get_f, "hatching_success") %>% as.numeric,
    emergence_success = obs %>% map(get_f, "emergence_success") %>% as.numeric,
    clutch_size = obs %>% map(get_f, "egg_count_calculated") %>% as.numeric
  ) %>% thv_filter 
save(tracks, file = "data/tracks.Rda")
load("data/tracks.Rda")

tags <- tag_records$content %>% 
  mutate(
    datetime = properties.when %>% utc_as_gmt08,
    longitude = properties.longitude %>% as.numeric,
    latitude = properties.latitude %>% as.numeric,
    date = properties.when %>% as_turtle_date,
    species = properties.species,
    health = properties.health,
    nesting = properties.nesting_event,
    name=properties.name
  ) %>% thv_filter
save(tags, file = "data/tags.Rda")
load("data/tags.Rda")

# Hold my beer, we'll do it in JSON
nests <- nest_json$content %>% {
  tibble::tibble(
    datetime = map_chr(., c("properties", "when")) %>% utc_as_gmt08,
    longitude = map_dbl(., c("properties", "longitude")),
    latitude = map_dbl(., c("properties", "latitude")),
    date = map_chr(., c("properties", "when")) %>% as_turtle_date,
    species = map_chr(., c("properties", "species")),
    obs = map(., c("properties", "observation_set")),
    hatching_success = obs %>% map(get_f, "hatching_success") %>% as.numeric,
    emergence_success = obs %>% map(get_f, "emergence_success") %>% as.numeric,
    clutch_size = obs %>% map(get_f, "egg_count_calculated") %>% as.numeric
    )
} %>% thv_filter
save(nests, file = "data/nests.Rda")
load("data/nests.Rda")

# listviewer::jsonedit(nests)
```

* Track observations span **`r min(tracks$datetime)` - `r max(tracks$datetime)`"**
* Tag observations span **`r min(tags$datetime)` - `r max(tags$datetime)`**

## Turtle dates
The column `date` is the reconstructed event date, as opposed to the `datetime`
of the encounter as seen by the observer.

For tagged nesting turtles:

* Assume nesting turtles are recorded between sunset and sunrise
* "turtle tagged 18 Dec 2016 22:15" = "emergence_date 18 Dec 2016"
* "turtle tagged 19 Dec 2016 02:13" = "emergence_date 18 Dec 2016"
* Count as nesting success (assuming all encounters are tagging after nesting)

For fresh tracks ("made last night"): 

* Assume tracks are recorded "the morning after" and before noon
* "fresh track recorded 19 Dec 2016 06:23" = "emergence_date 18 Dec 2016"
* Count as nesting success if nest was found.

For old tracks ("older than from last night"): 

* discard observations
* We do not check for nests on old tracks, so nesting success is "unknown" and 
  analysed as "not successful"
* Nests without clear tracks are recorded as "nest"

All dates are calculated as the calendar date (year, month, day) 12 hours before
the obesrvation's datetime. This subtracts one day from the date of all 
observations made between midnight and noon (mornings), and preserves the data of all
observations made between noon and midnight (afternoons).

All subsequent analyses bin the data by this reconstructed nesting event date.


## Interactive datatable: Tracks

```{r dt_tracks}
DT::datatable(tracks)
```

## Interactive datatable: Nesting turtles from tagging

```{r dt_tags}
DT::datatable(tags)
```

## Interactive datatable: Nests

`nests` are a subset of `tracks`. They were loaded separately to demonstrate
the capability of the WAStD API to filter data, but they could have equally been
filtered from the `tracks` dataframe.

The column `obs` contains a nested list of measurements, from which we have
extracted `hatching_success`, `emergence_success` and `clutch_size`.
There are more measurements in `obs` which we do not, but could extract and 
analyze in a similar fashion.

```{r dt_nests}
DT::datatable(nests)
```


## Interactive map

Hover-scroll to zoom, click features to inspect. 
Hold mouse outside map to scroll page.

Be patient when scrolling or zooming, this will cause a decent load on your computer.


```{r map}
#' makeAwesomeIcon factory
mkicon <- function(ico, col) makeAwesomeIcon(icon = ico, markerColor = col)

trackIcons <- awesomeIconList(
  "cheloniidae-fam" = mkicon('align-center', 'black'),
  "chelonia-mydas" = mkicon('align-center', 'green'),
  "eretmochelys-imbricata" = mkicon('align-center', 'blue'),
  "natator-depressus" = mkicon('align-center', 'red')
  )

tagIcons <- awesomeIconList(
  "cheloniidae-fam" = mkicon('tag', 'black'),
  "chelonia-mydas" = mkicon('tag', 'green'),
  "eretmochelys-imbricata" = mkicon('tag', 'blue'),
  "natator-depressus" = mkicon('tag', 'red')
  )

nestIcons <- awesomeIconList(
  "cheloniidae-fam" = mkicon('baby-formula', 'black'),
  "chelonia-mydas" = mkicon('baby-formula', 'green'),
  "eretmochelys-imbricata" = mkicon('baby-formula', 'blue'),
  "natator-depressus" = mkicon('baby-formula', 'red')
  )

leaflet(tracks) %>% 
  addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
  setView(lng=115.0, lat=-21.45, zoom=12) %>%
  addPolygons(data = grd, weight = 1, color = "orange", opacity = 0.5, 
              label = ~paste(id), group = "Grid") %>%
  addPolygons(data = tagging_area, weight = 1, color = "blue", opacity = 0.5, 
              group = "Tagging area") %>%
  addAwesomeMarkers(~longitude, ~latitude,
                    data = filter(tracks, nest_type != "hatched-nest"), 
                    # data=tracks, 
                    icon = ~trackIcons[species],
                    label = ~paste("Track", date, nest_age, species, nest_type),
                    popup = ~paste("Track", date, nest_age, species, nest_type),
                    group = "Tracks") %>%
  addAwesomeMarkers(~longitude, ~latitude, data = tags,
                    icon = ~tagIcons[species],
                    label = ~paste("Tag", date, species, name, "nesting:", nesting),
                    popup = ~paste("Tag", date, species, name, "nesting:", nesting),
                    group = "Tags") %>%
  # addHeatmap(data=tags, lng = ~longitude, lat = ~latitude, 
  #            blur = 20, max = 1, radius = 15) %>%
  addAwesomeMarkers(~longitude, ~latitude,
                    data = nests,
                    icon = ~nestIcons[species],
                    label = ~paste("Nest", date, species, 
                                   "HS", as.numeric(hatching_success), 
                                   "%, ES", as.numeric(emergence_success), "%"),
                    popup = ~paste("Nest", date, species, 
                                   "HS", as.numeric(hatching_success), 
                                   "%, ES", as.numeric(emergence_success), "%"),
                    group = "Nests") %>%
  addLayersControl(baseGroups = c("Aerial", "Place names"),
                   overlayGroups = c("Grid", "Tagging area", "Tracks", "Tags", "Nests"))
```

# Animal handling and throughput

The dataframe `tags_by_date` counts the number of applied or resighted tags by date.
Multiple encounters with the same tag on the same day are counted separately.
Therefore, a turtle tag applied by one team (recorded as applied new), and seen by 
another team (recorded as resighted) will be counted as two distinct encounters.
`tags_by_date` has one row per observed day.

The dataframe `animals_by_date` counts the number of encounters with distinct
animals per day. Other than `tags_by_date`, `animals_by_date` lists one row per
animal per date, so the total number of rows is the total number of distinct tags
handled. This number includes possible taggings of the same animal with different 
tags as duplicates.

```{r}
tags_by_date <- tags %>% group_by(date) %>% tally() %>% ungroup()
animals_by_date <- tags %>% group_by(date, name) %>% tally() %>% ungroup()
```

A total of `r nrow(tags)` turtles were processed (captured new or resighted) 
during the `r length(unique(tags_by_date$date))` days of field work on which
tagging took place.

This number excludes records in WAMTRAM with missing coordinates or time stamps. 
Re-sightings of the same turtle are counted as separate handling events.

On average, `r mean(tags_by_date$n) %>% round(digits = 2)` 
(SD `r sd(tags_by_date$n) %>% round(digits = 2)`, 
range `r min(tags_by_date$n)` to `r max(tags_by_date$n)`) turtles were processed 
per surveyed day.

`r length(unique(animals_by_date$name))` distinct tags were applied.
This number relates to new animals, but contains the tag "NA" and resightings.
This number also omits the invalid records from WAMTRAM with missing coordinates 
or dates.

**TODO**

* For each tag with nesting `present`, count number of of sightings before.
* For each tag without nesting `present`, count number of sightings within 7 days 
  from each other.
* Mean number of nests per night, sd, range
* Equivalent to ~ 600 nests over 1 km of beach in 4 months
* Nesting density as number of nests per km coastline
* Per species: CCL and CCW, mean, sd, min, max

## Animals processed per day

```{r}
datatable(tags_by_date, 
          caption = "Number of turtles processed per day")
```

## Tags processed per day

```{r}
datatable(animals_by_date, 
          caption = "Number of individual tags processed per day")
```


```{r fig_turtles_processed_per_day}
ggplot(tags_by_date, aes(x = date, y = n)) + 
  geom_bar(stat = "identity") +
  scale_x_date(breaks = scales::pretty_breaks(),
               labels = scales::date_format("%d %b %Y")) +
  labs(x="Date", y ="Number of turtles processed") +
  ggtitle("Turtles processed per day (new and recaptured)") +
  theme_light() +
  ggsave(paste0("data/turtles_processed_per_day.pdf"), width = 7, height = 5)
```


# Nesting intervals

```{r, fig.width=7, fig.height=20}
ggplot(tags, aes(x = date, y = name, color = nesting)) + 
  geom_point() + 
  geom_line(aes(date, name)) +
  scale_x_date(breaks = scales::pretty_breaks(),
               labels = scales::date_format("%d %b %Y")) +
  labs(x="Date", y ="Primary flipper tag label") +
  ggtitle("Nesting history of tagged turtles") +   
  theme_light() +
  ggsave(paste0("data/nesting_history.pdf"), width = 7, height = 20)
```

**TODO** How many days elapse to first resighting? 
This section depends on new functionality of WAStD (identifying recaptures).

# Nesting success

From `r nrow(filter(nests, hatching_success>=0))` conclusive nest excavations,
the calculated hatching and emergence success are shown summarised per day.

```{r fig_fb_hs_es, fig.width=7,fig.height=5}
nests_with_hs <- filter(nests, hatching_success >= 0, species == "natator-depressus")
ggplot(nests_with_hs, aes(group=date, x=date)) + 
  geom_boxplot(aes(y=as.numeric(hatching_success)), na.rm=TRUE, color="black") +
  geom_boxplot(aes(y=as.numeric(emergence_success)), na.rm=TRUE, color="gray") +
  scale_x_date(breaks = scales::pretty_breaks(),
               labels = scales::date_format("%d %b %Y")) +
  xlab("Date") +
  ylab("Success rate in percent") +
  ggtitle("Flatback hatching (black) and emergence (grey) success",
          subtitle = paste("From", nrow(nests_with_hs), "nests")) +
  theme_light() +
  ggsave(paste0("data/flatback_hs_es.pdf"), width = 7, height = 5)
```

## Nesting success summary table

Mean, SD and range for HS and ES and clutch size per species are calculated:


```{r}
success_summary <- nests %>%
  filter(hatching_success >= 0) %>% 
  group_by(species) %>%
  dplyr::summarize(
    "count" = n(),
    "clutch_size_mean" = mean(clutch_size) %>% round(digits = 2),
    "clutch_size_sd" = sd(clutch_size) %>% round(digits = 2),
    "clutch_size_min" = min(clutch_size),
    "clutch_size_max" = max(clutch_size),
    "hatching_success_mean" = mean(hatching_success) %>% round(digits = 2),
    "hatching_success_sd" = sd(hatching_success) %>% round(digits = 2),
    "hatching_success_min" = min(hatching_success),
    "hatching_success_max" = max(hatching_success),
    "emergence_success_mean" = mean(emergence_success) %>% round(digits = 2),
    "emergence_success_sd" = sd(emergence_success) %>% round(digits = 2),
    "emergence_success_min" = min(emergence_success),
    "emergence_success_max" = max(emergence_success)
               )
DT::datatable(success_summary, 
              caption = "Nesting success summary",
              options = list(paging = F))
```

## Nesting success map

Hatching success is calculated from excavations of hatched nests: 

* Size can correlate to clutch size, but is kept static here.
* Colour correlates to hatching success from low (red) to high (green).

```{r}
# Create a palette to bin continuous values into distinct colors
palette = rainbow(10, 
                  start = rgb2hsv(col2rgb('red'))[1], 
                  end = rgb2hsv(col2rgb('green'))[1])
pal <- colorBin(palette, nests_with_hs$hatching_success, 10, pretty = FALSE)

leaflet(
  filter(nests, hatching_success >= 0, species == "natator-depressus")) %>% 
  addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
  setView(lng=115.02, lat=-21.463, zoom=16) %>%
  addPolygons(data = tagging_area, 
              weight = 1, 
              color = "blue", 
              opacity = 0.5, 
              group = "Tagging area") %>%
  addCircleMarkers(~longitude, ~latitude,
    # radius = ~clutch_size/5,
    color = ~pal(hatching_success),
    label = ~paste("Nest", date, species, 
                   "CS", as.numeric(clutch_size), 
                   "HS", as.numeric(hatching_success), 
                   "%, ES", as.numeric(emergence_success), "%"),
    stroke = FALSE, 
    fillOpacity = 0.8,
    group = "Flatback nests") %>%
  addLegend("bottomright",
            pal = pal,
            values = ~hatching_success,
            title = "Hatching success", 
            opacity = 1) %>%
  addLayersControl(
    baseGroups = c("Aerial", "Place names"),
    overlayGroups = c("Tagging area", "Flatback nests")) %T>%
  mapshot(file = "nesting_success_map.png")
```

# Turtle tracks

On `r length(unique(tracks$date))` days, turtle tracks and nests were surveyed.

The following Table shows per species (including unknown shown as "Cheloniidae fam."):

* Total number of tracks, 
* mean number of tracks per day, 
* min and max number of tracks per day (cols)

## Track abundance summary table

```{r}
tracks_by_species <- tracks %>% 
  filter(nest_age == "fresh", 
         nest_type %in% c("track-not-assessed", 
                     "track-unsure", 
                     "false-crawl", 
                     "successful-crawl")) %>% 
  group_by(species, date) %>% 
  tally() %>% 
  group_by(species) %>% 
  dplyr::summarize("mean" = mean(n) %>% round(digits = 2),
                   "SD" = sd(n) %>% round(digits = 2),
                   "min" = min(n),
                   "max" = max(n))

datatable(tracks_by_species, 
          caption = "Fresh tracks per day (excluding nests without tracks)",
          options = list(paging = F))
```


# Nesting effort

To visualise and analyse nesting effort, data from track counts is combined with
data from tagging and missed turtles. 
Location and species are shared, while track / nest type (from track count) and
turtle nesting (from tagging) are combined into one column.

The column `date` is the "turtle date"; `diffdate` is the number of days since 1970,
which represents the date as integer number for analytic purposes.

```{r nesting_encounters}
nesting_encounters <- bind_rows(
  tracks %>% 
    filter(nest_age == "fresh") %>% 
    transmute(
      latitude = latitude, 
      longitude = longitude, 
      date = date, 
      species = species, 
      type = nest_type),
  tags %>% 
    transmute(
      latitude = latitude, 
      longitude = longitude, 
      date = date, 
      species = species, 
      type = nesting)
  ) %>%
  mutate(diffdate=difftime(date, origin))
```


Observations have to be summarised into spatial (rectangular grid cells) and 
temporal (days) units and pivoted into species profiles.

Here, a rectangular grid (roughly 100 x 100m) enclosing the study area was 
artificially created. The extent and rectangular shape of the grid has no link 
to the actual study site (the beaches of Thevenard Island), and was chosen solely
for convenience. A grid too narrow will introduce false similarity through 
shared absences, while a grid too wide will lose spatial detail.

It is to be decided whether the actual nesting beaches, gridded into equidistant
sections, would be a more appropriate spatial unit.

Turn nesting encounters into a SpatialPointsDataFrame (SPDF, R's "Shapefile"), 
then spatially join this SPDF to the grid cells the encounter lies in.
Only the grid cells over sandy beaches will contain encounters.

**Note** There may be a 
[tidier](https://github.com/tidyverse/ggplot2/wiki/plotting-polygon-shapefiles)
way of manipulating and displaying spatial data using `ggplot2`.
We are using the package `sp`.

```{r make_spdf}
wgs84 <- CRS('+proj=longlat +datum=WGS84 +no_defs')
d.xy <- select(nesting_encounters, longitude, latitude)
d.sp <- SpatialPoints(coords = d.xy, proj4string = wgs84)
d.spdf <- SpatialPointsDataFrame(d.sp, data = nesting_encounters, proj4string = wgs84)
d.spdf.grd <- d.spdf %>% spCbind(over(d.spdf, grd)) %>% spCbind(over(d.spdf, tagging_area))
qplot(longitude, latitude, data = d.spdf.grd@data, facets = ~species, color = date) + 
  coord_fixed() + theme_light() + 
  ggtitle("Total nesting over time") + 
  ggsave("data/total_nesting_over_time.pdf")
```

# Tagging area selection

Does the tagging area cover the main nesting area?

See also the interactive map above.

The proportion of tracks in the tagging area compared to the rest of the island is calculated as follows:

* Use all nesting encounters (tracks, nests and tagging) cast into 100 x 100 m grid cells (`d.spdf.grd@data`),
* discard data not within the grid, and therefore not on Thv (filter out data not matching a grid `id`),
* tally by species and location inside or outside the tagging area (inside has a `name` set, outside not),
* spread the tally into wide format to get one column for "tagging area" and one for "outside",
* calculate `all` encounters by adding encounters inside the `taggig_area` to those `outside`,
* calculate `perccent_in_tagging_area` from encounters inside the `taggig_area` to those `outside`.

```{r}
species_profiles_areas <- d.spdf.grd@data %>% 
  filter(!is.na(id)) %>%
  filter(type != "hatched-nest") %>%
  group_by(species, name) %>% 
  tally(sort = F) %>% 
  ungroup() %>% 
  spread(name, n, fill = 0) %>%
  set_colnames(c("species", "tagging_area", "outside")) %>%
  mutate(
    all = tagging_area + outside,
    percent_in_tagging_area = 100 * (
      tagging_area / (tagging_area + outside)) %>% round(digits = 3)
  )
datatable(species_profiles_areas, 
          caption = "Number of tracks (excl. nests) and taggings (excl. invalid WAMTRAM data) in tagging area vs all",
          options = list(paging = F))
```

Green turtles seem to concentrate on the NW corner and avoid the SE corner,
whereas Flatbacks seem to occur all around the island with a focus on the stretch
between west of the Jetty and the Chevron property on the East end.

* Graph: Temporal distribution of tracks (all and successful) per species (flat and green)


# Nesting effort by species (tracks and tags)

* Area: all of Thv
* Data: all nesting (tracks and tagging)
* y: Absolute numbers counted per day
* x: Date
* groups: nesting type - track types, nest types, 
  tagging (nesting "present", "absent" or "na")
* facet: species
* statistics: points, smooth (model: "loess" automatically chosen)
* warnings: data gap between field trips 

```{r plot_nesting_type_by_date, warning = F}
ggplot_nesting_type_by_date <- function(data, species_name) {
  nesting_encounters %>%
    filter(species == species_name) %>%
    group_by(date, type) %>% 
    tally() %>%
    rename(number_per_day = n) %>%
    ggplot(aes(x = date, y = number_per_day, colour = type)) + 
    geom_point() + 
    geom_smooth(method = "auto") +
    scale_x_date(breaks = scales::pretty_breaks(),
                 labels = scales::date_format("%d %b %Y")) +
    xlab("Date") +
    ylab("Number counted per day") +
    ggtitle(paste("Nesting activity of", species_name %>% humanize)) +
    theme_light() +
    ggsave(paste0("data/nesting_type_by_date_", species_name, ".pdf"))
}

ggplot_nesting_type_by_date(nesting_encounters, "natator-depressus")
ggplot_nesting_type_by_date(nesting_encounters, "chelonia-mydas")
```

## Nesting success by species (tracks and tags)

Nesting success is calculated as the ratio of tallied successful over total
nesting encounters per date and species. 
This combines nesting over the whole area for each surveyed day.

As successful nesting events are counted (this is negotiable):

* fresh tracks with nests (`successful-crawl`),
* hatched nests,
* tagged turtles with successful nesting (`present`).


* Area: all of Thv
* Data: all nesting (tracks and tagging)
* y: Absolute numbers counted per day
* x: Date
* groups: successful nesting, all nesting
* facet: species
* statistics: histogram
* warnings: data gap between field trips 

```{r plot_nesting_abundance, fig.height=5, fig.width=7, warning = F}
all_nesting_by_date <- nesting_encounters %>%
  filter(type != "hatched-nest" & type != "nest") %>%
  group_by(date, species) %>% tally() %>% ungroup() %>% 
  rename(all = n)

successful_nesting_by_date <- nesting_encounters %>% 
  filter(type == 'successful-crawl' | type == "present") %>%
  group_by(date, species) %>% tally() %>% ungroup() %>%
  transmute(date = date, 
            species = species, 
            successful = n)

# Join nesting emergences with nesting success or 0
nesting_success <- all_nesting_by_date %>%
  left_join(successful_nesting_by_date, by = c('date','species')) %>%
  mutate(successful = ifelse(is.na(successful), 0, successful)) %>%
  mutate(nesting_success_rate = 100 * successful/all)

ggplot_nesting_all_by_date <- function(data, species_name) {
  data %>%
    filter(species == species_name) %>%
    ggplot(aes(x=date)) + 
    geom_bar(aes(y=all), stat="identity", color="black", fill="black") +
    geom_bar(aes(y=successful), stat="identity", color="green", fill="green") +
    scale_x_date(breaks = scales::pretty_breaks(),
                 labels = scales::date_format("%d %b %Y")) +
    labs(x="Date", y="Number of all and successful nesting events") +
    ggtitle(paste("Nesting activity of", species_name %>% humanize),
            subtitle="Number of successful (green) and all (black) nesting events") +
  labs(x="Date", y="Number of all and successful nesting events") +
    theme_light() +
    ggsave(paste0("data/nesting_abundance_", species_name, ".pdf"), 
           width = 7, height = 5)
}

ggplot_nesting_all_by_date(nesting_success, "natator-depressus")
ggplot_nesting_all_by_date(nesting_success, "chelonia-mydas")

nesting_success_summary <- nesting_success %>%
  group_by(species) %>%
  dplyr::summarize(
    "observed_days" = n(),
    "nesting_events_mean" = mean(all) %>% round(digits = 2),
    "nesting_events_sd" = sd(all) %>% round(digits = 2),
    "nesting_events_min" = min(all),
    "nesting_events_max" = max(all),
    "nesting_success_rate_mean" = mean(nesting_success_rate) %>% round(digits = 2),
    "nesting_success_rate_sd" = sd(nesting_success_rate) %>% round(digits = 2),
    "nesting_success_rate_min" = min(nesting_success_rate) %>% round(digits = 2),
    "nesting_success_rate_max" = max(nesting_success_rate) %>% round(digits = 2)
    )

DT::datatable(nesting_success_summary, 
              caption = "Nesting event totals and success rate from tracks (excl. old nests) and tags")

```


**Note** Track and nest counts within the tagging sector (Jetty to Chevron) 
from the night of 20/11/2016 to 21/11/2016 were not recorded
in the ODK Collect app the morning after, as wind and foot traffic during the night
rendered the tracks unreadable. The tracks and nests were recorded on paper data
sheets and will turn up in the tagging database WAMTRAM2.

November 20-22 were windy, and tracks on the south side of the western end were
likely blown over. This will result in false absence of tracks on 20/11/2016-22/11/2016.
Wind conditions after 22/11/2016 are unknown and may also decrease the number of 
detected tracks.

No track count survey on morning of 2016-11-25. Track count on 2016-11-26 
(turtle date 2016-11-25) artificially inflated track numbers by counting all 
tracks as "fresh", including tracks from non-surveyed day before. 
It would be accetable to split tracks from turtle date 2016-11-25 across turtle 
dates 2016-11-25 and 2016-11-24.

**TODO** 

* Identify periods of artificially low track count from field diary and highlight
  in figures.
* Enter surveys into WAStD, annotate with observations on weather, methodology,
  any known biases on recorded data.
* Insert here comments from expedition and surveys.


## Nesting success by species (tracks only)

Nesting success is calculated as the ratio of tallied successful over total
nesting encounters per date and species. 
This combines nesting over the whole area for each surveyed day.

As successful nesting events are counted (this is negotiable):

* fresh tracks with nests (`successful-crawl`).

* Area: all of Thv
* Data: **tracks only**, no nests, no tagging
* y: Absolute numbers counted per day
* x: Date
* groups: successful nesting (successful crawls), all nesting 
  (false crawls, successful crawls, unsure crawls, tracks not assessed for success)
* facet: species
* statistics: points, smooth (model: "loess" automatically chosen)
* warnings: data gap between field trips 

```{r plot_track_effort, fig.height=5, fig.width=7, warning = F}
fresh_tracks <- tracks %>%
  filter(properties.nest_type %in% c("successful-crawl", 
                                     "false-crawl", 
                                     "track-unsure", 
                                     "track-not-assessed"))

all_tracks_by_date <- fresh_tracks %>%
  group_by(date, species) %>% tally() %>% ungroup() %>% rename(all = n)

successful_tracks_by_date <- tracks %>% 
  filter(properties.nest_type == 'successful-crawl') %>%
  group_by(date, species) %>% tally() %>% ungroup() %>% rename(successful = n)

# Join nesting emergences with nesting success or 0
track_success <- all_tracks_by_date %>%
  left_join(successful_tracks_by_date, by = c('date','species')) %>%
  mutate(successful = ifelse(is.na(successful), 0, successful),
         track_success = 100 * successful/all)

ggplot_track_success_by_date <- function(data, species_name) {
  data %>%
    filter(species == species_name) %>%
    ggplot(aes(x = date)) + 
    geom_bar(aes(y = all), stat = "identity", color = "black", fill = "black") +
    geom_bar(aes(y = successful), stat = "identity", color = "green", fill = "green") +
    scale_x_date(breaks = scales::pretty_breaks(),
                 labels = scales::date_format("%d %b %Y")) +
    labs(x = "Date", y = "Number of all and successful tracks") +
    ggtitle(paste("Nesting effort of", species_name %>% humanize),
            subtitle = "Number of all (black) and successful (green) tracks") +
  labs(x = "Date", y = "Number of all and successful tracks") +
    theme_light() +
    ggsave(paste0("data/track_effort_", species_name, ".pdf"), 
           width = 7, height = 5)
}
ggplot_track_success_by_date(track_success, "natator-depressus")
ggplot_track_success_by_date(track_success, "chelonia-mydas")

ggplot_track_successrate_by_date <- function(data, species_name) {
  data %>%
    filter(species == species_name) %>%
    ggplot(aes(x = date)) + 
    geom_bar(aes(y = track_success), stat = "identity") +
    scale_x_date(breaks = scales::pretty_breaks(),
                 labels = scales::date_format("%d %b %Y")) +
    labs(x = "Date", y = "Fraction of tracks with nest") +
    ggtitle(paste("Nesting success of", species_name %>% humanize),
            subtitle = "Fraction of successful over total nesting crawls") +
    theme_light() +
    ggsave(paste0("data/track_success_", species_name, ".pdf"), 
           width = 7, height = 5)
}

ggplot_track_successrate_by_date(track_success, "natator-depressus")
ggplot_track_successrate_by_date(track_success, "chelonia-mydas")

# TODO mean, sd, min, max of track_success$track_success for each species

track_success_summary <- track_success %>%
  group_by(species) %>%
  dplyr::summarise(
    mean_nesting_success = mean(track_success) %>% round(digits = 2),
    sd_nesting_success = sd(track_success) %>% round(digits = 2)
  )
datatable(track_success_summary, 
          caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)")

```

The total number of fresh tracks was `r nrow(fresh_tracks)`.

# Spatial and temporal structures in nesting
This section conducts a PCNM analysis to detect spatial and temporal patterns
in the observed abundance.

Create species profiles for species track abundance per grid cell id and encounter date.
```{r dt_species_profiles}
species_profiles <- d.spdf.grd@data %>% 
  filter(!is.na(id)) %>%
  group_by(diffdate, date, id, species) %>% 
  tally(sort=F) %>% 
  ungroup() %>%
  spread(species, n, fill=0) %>% left_join(grd, by="id", copy=T) %>%
  set_colnames(make.names(colnames(.)))

DT::datatable(species_profiles, 
              caption = "Species profiles (100 x 100m grid)",
              options = list(paging = T, pageLength = 10))
```

Separate the species profile data into 

* initial responding data Y.in (nesting event tallies per species), and
* point coordinates of each grid cell X.coords.

These data frames have one row per surveyed combination of grid cell and date.

Listing only grid cells with encounters, Xs contain the coordinates, Xs.id contain
coordinates plus grid cell ID.
Both Xs and Xs.id have one row per grid cell with encounters.

Lastly, Xt lists all unique surveyed turtle dates as days since 1970, a numerical
value as required by spatial analysis tools which cannot parse comlex data types 
such as  dates.

```{r split_species_profiles}
Y.in <- species_profiles %>% 
  dplyr::select(-id, -diffdate, -date, -xmin, -xmax, -ymin, -ymax)

X.coords <- species_profiles %>% dplyr::select(xmin, ymin)

Xs.id <- grd@data %>% 
  filter(id %in% unique(species_profiles$id)) %>% 
  dplyr::select(id, xmin, ymin)
Xs <- Xs.id %>% dplyr::select(-id)

Xt <- unique(species_profiles$diffdate) %>% tbl_df()
colnames(Xt) <- "diffdate"
```



```{r, echo=F}
# totals_all <- site_summary %>% 
#   group_by(species) %>% 
#   tally %>% 
#   dplyr::rename(all = nn)
# 
# totals_tagging_area <- site_summary %>% 
#   filter(!is.na(name)) %>% 
#   group_by(species) %>% 
#   tally %>% 
#   dplyr::rename(in_tagging_area = nn)
# 
# encounters_all_vs_tagging <- totals_all %>% 
#   join(totals_tagging_area, by="species") %>%
#   mutate(percent_in_tagging_area = round(100 * in_tagging_area / all, digits=3))
# 
# DT::datatable(encounters_all_vs_tagging)
```


```{r plot_pairs_Y}
pairs(Y.in, main="Correlations between species")
```

The nesting emergence abundance between Flatbacks and Greens seem to correlate inversely.
Observations of Hawksbills and turtles with unknown species ID are sparse.

* Histograms of Y and X will show the distribution of variable values.
* Expected: normal or uniform distribution.
* Right-skewed distributions can be log-transformed.
* Left-skewed distributions can be sqrt-transformed.
* Sparse observations should be Hellinger-transformed to reduce false similarity 
  between sites through shared absence (which might be undersampling rather than 
  true absence).

```{r plot_Y_abundance}
qplot(Y.in$"chelonia.mydas", 
      geom="histogram", 
      main="Abundance of Chelonia mydas", 
      xlab="Absolute abundance per grid cell",
      ylab="Count")
qplot(Y.in$"natator.depressus", 
      geom="histogram", 
      main="Abundance of Natator depressus", 
      xlab="Absolute abundance per grid cell",
      ylab="Count")
```

The two main species look a bit left-skewed and show many absences.

The Hellinger transformation will help reduce false similarity through shared absences.
This assumes sparse observations, where shared absence is due to undersampling
or low population density. The high number of nesting events may violate this assumption,
so that Hellinger transformation may not be appropriate.

Alternatively, relative abundance is calculated through division by row (site) sum.

```{r plot_Yhel}
Y.hel <- Y.in %>% decostand("hellinger")
pairs(Y.hel)
qplot(Y.hel$"chelonia.mydas", 
      geom="histogram", 
      main="Abundance of Chelonia mydas", 
      xlab="Hellinger-transformed abundance per grid cell",
      ylab="Count")
qplot(Y.hel$"natator.depressus", 
      geom="histogram", 
      main="Abundance of Natator depressus", 
      xlab="Hellinger-transformed abundance per grid cell",
      ylab="Count")

Y.tot <- Y.in %>% decostand("total", MARGIN=1) # relative abundance as % of site total count
pairs(Y.tot)
qplot(Y.tot$"chelonia.mydas", 
      geom="histogram", 
      main="Abundance of Chelonia mydas", 
      xlab="Relative abundance (% of site total) per grid cell",
      ylab="Count")
qplot(Y.tot$"natator.depressus", 
      geom="histogram", 
      main="Abundance of Chelonia mydas", 
      xlab="Relative abundance (% of site total) per grid cell",
      ylab="Count")
```

Remove linear spatial trends, lest they eat up explanatory power.

```{r plot_detrended_Y}
Y <- residuals(rda(Y.tot, X.coords))
pairs(Y)
```

Following [this tutorial](https://sites.ualberta.ca/~ahamann/teaching/renr690/Lab9.pdf).

```{r pcnm}
plot(Xs)
Xs.site.pcnm <- PCNM(dist(Xs))
summary(Xs.site.pcnm)
plot(Xs.site.pcnm$spanning, main="Spanning tree of sites")
Xs.site.pcnm.pos <- as.data.frame(Xs.site.pcnm$vectors)[,which(Xs.site.pcnm$Moran_I$Positive == T)]
sites_pcnm <- cbind(Xs.id, scores(Xs.site.pcnm.pos))
surveys_pcnm <- species_profiles %>% dplyr::left_join(sites_pcnm, by="id", copy=T)
Xs.scores <- dplyr::select(surveys_pcnm, starts_with("V"))

Xt.pcnm <- PCNM(dist(Xt))
summary(Xt.pcnm)
plot(Xt.pcnm$spanning, main="Spanning tree of observation dates")
Xt.pcnm.pos <- as.data.frame(Xt.pcnm$vectors)[,which(Xt.pcnm$Moran_I$Positive == T)]
dates_pcnm <- cbind(Xt, scores(Xt.pcnm.pos))
surveys_pcnm_t <- species_profiles %>% dplyr::left_join(dates_pcnm, by="diffdate", copy=T)
Xt.scores <- dplyr::select(surveys_pcnm_t, starts_with("V"))

Xs.sel <- forward.sel(Y, Xs.scores)
Xt.sel <- forward.sel(Y, Xt.scores)
XS <- Xs.scores %>% dplyr::select(Xs.sel$order)
XT <- Xt.scores %>% dplyr::select(Xt.sel$order)

Xs.sel
Xt.sel
Y.vp <- varpart(Y, ~as.matrix(XS), ~as.matrix(XT))
Y.vp
plot(Y.vp)
```

## Variance explained by spatial and temporal patterns

Outcome: 

* Spatial patterns explain a total of `r round(Y.vp$part$fract$Adj.R.squared[1] * 100, 2)`% of the variance.
* Temporal patterns explain a total of `r round(Y.vp$part$fract$Adj.R.squared[2] * 100, 2)`% of the variance.
* Another `r round(Y.vp$part$indfract$Adj.R.squared[2] * 100, 2)`% of the variance 
  is co-explained by both spatial and temporal patterns.
* The remaining variation that is unexplained through spatial or temporal patterns is
  `r round(Y.vp$part$indfract$Adj.R.squared[4] * 100, 2)`%.
  
Caveat:

This analysis can detect patterns from the size of the widest gap (small spatially,
but wide temporal gap between field trips) and up to the extent of the study (whole island,
and November to Feb).

The homogeneous density of tracks around the whole island agrees with the finding
of very little influence of spatially structured processes.

The large temporal gap between field trips however will play a significant role
in hiding the strong seasonality in nesting abundance from the analysis.

## Spatial patterns

```{r plot_xs_pcnm}
plot.xs <- function(nr, coords, scores, fs){
  pdf(paste0("data/spatial_pattern_", nr, ".pdf"), width = 7, height = 5)
  ordisurf(coords, scores[,nr], bubble = 3,
                  main=paste("Spatial pattern", fs[nr,2]), 
                  sub=paste("explains", 
                            round(as.numeric(fs[nr,3]), digits = 3) * 100, 
                            "% variance"))
  dev.off()
  ordisurf(coords, scores[,nr], bubble = 3,
                  main=paste("Spatial pattern", fs[nr,2]), 
                  sub=paste("explains", 
                            round(as.numeric(fs[nr,3]), digits = 3) * 100, 
                            "% variance"))
}

lapply(seq_len(nrow(Xs.sel)), 
       plot.xs, 
       dplyr::select(sites_pcnm, xmin, ymin), 
       dplyr::select(sites_pcnm, starts_with("V")), 
       Xs.sel)
```

## Temporal patterns

```{r lot_xt_pcnm}
plot_pcnm <- function(no){
  temp_pcnm_2 <- data.frame(date = species_profiles$date, pcnm = XT[,no][[1]])
  ggplot(temp_pcnm_2, aes(date, pcnm)) + 
    geom_line() + 
    theme_minimal() + 
    labs(x = "Date", y = paste("PCNM variable", no)) +
    ggtitle(paste("Temporal pattern", no),
            subtitle = paste("Explains", 
                             round((Xt.sel$R2[no] * 100), digits = 2), 
                             "% variance")) +
    ggsave(paste0("data/temporal_pattern", no, ".pdf"), width = 7, height = 5)
}
lapply(Xt.sel$order, plot_pcnm)
```

A spot of modelling! Pure nonsense at this stage.

```{r ordination}
XSm <- as.matrix(XS)
XTm <- as.matrix(XT)

Y.rda <- rda(Y ~ XSm + XTm)
fig <- ordiplot(Y.rda, 
                main = "Constrained ordination (RDA) all species", 
                type = "none")
points(fig, "sites", pch = 21, col = "red", bg = "yellow")
text(fig, "species", col = "blue", cex = 0.9)
identify(fig, "spec")
```

```{r, reset_workdir, echo=FALSE}
setwd("~/projects/turtle-scripts")
```
