---
title: "Turtle Sightings Roebuck Bay 2018"
author: "Florian Mayer, WA Turtles, DBCA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(here::here("wa-turtle-programs", "tracks_setup.R"))
source(here::here("wa-turtle-programs", "tracks_helpers.R"))
# source(here::here("wa-turtle-programs", "load_data.R"))
```

# Data

```{r data}
s18a <- readr::read_csv(
  here::here("data", "2018_sightings.csv")) %>% 
  dplyr::mutate(
    date = lubridate::parse_date_time(
      date, orders="%d/%m/%Y", tz="Australia/Perth"),
    latitude = latitude %>% as.double(),
    longitude = longitude %>% as.double()
  )

s18b_records <- wastdr::wastd_GET(
  "animal-encounters", 
  query = list(taxon = "Cheloniidae", format = "json", site=57)
)
s18b <- wastdr::parse_animal_encounters(s18b_records)
```

This workbook analyses turtle sightings from two fieldtrips.

In the first fieldtrip, `r s18a %>% nrow` turtles were encountered on 
`r s18a$date %>% unique %>% length` days between `r s18a$date %>% min` and 
`r s18a$date %>% max`.
Data from this fieldtrip was captured on paper datasheets and transcribed into a spreadsheet.
This workbook reads directly from a copy of this spreadsheet.

In the second fieldtrip, `r s18b %>% nrow` turtles were encountered on 
`r s18b$turtle_date %>% unique %>% length` days between `r s18b$datetime %>% min` and 
`r s18b$datetime %>% max`.
Data from this fieldtrip was captured digitally using 
[OpenDataKit](https://opendatakit.org/)("ODK"")
software and our ODK form "Turtle Sighting 0.1".
The data from all ODK forms we use is automatically ingested into the 
[WA Sea Turtle Database ("WAStD")](https://tsc.dbca.wa.gov.au/) and is 
accessible to authorised users through its [API](https://tsc.dbca.wa.gov.au/api/1/).
This workbook reads and parses turtle sightings directly from the 
[WAStD API](https://tsc.dbca.wa.gov.au/api/1/) using the R package 
[wastdr](https://dbca-wa.github.io/wastdr/).

# Map

The spatial distribution of the turtle sightings shown here is driven by

* the survey effort (kms driven, hrs time spent looking),
* observer detection bias and efficiency,
* the turtle abundance, and
* turtle behaviour (fraction of time spent on surface where visible to observers).

```{r map}
l <- leaflet::leaflet(width = 800, height = 600) %>%
    leaflet::addProviderTiles("Esri.WorldImagery", 
                              group = "Aerial") %>%
    leaflet::addProviderTiles("OpenStreetMap.Mapnik", 
                              group = "Place names") %>%
    leaflet::clearBounds() %>%
    leaflet::addAwesomeMarkers(
      data = s18a,
      lng = ~longitude, lat = ~latitude,
      icon = leaflet::makeAwesomeIcon(icon = "tag", 
                                      markerColor = "red"),
      label = ~glue::glue(
        '{format(date, "%d/%m/%Y")} {lifestage} ',
        '{capture} Depth {depth %||% NA} m'),
      group = "2018-06"
    ) %>%
      leaflet::addAwesomeMarkers(
      data = s18b,
      lng = ~longitude, lat = ~latitude,
      icon = leaflet::makeAwesomeIcon(icon = "tag", 
                                      markerColor = "blue"),
      label = ~glue::glue(
        '{format(datetime, "%d/%m/%Y")} {maturity} ',
        '{species} {activity}'),
      group = "2018-08"
    ) %>%
    leaflet::addLayersControl(
      baseGroups = c("Aerial", "Place names"),
      overlayGroups = c("2018-06", "2018-08"),
      options = leaflet::layersControlOptions(
        collapsed = FALSE)
)

mapview::mapshot(
  l, 
  selfcontained = TRUE,
  url = here::here("wastd", "turtle-sightings-rbb2018.html"),
  file = here::here("wastd", "turtle-sightings-rbb2018.png"))
l
```

