---
title: "DKAN and ckanR"
author: "ckanR contributors"
date: "`r Sys.Date()`"
output: html_document
---

# Setup
Install latest ckanR, dplyr and DT.

```{r, eval=FALSE, message=FALSE}
install.packages("devtools")
devtools::install_github("rstudio/DT")
devtools::install_github("hadley/dplyr")
devtools::install_github("ropensci/ckanr")
```

# Config
Save yourself specifying the URL with each ckanR call by setting a default url.
```{r}
library(ckanr)
ckanr::ckanr_setup(url="http://datos.imss.gob.mx/")
```

# Reading data
Reading the first dataset and the first resource of IMSS.gov.mx:

```{r}
packages <- ckanr::package_list()
p1 <- ckanr::package_show(packages[[1]])
r1 <- ckanr::resource_show(p1$resources[[1]]$id)
# did that work? print the resulting objects
p1
r1
```

Using the data retrieved by ckanR, you can now 
[view the first dataset](`r p1$url`) or [download its first resource](`r r1$url`).

Reading a specific dataset:
```{r}
p <- ckanr::package_show("asegurados-2013")
r <- ckanr::resource_show("57f8e388-7cad-4105-a5b0-6eae49f1e96e")
```


You can now read data from the package `r p$name`, e.g. resource `r r$name` into R.
Note: As the [CSV file](`r r1$url`) is `r r$size`, this may take a while, 
so we only read in the first 100 lines.

```{r}
data <- read.csv(r$url, header=T, as.is=T, nrows = 100, sep="|")
dplyr::glimpse(data)
DT::datatable(data)
```
