---
title: "Cable Beach Broome Turtle Nesting Census"
author: "Marine Turtles WA, Sarah Mullineux, Florian Mayer"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
source("tracks_setup.R")
source("tracks_helpers.R")
source("load_data.R")
```

# Data

```{r filter_data}
tracks_bme <- tracks_all %>% filter_bme() %>% exclude_training_species() %>% add_nest_labels()
surveys_bme <- surveys %>% filter_bme()
surveys_bme_prod <- surveys_bme %>% exclude_training_surveys()
dist_bme <- disturbance %>% filter_bme()
nests_bme <- nests_all %>% filter_bme()
prefix_bme <- "BME"
placename_bme <- "Cable Beach Broome"
```

# QA
Most QA is done across all WA records and programs in the separate QA workbook.

This section identifies data requiring editing.

Training dates:

* turtle date 2017-11-13, 2017-10-28, 2017-10-17

## Provided username mismatches WAStD users
```{r name_mismatch_bme}
surveys_bme %>% filter_surveys_requiring_qa() %>% dt()
```

# Season summary

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.
Overall, the Volunteers have saved 
`r round((nrow(tracks_bme) + nrow(surveys_bme)*2 + nrow(dist_bme)) * 2 / 60)` 
hours of mind-numbing data entry and proofreading work for 
`r nrow(tracks_bme)` Tracks, 
`r nrow(dist_bme)` Disturbances, and 
`r nrow(surveys_bme)*2` Site Visit Starts/Ends. 

## Maps
```{r map_bme}
tracks_bme %>% filter_2017() %>% map_tracks()
tracks_bme %>% filter_2018() %>% map_tracks()
```

### Animated map
A [Google API key](https://developers.google.com/maps/documentation/geocoding/start?csw=1) must
be present as R environment variable `GOOGLE_MAPS_APIKEY` for the next step.

```{r map_bme_animated, eval=F, warning=FALSE, message=FALSE}
tracks_bme_2017 <- tracks_bme %>% filter_2017()
# gganimate_tracks(tracks_bme_2017, placename_bme, prefix_bme)
```

### Tagged nests
The map is saved as a PNG file `r glue::glue(prefix_bme, "_tagged_nests.png")`.

Coming 2018.

```{r nests_bme}
nests_bme %>%
  filter_2018() %>%
  map_nests() %>%
  mapview::mapshot(file = glue::glue("{prefix_bme}_tagged_nests_2018.png"))
```

## Nesting abundance
### All beaches
```{r tracks_all_sectors_bme}
tracks_bme %>% nesting_type_by_season_species() %>% kable()
```

## Nesting by week
### All beaches
The next table show data from all sites.

```{r tracks_all_beaches_by_week_bme}
tracks_bme %>% nesting_type_by_season_week_species() %>% kable()
```

## Nesting by day
This section shows data from all sites.

```{r nesting_success_bme, warning = F}
tracks_ts(tracks_bme, surveys_bme_prod, placename_bme, prefix_bme)
nests <- tracks_bme %>% track_success()
nests %>% ggplot_track_success_by_date("natator-depressus", placename_bme, prefix_bme)
nests %>% ggplot_track_successrate_by_date("natator-depressus", placename_bme, prefix_bme)
nests %>% ggplot_track_success_by_date("chelonia-mydas", placename_bme, prefix_bme)
nests %>% ggplot_track_successrate_by_date("chelonia-mydas", placename_bme, prefix_bme)
nests %>% track_success_by_species() %>% DT::datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```

## Hatching and emergence success
```{r hatching_emergence_success_bme}
tracks_bme %>%
  hatching_emergence_success() %>%
  DT::datatable(., caption = "Hatching and emergence success summary", options = list(paging = F))
```

## Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Fox Sake" and appear here as "dist_(place)".

### Disturbance and predation of nests
Turtle nests with recorded disturbance or predation.

```{r dist_nests_bme}
disturbed_nests_bme_2017 <- tracks_bme %>% filter_2017() %>% filter(disturbance == "present")
disturbed_nests_bme_2017 %>% map_tracks()

disturbed_nests_bme_2018 <- tracks_bme %>% filter_2018() %>% filter(disturbance == "present")
disturbed_nests_bme_2018 %>% map_tracks()
```

There were **`r count(disturbed_nests_bme_2017)` reports** of disturbed nests in 2017 and 
**`r count(disturbed_nests_bme_2018)` reports** in 2018. 
Caveat: QA to exclude possible training records.

Coming soon: details of disturbance to nests.

### General disturbance
There were **`r count(dist_bme)` reports** of general disturbance or predator presence.

```{r map_dist_bme}
dist_bme %>% filter_2017() %>% group_by(season, disturbance_cause) %>% tally() %>% arrange(-n) %>% kable()
dist_bme %>% filter_2017() %>% map_dist()

dist_bme %>% filter_2018() %>% group_by(season, disturbance_cause) %>% tally() %>% arrange(-n) %>% kable()
dist_bme %>% filter_2018() %>% map_dist()
```

# Surveys
Curators are to QA 
[all `r placename_bme` surveys on WAStD](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__in=22,23,24)
following the 
[QA instructions for surveys](https://wastd.readthedocs.io/data_curators.html#data-flow-of-surveys).

### Per season effort
Highest level of aggregation: entire season.
Assumption: all training surveys have been marked as such in WAStD.

#### Excluding training

```{r survey_season_stats_all}
surveys_bme %>% survey_season_stats %>% dt
```

### Per day effort
Lowest level of aggregation: daily.

```{r survey_count}
surveys_bme %>% survey_count_heatmap(placename_bme, prefix_bme)
surveys_bme %>% list_survey_count(placename_bme)
surveys_bme %>% survey_hours_heatmap(placename_bme, prefix_bme)
surveys_bme %>% list_survey_effort(placename_bme)
```

### Individual surveys
No aggregation: individual surveys.

```{r surveys_list}
surveys_bme %>% survey_show_detail %>% dt()
```


## Survey effort per primary reporter

### Caveat
This list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text.
To correctly report on team member effort, curators have to read the list of 
team members in the Site Visit "comments" field and attach the correct users 
to the Survey team list.
Once the surveys are comprehensively attributed to all volunteers present, this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with
team rosters.

```{r survey_effort_by_person}
personal_effort_bme <- surveys_bme %>% survey_hours_per_person()
personal_effort_bme %>% kable()
```

## Volunteer effort 
### Season 2017-18
Number of surveys per sector times length of site.

In sector 1, `r survey_ground_covered(surveys, 22, 4, 2017)` km were walked 
in `r survey_count(surveys, 22, 2017)` surveys.

In sector 2, `r survey_ground_covered(surveys, 23, 4, 2017)` km were walked 
in `r survey_count(surveys, 23, 2017)` surveys.

In sector 3, `r survey_ground_covered(surveys, 24, 7, 2017)` km were walked 
in `r survey_count(surveys, 24, 2017)` surveys.

### Season 2018-19
Number of surveys per sector times length of site.

In sector 1, `r survey_ground_covered(surveys, 22, 4, 2018)` km were walked 
in `r survey_count(surveys, 22, 2018)` surveys.

In sector 2, `r survey_ground_covered(surveys, 23, 4, 2018)` km were walked 
in `r survey_count(surveys, 23, 2018)` surveys.

In sector 3, `r survey_ground_covered(surveys, 24, 7, 2018)` km were walked 
in `r survey_count(surveys, 24, 2018)` surveys.



# Data upload

```{r data_upload}
tracks_bme %>% dplyr::select(-obs, -photos) %>% 
  readr::write_csv(glue::glue("{prefix_bme}_tracks.csv"))
surveys_bme %>% readr::write_csv(glue::glue("{prefix_bme}_surveys.csv"))
dist_bme %>% dplyr::select(-photos) %>% 
  readr::write_csv(glue::glue("{prefix_bme}_disturbance.csv"))

# Package all output files into one ZIP archive
products_fn_bme <- glue::glue("{prefix_bme}_products.zip")
if (file.exists(products_fn_bme)) file.remove(products_fn_bme)
zip(zipfile = products_fn_bme, files = list.files(pattern = prefix_bme))

# Create a resource for the ZIP archives
d <- ckanr::package_show("turtle-tracks")
# r <- resource_create(package_id=d$id, name="Cable Beach Broome Turtle Nesting Census", upload="broome.html")
# r <- resource_create(package_id=d$id, name="Cable Beach Broome outputs", upload=products_fn_bme)

ckanr::resource_update("9572e76b-55e9-4afb-b224-7cce7beb4959", "broome.html")
ckanr::resource_update("fa5d9fff-90d4-4ac2-a578-18cc2964f6f6", products_fn_bme)
```

