---
title: "Conzinc Bay Turtle Nesting Census"
author: "Murujuga Aboriginal Corporation, DBCA Karratha and North West Shelf Flatbacks WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
source("tracks_setup.R")
source("tracks_helpers.R")
source("load_data.R")
```

# Caveat
Data from 2017/18 at Rosemary Islands was recovered from unfinalised forms on devices.

# Data
```{r load_data}
tracks_con <- tracks_all %>% filter_con() %>% exclude_training_species() %>% add_nest_labels()
surveys_con <- surveys %>% filter_con()
surveys_con_prod <- surveys_con %>% exclude_training_surveys()
dist_con <- disturbance %>% filter_con()
nests_con <- nests_all %>% filter_con()
prefix_con <- "CON"
placename_con <- "Conzinc Bay"
```

# QA
Most QA is done across all WA records and programs in the separate QA workbook.

This section identifies data requiring editing.

## Provided username mismatches WAStD users
```{r name_mismatch_con}
surveys_con %>% filter_surveys_requiring_qa() %>% dt()
```

# Season summary

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.

Time savings Rosemary: `r (nrow(tracks_con) + nrow(surveys_con)*2 + nrow(dist_con)) * 2 / 60` 
hours for `r nrow(tracks_con)` Track or Treats and  `r nrow(surveys_con)*2` Site Visit Starts/Ends. 

## Maps
### 2018-19
```{r map_con_2018}
tracks_con %>% filter_2018() %>% map_tracks()
```


### Animated map
A [Google API key](https://developers.google.com/maps/documentation/geocoding/start?csw=1) must
be present as R environment variable `GOOGLE_MAPS_APIKEY` for the next step.

```{r map_con_animated, warning=FALSE, eval=F}
tracks_con_2018 <- tracks_con %>% filter_2018()
gganimate_tracks(tracks_con_2018, placename_con, prefix_con)
```

### Tagged nests
The map is saved as a PNG file `r glue::glue(prefix_con, "_tagged_nests.png")`.

Coming 2020.

```{r nests_con, eval=F}
nests_con %>%
  filter_2018() %>%
  map_nests() %>%
  mapview::mapshot(file = glue::glue("{prefix}_tagged_nests_2018.png"))
```

## Nesting abundance
### All beaches
```{r tracks_all_sectors_con}
tracks_con %>% nesting_type_by_season_species() %>% kable()
```

## Nesting by week
### All beaches
The next table show data from all sites.

```{r tracks_all_beaches_by_week_con}
tracks_con %>% nesting_type_by_season_week_species() %>% kable()
```

## Nesting by day
This section shows data from all sites.

```{r nesting_success_con, warning = F}
tracks_con %>% tracks_ts(surveys_con_prod, placename_con, prefix_con)
nests <- tracks_con %>% track_success()
# nests %>% ggplot_track_success_by_date("natator-depressus", placename_con, prefix_con)
# nests %>% ggplot_track_successrate_by_date("natator-depressus", placename_con, prefix_con)
# nests %>% ggplot_track_success_by_date("chelonia-mydas", placename_con, prefix_con)
# nests %>% ggplot_track_successrate_by_date("chelonia-mydas", placename_con, prefix_con)
nests %>% ggplot_track_success_by_date("eretmochelys-imbricata", placename_con, prefix_con)
nests %>% ggplot_track_successrate_by_date("eretmochelys-imbricata", placename_con, prefix_con)
nests %>% track_success_by_species() %>% DT::datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```

## Hatching and emergence success
```{r hatching_emergence_success_con}
tracks_con %>%
  hatching_emergence_success() %>%
  DT::datatable(., caption = "Hatching and emergence success summary", options = list(paging = F))
```

## Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Fox Sake" and appear here as "dist_(place)".

### Disturbance and predation of nests 2018-19
Turtle nests with recorded disturbance or predation.

```{r dist_nests_con_2018}
disturbed_nests_con_2018 <- tracks_con %>% filter_2018() %>% filter(disturbance == "present")
disturbed_nests_con_2018 %>% map_tracks()
```

There were **`r count(disturbed_nests_con_2018)` reports** of disturbed nests. 

Coming soon: details of disturbance to nests.

### General disturbance 2018-19
```{r map_dist_con_2018}
dist_con_2018 <- dist_con %>% filter_2018() 
dist_con_2018 %>% group_by(disturbance_cause) %>% tally() %>% arrange(-n) %>% kable()
dist_con %>% filter_2018() %>% map_dist()
```
There were **`r count(dist_con_2018)` reports** of general disturbance or predator presence.

# Surveys
Curators are to QA 
[all `r placename_con` surveys on WAStD](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=115)
following the 
[QA instructions for surveys](https://wastd.readthedocs.io/data_curators.html#data-flow-of-surveys).

### Per season effort
Highest level of aggregation: entire season.
Assumption: all training surveys have been marked as such in WAStD.

#### Excluding training

```{r survey_season_stats_all}
surveys_con %>% survey_season_stats %>% dt0
```

### Per day effort
Lowest level of aggregation: daily.

```{r survey_count}
surveys_con %>% survey_count_heatmap(placename_con, prefix_con)
surveys_con %>% list_survey_count(placename_con)
surveys_con %>% survey_hours_heatmap(placename_con, prefix_con)
surveys_con %>% list_survey_effort(placename_con)
```

### Individual surveys
No aggregation: individual surveys.

```{r surveys_list}
surveys_con %>% survey_show_detail %>% dt0
```

## Survey effort per primary reporter

### Caveat
This list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text.
To correctly report on team member effort, curators have to read the list of 
team members in the Site Visit "comments" field and attach the correct users 
to the Survey team list.
Once the surveys are comprehensively attributed to all volunteers present, this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with
team rosters.

```{r survey_effort_by_person}
personal_effort_con <- surveys_con %>% survey_hours_per_person()
personal_effort_con %>% kable()
```

## Survey effort in kilometers walked
Number of surveys per sector times length of site.

### Season 2018-19
At Conzinc Bay, assuming the trip was 4 km,
`r survey_ground_covered(surveys, 115, 4, 2018)` km were walked 
in `r survey_count(surveys, 115, 2018)` surveys.

# Data upload
## Raw data
* Raw data are exported into CSV spreadsheets.
* Key figures are exported into .png raster files.
* This report is rendered into a single HTML page.

```{r data_export_csv}
tracks_con %>%
  select(-obs, -photos) %>%
  readr::write_csv(glue::glue("{prefix_con}_tracks.csv"))
surveys_con %>% readr::write_csv(glue::glue("{prefix_con}_surveys.csv"))
dist_con %>%
  select(-photos) %>%
  readr::write_csv(glue::glue("{prefix_con}_disturbance.csv"))
```

## Data catalogue
Data are uploaded to the [turtle nest census dataset](https://data.dpaw.wa.gov.au/dataset/turtle-tracks) 
on the departmental data catalogue, accessible from the DBCA intranet only.

```{r data_upload}
# Package all output files into one ZIP archive
products_fn_con <- glue::glue("{prefix_con}_products.zip")
if (file.exists(products_fn_con)) file.remove(products_fn_con)
zip(zipfile = products_fn_con, files = list.files(pattern = prefix_con))

# Create a resource for the ZIP archives
# d <- ckanr::package_show("turtle-tracks")
# r <- resource_create(package_id=d$id, name="Conzinc Bay Turtle Nesting Census", upload="conzinc.html")
# r <- resource_create(package_id=d$id, name="Conzinc Bay outputs", upload=products_fn_con)

# Update resources on data catalogue
ckanr::resource_update("55595ff0-092a-454c-8208-decbb7c789ba", "conzinc.html")
ckanr::resource_update("e494eaa3-a8e6-4417-8d2b-5665e5696c2b", products_fn_con)
```

## Google Drive
A copy of this report and all generated outputs is uploaded to Google Drive
and shared via 
[this link](https://drive.google.com/drive/folders/12qGo66U7iTU5vGIkVjKOIOoKHkurn4MG?usp=sharing) 
with DBCA external collaborators.

```{r google_drive_auth, eval=F}
googledrive::drive_auth(reset = TRUE, use_oob = TRUE)
```

```{r data_upload_google}
googledrive::drive_ls(prefix_con) %>% googledrive::drive_rm(.)
googledrive::drive_upload("conzinc.html", path = "CON/conzinc.html")
list.files(pattern = prefix_con) %>% purrr::map(googledrive::drive_upload, path = as_dribble(prefix_con))
```
