---
title: "Delambre Island Turtle Nesting Census"
author: "Marine Turtles WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: show
    theme: lumen
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
source("tracks_setup.R")
source("tracks_helpers.R")
source("load_data.R")
```

# Caveat
Training and test data from Delambre Island are not yet separated out.
Data from Rosemary Islands was recovered from unfinalised forms on devices.

# Data
```{r load_data}
tracks_del <- tracks_all %>% filter_del() %>% exclude_training_species() %>% add_nest_labels()
tags_del <- animals %>% filter_del() %>% exclude_training_species()
tags_del_2017 <- tags_del %>% filter_2017()
surveys_del <- surveys %>% filter_del()
surveys_del_prod <- surveys_del %>% exclude_training_surveys()
dist_del <- disturbance %>% filter_del()
nests_del <- nests_all %>% filter_del()
prefix_del <- "DEL"
placename_del <- "Delambre Is"
```


# QA
Most QA is done across all WA records and programs in the separate QA workbook.

This section identifies data requiring editing.

## Provided username mismatches WAStD users
```{r name_mismatch_del}
surveys_del %>% filter_surveys_requiring_qa() %>% dt()
```

# Background
Source: Milani Chaloupka, presentation to DBCA / NWS committee 8/5/2018.

Demographic focus is essential for diagnosing population status and trend.
Survival rates, conditional breeding rates, hatchling production rates, recruitment rates and population abundance.

Three major flatback pops in Pilbara: Delambre 3.5k nesting/y, 
Mundabullangana 2.5k nesting/y, Barrow Is 2k/y.
Biggest pop in GBR is 700 nests/y.

Milani presented: 

* Bayesian gompertz state-space model trend
* Random effects forest plot with rookery specific posterior distribution

# Season summary

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.

Time savings Delambre: `r (nrow(tracks_del) + nrow(surveys_del)*2 + nrow(dist_del)) * 2 / 60` 
hours for `r nrow(tracks_del)` Track or Treats and  `r nrow(surveys_del)*2` Site Visit Starts/Ends. 

# Delambre Nesting
## Tagged animals
```{r}
tags_by_date <- tags_del %>% filter_2017 %>% group_by(turtle_date) %>% tally() %>% ungroup()
animals_by_date <- tags_del %>% filter_2017 %>% group_by(turtle_date, name) %>% tally() %>% ungroup()
```

A total of `r tags_del %>% filter_2017 %>% nrow()` turtles were processed (captured new or resighted)
during the `r length(unique(tags_by_date$turtle_date))` days of field work on which tagging took place.

This number excludes records in WAMTRAM with missing coordinates or time stamps. 
Re-sightings of the same turtle are counted as separate handling events.

On average, `r mean(tags_by_date$n) %>% round(digits = 2)` 
(SD `r sd(tags_by_date$n) %>% round(digits = 2)`, 
range `r min(tags_by_date$n)` to `r max(tags_by_date$n)`) turtles were processed 
per surveyed day.

`r length(unique(animals_by_date$name))` distinct tags were applied.
This number relates to new animals, but contains the tag "NA" and resightings.
This number also omits the invalid records from WAMTRAM with missing coordinates 
or dates.

**TODO**

* For each tag with nesting `present`, count number of of sightings before.
* For each tag without nesting `present`, count number of sightings within 7 days 
  from each other.
* Mean number of nests per night, sd, range
* Equivalent to ~ 600 nests over 1 km of beach in 4 months
* Nesting density as number of nests per km coastline
* Per species: CCL and CCW, mean, sd, min, max

### Animals processed per day

```{r}
datatable(tags_by_date,
  caption = "Number of turtles processed per day"
)
```

### Tags processed per day

```{r}
datatable(animals_by_date,
  caption = "Number of individual tags processed per day"
)
```

```{r fig_turtles_processed_per_day}
ggplot2::ggplot(tags_by_date, ggplot2::aes(x = turtle_date, y = n)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::scale_x_date(
    breaks = scales::pretty_breaks(),
    labels = scales::date_format("%d %b %Y")
  ) +
  ggplot2::labs(x = "Turtle date", y = "Number of turtles processed") +
  ggplot2::ggtitle("Turtles processed per day (new and recaptured)") +
  ggplot2::theme_classic() +
  # ggplot2::ggsave(ggplot2::ggsave(glue::glue("{prefix_del}_turtles_processed_per_day.png"), 
  # device = png(), width = 10, height = 15)) + 
  NULL
```

## Nesting intervals

```{r, fig.width=10, fig.height=30}
tags_del_2017_tally <- tags_del_2017 %>% group_by(name) %>% tally() 

tags_del_2017_resighted_records <- tags_del_2017 %>% 
  filter(name %in% filter(tags_del_2017_tally, n>1)$name) %>% 
  tidyr::replace_na(list(name = "Tag unknown"))

tags_del_2017_single_records <- tags_del_2017 %>% 
  filter(name %in% filter(tags_del_2017_tally, n==1)$name) %>% 
  mutate(name="Seen once")

ggplot(tags_del_2017_resighted_records, 
       aes(x = turtle_date, y = name, color = nesting_event)) +
  geom_point(data = tags_del_2017_single_records) +
  geom_point(data = tags_del_2017_resighted_records) +
  geom_line(data = tags_del_2017_resighted_records, aes(turtle_date, name)) +
  scale_x_date(
    breaks = scales::pretty_breaks(),
    labels = scales::date_format("%d %b %Y")
  ) +
  ggtitle("Nesting history of tagged turtles") +
  theme_classic() + 
  ggsave(glue::glue("{prefix_del}_nesting_intervals.png"), 
         width = 10, height = 30, limitsize = FALSE)
```


The chart shows encounters with the given tags over the season. Every dot is a tag (tag ID is y-axis) being applied or re-sighted at a given turtle date (x-axis).

Re-sighted tags are connected with a line.

The categories "nesting event" or not indicate whether the turtle was confirmed nesting or maybe just identified while false crawling (and tag still recorded).

There is a heap of single encounters which I didn't want to filter out as not to imply that all tags were re-sighted.

The natural progression of new tag IDs versus date reflects the tagging teams going through their sorted tag batches over the season.

The chart is immensely tall (60 inches printed) for this huge number of tags. Looking for clever ways to shrink the chart height.


**TODO** How many days elapse to first resighting? 
This section depends on new functionality of WAStD (identifying recaptures).
## Maps
```{r map_del}
tracks_del %>% filter_2017() %>% map_tracks()
```

### Animated map
A [Google API key](https://developers.google.com/maps/documentation/geocoding/start?csw=1) must
be present as R environment variable `GOOGLE_MAPS_APIKEY` for the next step.

```{r map_del_animated, warning=FALSE, message=FALSE}
# tracks_del %>% filter_2017() %>% gganimate_tracks(placename_del, prefix_del)
```

### Tagged nests
The map is saved as a PNG file `r glue::glue(prefix_del, "_tagged_nests.png")`.

```{r nests_del}
nests_del %>%
  filter_2017() %>%
  map_nests() %T>%
  mapview::mapshot(file = glue::glue("{prefix_del}_tagged_nests_2017.png"))
```

## Nesting abundance
### All beaches
```{r tracks_all_sectors_del}
tracks_del %>% nesting_type_by_season_species() %>% kable()
```

## Nesting by week
### All beaches
The next table show data from all sites.

```{r tracks_all_beaches_by_week_del}
tracks_del %>% nesting_type_by_season_week_species() %>% kable()
```

## Nesting by day
This section shows data from all sites.

```{r nesting_success_del, warning = F}
tracks_ts(tracks_del, surveys_del_prod, placename_del, prefix_del)
nests <- tracks_del %>% track_success()
nests %>% ggplot_track_success_by_date("natator-depressus", placename_del, prefix_del)
nests %>% ggplot_track_successrate_by_date("natator-depressus", placename_del, prefix_del)
nests %>% ggplot_track_success_by_date("chelonia-mydas", placename_del, prefix_del)
nests %>% ggplot_track_successrate_by_date("chelonia-mydas", placename_del, prefix_del)
nests %>% track_success_by_species() %>% DT::datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```

## Hatching and emergence success
```{r hatching_emergence_success_del}
tracks_del %>%
  hatching_emergence_success() %>%
  DT::datatable(., caption = "Hatching and emergence success summary", options = list(paging = F))
```

## Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Fox Sake" and appear here as "dist_(place)".

### Disturbance and predation of nests
Turtle nests with recorded disturbance or predation.

```{r dist_nests_del}
disturbed_nests_del <- tracks_del %>% filter_2017() %>% filter(disturbance == "present")
disturbed_nests_del %>% map_tracks()
```

There were **`r count(disturbed_nests_del)` reports** of disturbed nests. 
Caveat: QA to exclude possible training records.

Coming soon: details of disturbance to nests.

### General disturbance
There were **`r count(dist_del)` reports** of general disturbance or predator presence.

```{r map_dist_del}
dist_del %>% filter_2017() %>% group_by(disturbance_cause) %>% tally() %>% arrange(-n) %>% kable()
dist_del %>% map_dist()
```


# Surveys
Curators are to QA 
[all `r placename_del` surveys on WAStD](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=39)
following the 
[QA instructions for surveys](https://wastd.readthedocs.io/data_curators.html#data-flow-of-surveys).

### Per season effort
Highest level of aggregation: entire season.
Assumption: all training surveys have been marked as such in WAStD.

#### Excluding training

```{r survey_season_stats_all}
surveys_del %>% survey_season_stats %>% dt
```

### Per day effort
Lowest level of aggregation: daily.

```{r survey_count}
surveys_del %>% survey_count_heatmap(placename_del, prefix_del)
surveys_del %>% list_survey_count(placename_del)
surveys_del %>% survey_hours_heatmap(placename_del, prefix_del)
surveys_del %>% list_survey_effort(placename_del)
```

### Individual surveys
No aggregation: individual surveys.

```{r surveys_list}
surveys_del %>% survey_show_detail %>% dt()
```

## Survey effort per primary reporter
Q Which volunteer did the most hours overall on survey?

### Caveat
This list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text.
To correctly report on team member effort, curators have to read the list of 
team members in the Site Visit "comments" field and attach the correct users 
to the Survey team list.
Once the surveys are comprehensively attributed to all volunteers present, this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with
team rosters.

```{r survey_effort_by_person}
personal_effort_del <- surveys_del %>% survey_hours_per_person()
personal_effort_del %>% kable()
```

## Survey effort
### Season 2017-18
Number of surveys per sector times length of site.

At Delambre, assuming the trip was 4 km,
`r survey_ground_covered(surveys, 39, 4, 2017)` km were walked 
in `r survey_count(surveys, 39, 2017)` surveys.

### Season 2018-19
Number of surveys per sector times length of site.

At Delambre, assuming the trip was 4 km,
`r survey_ground_covered(surveys, 39, 4, 2018)` km were walked 
in `r survey_count(surveys, 39, 2018)` surveys.

# Data upload
## Raw data
* Raw data are exported into CSV spreadsheets.
* Key figures are exported into .png raster files.
* This report is rendered into a single HTML page.

```{r data_export_csv}
tracks_del %>%
  select(-obs, -photos) %>%
  readr::write_csv(glue::glue("{prefix_del}_tracks.csv"))
surveys_del %>% readr::write_csv(glue::glue("{prefix_del}_surveys.csv"))
dist_del %>%
  select(-photos) %>%
  readr::write_csv(glue::glue("{prefix_del}_disturbance.csv"))
```

## Data catalogue
Data are uploaded to the [turtle nest census dataset](https://data.dpaw.wa.gov.au/dataset/turtle-tracks) 
on the departmental data catalogue, accessible from the DBCA intranet only.

```{r data_upload}
# Package all output files into one ZIP archive
products_fn_del <- glue::glue("{prefix_del}_products.zip")
if (file.exists(products_fn_del)) file.remove(products_fn_del)
zip(zipfile = products_fn_del, files = list.files(pattern = prefix_del))

# Create a resource for the ZIP archives
# d <- ckanr::package_show("turtle-tracks")
# r <- resource_create(package_id=d$id, name="Delambre Island Turtle Nesting Census", upload="delambre.html")
# r <- resource_create(package_id=d$id, name="Delambre outputs", upload=products_fn_del)

# Update resources on data catalogue
ckanr::resource_update("69d5585a-3307-41ee-980e-f829f7eef004", "delambre.html")
ckanr::resource_update("fd1c15a4-4766-4039-b3a7-5a3dd7b403a2", products_fn_del)
```
