---
title: "Eighty Mile Beach Season Review 2017-2018"
author: "Marine Turtles WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
source("tracks_setup.R")
source("tracks_helpers.R")
source("load_data.R")
```

# Caveat
The data shown are pending QA and reconciliation with paper-based records. 
Until this warning disappears, this dataset may contain duplicates from training runs, effort attributed to the wrong operators, and possibly other issues and biases.

# Data
Eighy Mile Beach is surveyed at sites Anna Plains and Caravan Park.
```{r load_data}
tracks_emb <- tracks_all %>% filter_emb() %>% filter_realspecies() %>% add_nest_labels()
tracks_emb_ap <- tracks_emb %>% filter_emb_ap()
tracks_emb_cv <- tracks_emb %>% filter_emb_cvp()
dist_emb <- disturbance %>% filter_emb()
surveys_emb <- surveys %>% filter_emb()
surveys_emb_prod <- surveys_emb %>% filter_realsurveys()
nests_emb <- nests_all %>% filter_emb()

prefix <- "EMB"
placename <- "Eighty Mile Beach"
```


# QA
This section identifies tracks and disturbances outside known sites, missing surveys, and surveys requiring editing.
We show data from all regions as not to miss any stray records - ignore data not relevant to this analysis.

## Tracks outside known sites
Review whether there are stray tracks close to this program's sites.

* WAStD sites missing (new sites) or boundaries set too narrow. 
  Extend sites, re-save stray tracks to auto-repair their site affiliation.
* Training records close to site (but not meant to be within site).
  Change confirmed training records outside training areas to "Hatchback turtles" 
  and mark as "curated" to prevent from being overwritten by repeated data loading.
* Note: re-run uncached data import after changing tracks.

```{r tracks_missing_sites}
tracks_all %>% 
  filter_nosite() %>%
  filter_realspecies() %>%
  add_nest_labels() %>% 
  map_tracks()
```

## Disturbance outside known sites
Review disturbances close to this program's sites. It is safe to ignore other disturbances.

Disturbances can either be legitimate, although opportunistic, records, or training
records.

* Training records: set to "Other, see comments", mention "training" in comments,
  ideally instruct data collectors to photograph their writing board with "training"
  written on it to make absolutely clear that this record is a training record.

```{r dist_missing_sites}
disturbance %>% filter_nosite() %>% map_dist()
```

## Surveys in WAStD
Curators are to QA 
[all `r placename` surveys on WAStD](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__in=36,37) following the 
[QA instructions for surveys](https://wastd.readthedocs.io/data_curators.html#data-flow-of-surveys).

Split up per site:

* [Caravan Park](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=36)
* [Anna Plains](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=37)

## Surveys missing end
Where a "Site Visit End" form was forgotten, `end_source_id` is empty, and the duration is set to 6h. 
These surveys should be closed a few minutes after the last recorded track or disturbance.

Surveys with missing end contribute to an over-estimation of survey effort.

This task will be automated soon.
  
```{r surveys_missing_end}
surveys_emb %>% filter_surveys_missing_end %>% dt()
```

## Provided username mismatches WAStD users
TODO: define QA rules for mismatches of usernames as (possibly mis)typed by data collectors vs 
[WAStD users](https://tsc.dbca.wa.gov.au/admin/users/user/).

```{r name_mismatch}
surveys_emb %>% filter_surveys_requiring_qa %>% dt()
```

# Season summary

## General notes
Nests without a clearly visible emergence point were not recorded.
This led to under-reporting of hatched nests this season.

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.
This season, the Volunteers have saved 
`r (nrow(tracks_emb) + nrow(surveys_emb)*2 + nrow(dist_emb)) * 2 /60` hours of 
mind-numbing data entry and proofreading work for `r nrow(tracks_emb)` Track or Treats and 
`r nrow(surveys_emb)*2` Site Visit Starts/Ends.

## Maps
```{r map_overall}
tracks_emb %>% map_tracks()
# tracks_ap %>% map_tracks()
# tracks_cv %>% map_tracks()
```


### Animated map
A [Google API key](https://developers.google.com/maps/documentation/geocoding/start?csw=1) must
be present as R environment variable `GOOGLE_MAPS_APIKEY` for the next step.

```{r map_animated}
ggmap::register_google(key = Sys.getenv("GOOGLE_MAPS_APIKEY"))

# basemap
bbx <- ggmap::make_bbox(longitude, latitude, tracks_emb, f = 0.05)
ctr <- c(mean(bbx["left"], bbx["right"]), mean(bbx["top"], bbx["bottom"]))
# basemap <- ggmap::get_map(ctr, zoom=17) %>% ggmap::ggmap()
basemap <- ggmap::get_googlemap(ctr, zoom = 15, scale = 2, maptype = "hybrid") %>%
  ggmap::ggmap()

# tracks
tracks_map <- suppressWarnings(
  basemap +
    geom_point(aes(longitude, latitude, colour = nest_type),
      data = tracks_pth, alpha = 0.8
    )
) +
  ggplot2::ggtitle("Turtle Nesting", subtitle = "Turtle date: {frame_time}") +
  gganimate::transition_time(date) +
  gganimate::ease_aes("linear")

# render and save
gganimate::animate(tracks_map, fps = 2) %T>%
  gganimate::anim_save(glue::glue("{prefix}_nesting.gif"), .)
```


### Tagged nests
2017-18: no tagged nests.

The map (once enabled) is saved as a PNG file `r glue::glue(prefix, "_tagged_nests.png")`.

```{r nests_all, eval=F}
nests_emb %>% map_nests() %>% mapview::mapshot(file=glue::glue("{prefix}_tagged_nests.png"))
```


## Nesting abundance
```{r tracks_all_sectors}
tracks_emb %>% species_by_type() %>% kable(.)
```

```{r tracks_ap}
tracks_emb_ap %>% species_by_type() %>% kable()
```

```{r tracks_cv}
tracks_emb_cv %>% species_by_type() %>% kable()
```

## Nesting success 
Split into CVP and APL?

```{r nesting_success, warning = F}
tracks_emb %>% tracks_ts()

tracks_emb %>% tracks_ts(placename, prefix)
tracks_emb_ap %>% tracks_ts("Cemetery Beach", prefix)
tracks_emb_cp %>% tracks_ts("Pretty Pool Beach", prefix)

nests <- tracks_emb %>% track_success()

nests %>% ggplot_track_success_by_date("natator-depressus", placename, prefix)
nests %>% ggplot_track_successrate_by_date("natator-depressus", placename, prefix)
nests %>% track_success_by_species() %>% DT::datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```


```{r}
nests <- tracks_cv %>% track_success()
label <- "Caravan Park"

nests %>% ggplot_track_success_by_date("natator-depressus", label)
# nests %>% ggplot_track_success_by_date("chelonia-mydas", label)
nests %>% ggplot_track_successrate_by_date("natator-depressus", label)
# nests %>% ggplot_track_successrate_by_date("chelonia-mydas", label)
nests %>% track_success_by_species() %>% datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```

## Hatching and emergence success
```{r hatching_emergence_success}
tracks_ap %>%
  hatching_emergence_success() %>%
  DT::datatable(.,
    caption = "Hatching and emergence success summary Anna Plains",
    options = list(paging = F)
  )

tracks_cv %>%
  hatching_emergence_success() %>%
  DT::datatable(.,
    caption = "Hatching and emergence success summary Caravan Park",
    options = list(paging = F)
  )

```


## Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Fox Sake" and appear here as "dist_(place)".

### Disturbance and predation of nests
Turtle nests with recorded disturbance or predation.

```{r}
disturbed_nests_emb <- tracks_emb %>% filter(disturbance == "present")
disturbed_nests_emb %>% add_nest_labels() %>% map_tracks()
```

There were **`r count(disturbed_nests_emb)` reports** of disturbed nests. Caveat: QA to exclude possible training records.

Coming soon: details of disturbance to nests.

### General disturbance

There were **`r count(dist_emb)` reports** of general disturbance or predator presence:

```{r}
dist_emb %>% group_by(disturbance_cause) %>% tally() %>% arrange(-n) %>% kable()
dist_emb %>% map_dist
```


## Survey count

### Caveat
* Data includes training surveys.

This contributes to over-estimation of effort, but shows actual boots on the 
ground - training is also an effort!

```{r survey_count}
surveys_emb %>% survey_count_heatmap(placename, prefix)
surveys_emb %>% list_survey_count(placename)
surveys_emb %>% survey_hours_heatmap(placename, prefix)
surveys_emb %>% list_survey_effort(placename)
```

## Survey effort in hours
Q How many hours did volunteers survey, how many kilometers were walked?

### Caveat
* Surveys with missing end points are auto-closed after 5 hours. 
* Data includes training surveys.

Both factors contribute to over-estimation of effort.

```{r survey_effort}
surveys_emb %>% plot_survey_effort("Eighty Mile Beach")
surveys_emb %>% list_survey_effort("Eighty Mile Beach")
```

## Survey effort per primary reporter
Q Which volunteer did the most hours overall on survey?

### Caveat
This list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text.
To correctly report on team member effort, curators have to read the list of 
team members in the Site Visit "comments" field and attach the correct users 
to the Survey team list.
Once the surveys are comprehensively attributed to all volunteers present, this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with
team rosters.


Note: Florian Mayer was logged into the device during training, 
so hours logged against Florian should correctly be attributed to the actual
data collectors.

```{r survey_effort_by_person}
personal_effort <- surveys_emb %>% survey_hours_per_person()
personal_effort %>% kable()
```

## Survey effort in kilometers walked
Number of surveys per sector times length of site.

**Note** adjust length estimates.

At Anna Plains, assuming the round trip was 6 km,
`r survey_ground_covered(surveys, 37, 6)` km were covered 
in `r survey_count(surveys, 37)` surveys.

At Caravan Park Beach, assuming the round trip was 6 km,
`r survey_ground_covered(surveys, 36, 6)` km were covered 
in `r survey_count(surveys, 36)` surveys.

# Data upload

```{r data_upload}
# d <- ckanr::package_show("turtle-tracks")
ckanr::resource_update("56f1071d-b8ed-4ac7-bb37-a5151978a9f1", "emb_2017-18.html")
```

```{r data_export}
tracks_emb %>% 
  select(-obs, -photos) %>% 
  readr::write_csv(here::here("data", "tracks_emb.csv"))
surveys_emb %>% 
  readr::write_csv(here::here("data", "surveys_emb.csv"))
dist_emb %>% select(-photos) %>% 
  readr::write_csv(here::here("data", "dist_emb.csv"))
```

