---
title: "Ningaloo Turtle Nesting Census"
author: "Ningaloo Turtle Program and North West Shelf Flatbacks WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
source("load_data.R")
source("tracks_setup.R")
source("tracks_helpers.R")
```

# Data

```{r filter_data}
tracks_nin <- tracks_all %>% 
  filter_nin() %>% 
  exclude_training_species() %>% 
  add_nest_labels()
tracks_nin_mi <- tracks_nin %>% filter_nin_mi
tracks_nin_nwc <- tracks_nin %>% filter_nin_nwc
tracks_nin_cr <- tracks_nin %>% filter_nin_cr
tracks_nin_bn <- tracks_nin %>% filter_nin_bn
tracks_nin_cb <- tracks_nin %>% filter_nin_cb
tracks_nin_wr <- tracks_nin %>% filter_nin_wr
tracks_nin_gn <- tracks_nin %>% filter_nin_gn

surveys_nin <- surveys %>% filter_nin()
surveys_nin_prod <- surveys_nin %>% exclude_training_surveys()

dist_nin <- disturbance %>% filter_nin() %>% filter_disturbance()
dist_nin_mi <- dist_nin %>% filter_nin_mi
dist_nin_nwc <- dist_nin %>% filter_nin_nwc
dist_nin_cr <- dist_nin %>% filter_nin_cr
dist_nin_bn <- dist_nin %>% filter_nin_bn
dist_nin_cb <- dist_nin %>% filter_nin_cb
dist_nin_wr <- dist_nin %>% filter_nin_wr
dist_nin_gn <- dist_nin %>% filter_nin_gn

pred_nin <- disturbance %>% filter_nin() %>% filter_predation()
pred_nin_mi <- pred_nin %>% filter_nin_mi
pred_nin_nwc <- pred_nin %>% filter_nin_nwc
pred_nin_cr <- pred_nin %>% filter_nin_cr
pred_nin_bn <- pred_nin %>% filter_nin_bn
pred_nin_cb <- pred_nin %>% filter_nin_cb
pred_nin_wr <- pred_nin %>% filter_nin_wr
pred_nin_gn <- pred_nin %>% filter_nin_gn

nests_nin <- nests_all %>% filter_nin()
prefix_nin <- "NIN"
placename_nin <- "Ningaloo"
```


# Products for Ningaloo

* Predation separate from disturbance
* Add filters for sections as well
* For each of NW Cape, Bungelup, Janes Bay, Gnaraloo: species, number false 
  crawls, number of nests (table > CSV, figure > PNG)
* Tour guides want to know highest likelihood of finding hatchlings.
* Nat cons teams want to know where to put predator traps.

We need the following issues to be addressed to be able to use the data:

* Survey dates. Clarify confusion between turtle date and calendar date.
  * Calendar date of encounter_datetime in local timezone. 
  
* Predation versus predator tracks: Clarify predated nests vs disturbed nests vs 
  signs of predator presence and signs of disturbances
* Survey effort: 
  * list dates surveyed (calendar date) per site. 
  * list dates surveyed (calendar date) per subsection.
  * tally dates surveyed per site.
  * tally dates surveyed per subsection.

* Reliable records of which subsections were monitored and when. At present the 
  data provided indicate that some subsections were monitored twice on some days 
  and adjacent subsections weren’t monitored at all. This is an artifact of 
  sites being started (site visit start geolocation) incorrectly against training.
* Gnaraloo Bay. We need to be able to distinguish among Divisions and Gnaraloo Bay from Janes Bay from the North West Cape, etc. Don’t appear to be able to distinguish Gnaraloo Bay at present.
* Training records. Training records appear to still be in the data we’ve been provided. These need to be removed.
* Dead turtles and rescued turtles. We need to be able to distinguish between dead turtles and rescued turtles and at which subsection they were found.
* We have records of some turtle rescues that weren’t recorded with the tablets. Need to have the facility to add these to the database.

All good points, some are gaps in training, some in QA, others in reports. Nothing impossible.

1 Likely a misunderstanding of turtle vs calendar date. will make sure Peter uses the CSV with both dates in them.

2 Those are split up correctly in reports: predated/disturbed nests vs stand-alone signs of pred/dist. will make two separate CSVs.

3 Depends on QA of surveys. Some work needed there from my end. Corrine could help once I have specified work to be done.

4 See report, should be there, but will make sure it's in there

5 Need to review. Shouldn't be hard.

6 Will review and remove.

7 Working on that. NTP have been reporting rescues as strandings but will find a way to distinguish. That's a new CSV I'll add for them.

8 Dani Rob is trained how to enter those.



# QA
Most QA is done across all WA records and programs in the separate QA workbook.

This section identifies data requiring editing.

Training dates:

* TODO insert

## Provided username mismatches WAStD users
```{r name_mismatch_nin}
surveys_nin %>% filter_surveys_requiring_qa() %>% dt()
```

# Season summary

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.
Overall, the Volunteers have saved 
`r round((nrow(tracks_nin) + nrow(surveys_nin)*2 + nrow(dist_nin)) * 2 / 60)` 
hours of mind-numbing data entry and proofreading work for 
`r nrow(tracks_nin)` Tracks, 
`r nrow(dist_nin)` Disturbances, and 
`r nrow(surveys_nin)*2` Site Visit Starts/Ends. 

## Maps
### 2018-19
All  turtle tracks or nests recorded in season 2018-19. 
Click markers to view popup with details
and link to original record on WAStD.
Toggle visibility of individual species in the legend.

```{r map_nin_2018}
tracks_nin %>% filter_2018() %>% map_tracks()
```

### Animated map
A [Google API key](https://developers.google.com/maps/documentation/geocoding/start?csw=1) must
be present as R environment variable `GOOGLE_MAPS_APIKEY` for the next step.

```{r map_nin_animated, eval=F, warning=FALSE, message=FALSE}
tracks_nin_2018 <- tracks_nin %>% filter_2018()
# gganimate_tracks(tracks_nin_2018, placename_nin, prefix_nin)
```

### Tagged nests
The map is saved as a PNG file `r glue::glue(prefix_nin, "_tagged_nests.png")`.

Currently unused.

```{r nests_nin, eval=F}
nests_nin %>%
  filter_2018() %>%
  map_nests() %>%
  mapview::mapshot(file = glue::glue("{prefix_nin}_tagged_nests_2018.png"))
```

## Nesting abundance
### All beaches
```{r tracks_all_sectors_nin}
tracks_nin %>% nesting_type_by_season_species() %>% knitr::kable()
```

## Nesting by week
### All beaches
The next table show data from all sites.

```{r tracks_all_beaches_by_week_nin}
tracks_nin %>% nesting_type_by_season_week_species() %>% kable()
```

## Nesting by day
This section shows data from all sites.

```{r nesting_success_nin, warning = F}
tracks_ts(tracks_nin, surveys_nin_prod, placename_nin, prefix_nin)
nests <- tracks_nin %>% track_success()
nests %>% ggplot_track_success_by_date("caretta-caretta", placename_nin, prefix_nin)
nests %>% ggplot_track_successrate_by_date("caretta-caretta", placename_nin, prefix_nin)
nests %>% ggplot_track_success_by_date("chelonia-mydas", placename_nin, prefix_nin)
nests %>% ggplot_track_successrate_by_date("chelonia-mydas", placename_nin, prefix_nin)
nests %>% track_success_by_species() %>% DT::datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```

## Hatching and emergence success
Not measured in 2018.

```{r hatching_emergence_success_nin}
tracks_nin %>%
  hatching_emergence_success() %>%
  DT::datatable(., 
                caption = "Hatching and emergence success summary", 
                options = list(paging = F))
```

## Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Disturbance or Predation" and appear here as "dist_(place)".

### Disturbance and predation of nests
Turtle nests with recorded disturbance or predation.

```{r dist_nests_nin}
disturbed_nests_nin_2018 <- tracks_nin %>% filter_2018() %>% filter(disturbance == "present")
disturbed_nests_nin_2018 %>% map_tracks()
```

There were **`r count(disturbed_nests_nin_2018)` reports** in 2018. 
Caveat: QA to exclude possible training records.

Coming soon: details of disturbance to nests.

### General and nest disturbance and predator presence
There were **`r count(dist_nin)` reports** of general disturbance or predator presence.
These counts include both disturbed nests (encounter type "nest") and stand-alone signs
of predator presence or disturbance (encounter type "other").
Those records deemed (by observer or QA) having "negligible" impact or significance are excluded here.

```{r map_dist_nin}
dist_nin %>% filter_2018() %>% group_by(season, disturbance_cause, encounter_type) %>% tally() %>% arrange(-n) %>% kable()
dist_nin %>% filter_2018() %>% group_by(season, site_name, disturbance_cause, encounter_type) %>% tally() %>% arrange(-n) %>% dt
dist_nin %>% filter_2018() %>% map_dist()
```

# Surveys
Curators are to QA 
[all `r placename_nin` surveys on WAStD](https://tsc.dbca.wa.gov.au/admin/observations/survey/)
following the 
[QA instructions for surveys](https://wastd.readthedocs.io/data_curators.html#data-flow-of-surveys).

### Per season effort
Highest level of aggregation: entire season.
Assumption: all training surveys have been marked as such in WAStD.

#### Excluding training

```{r survey_season_stats_all}
surveys_nin %>% survey_season_stats %>% dt
```

### Per day effort
Lowest level of aggregation: daily.

```{r survey_count}
surveys_nin %>% survey_count_heatmap(placename_nin, prefix_nin)
surveys_nin %>% list_survey_count(placename_nin)
surveys_nin %>% survey_hours_heatmap(placename_nin, prefix_nin)
surveys_nin %>% list_survey_effort(placename_nin)
```

### Individual surveys
No aggregation: individual surveys.

```{r surveys_list}
surveys_nin %>% survey_show_detail %>% dt()
```


## Survey effort per primary reporter

### Caveat
This list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text.
To correctly report on team member effort, curators have to read the list of 
team members in the Site Visit "comments" field and attach the correct users 
to the Survey team list.
Once the surveys are comprehensively attributed to all volunteers present, this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with
team rosters.

```{r survey_effort_by_person}
personal_effort_nin <- surveys_nin %>% survey_hours_per_person()
personal_effort_nin %>% kable()
```

## Volunteer effort 

### Season 2018-19
Number of surveys per sector times length of site.

In subsection "North West Cape - Bundegi - South to North", 
`r survey_ground_covered(surveys, 60, 2, 2018)` km were walked 
in `r survey_count(surveys, 60, 2018)` surveys.

TODO: table of survey volunteer effort at all subsections (sites).

# Data upload

```{r data_upload}
tracks_nin %>% dplyr::select(-obs, -photos) %>% 
  readr::write_csv(glue::glue("{prefix_nin}_tracks.csv"))
surveys_nin %>% readr::write_csv(glue::glue("{prefix_nin}_surveys.csv"))
dist_nin %>% dplyr::select(-photos) %>% 
  readr::write_csv(glue::glue("{prefix_nin}_disturbance.csv"))

# Package all output files into one ZIP archive
products_fn_nin <- glue::glue("{prefix_nin}_products.zip")
if (file.exists(products_fn_nin)) file.remove(products_fn_nin)
products <- list.files(pattern=prefix_nin)
zip(zipfile = products_fn_nin, files = products)

# Create a resource for the ZIP archives
# d <- ckanr::package_show("turtle-tracks")
# r <- resource_create(package_id=d$id, name="Ningaloo Turtle Nesting Census", upload="ningaloo.html")
# r <- resource_create(package_id=d$id, name="Ningaloo outputs", upload=products_fn_nin)

# ckanr::resource_update("0c35791e-c478-451e-94c6-9c82d03c721d", "ningaloo.html")
# ckanr::resource_update("7c141bdb-f3b0-4a89-8465-b6aff1d43679", products_fn_nin)
```

## Google Drive
A copy of this report and all generated outputs is uploaded to 
Google Drive and shared via 
[this link](https://drive.google.com/drive/folders/1ovTwAfUXm8hCK7ct_neemwHCQ3TAHbfF?usp=sharing) 
with DBCA external collaborators.

Every machine has to be authenticated with Google Drive once.
As this process involves browser windows and pasting of approval codes into the 
R Console, it is not run automatically when compiling this workbook.

Running the following chunk manually once per machine will cache the authentication
token locally for future use.

```{r google_drive_auth, eval=F}
googledrive::drive_auth(use_oob = TRUE)
```

The following code refreshes the [shared folder on Google Drive](https://drive.google.com/drive/folders/1ovTwAfUXm8hCK7ct_neemwHCQ3TAHbfF?usp=sharing) 
with this report and a zip archive of all products.

```{r google_drive_upload, eval=F}
googledrive::drive_ls(prefix_nin) %>% googledrive::drive_rm(.)
googledrive::drive_upload("ningaloo.html", path=glue::glue("{prefix_nin}/ningaloo.html"))
products %>% purrr::map(googledrive::drive_upload, path = googledrive::as_dribble(prefix_nin))
```

