---
title: "Ningaloo Turtle Nesting Census"
author: "Ningaloo Turtle Program and North West Shelf Flatbacks WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
source("tracks_setup.R")
source("tracks_helpers.R")
source("load_data.R")
```


```{r filter_data}
animals_nin <- animals %>% filter_nin
animals_nin_rescued <- animals_nin %>% 
  dplyr::filter(
    health %in% c("na", "other", "alive", "alive-injured"), 
    encounter_type == "inwater"
  )

animals_nin_stranded <- animals_nin %>% 
  dplyr::filter(
    health %in% c(
      "alive-then-died",
      "dead-advanced",
      "dead-organs-intact",
      "dead-edible",
      "dead-mummified",
      "dead-disarticulated",
      "deadedible",
      "deadadvanced"
    )
  )

split_site_name <- . %>% 
  dplyr::mutate(
    division = stringr::str_split(site_name, " - ")[[1]][1],
    section = stringr::str_split(site_name, " - ")[[1]][2],
    subsection = stringr::str_split(site_name, " - ")[[1]][3]
    )

tracks_nin <- tracks_all %>% 
  filter_nin() %>% 
  exclude_training_species() %>% 
  add_nest_labels() %>% 
  split_site_name()
tracks_nin_mi <- tracks_nin %>% filter_nin_mi
tracks_nin_nwc <- tracks_nin %>% filter_nin_nwc
tracks_nin_nwc_tantabiddi <- tracks_nin %>% filter_nin_nwc_tantabiddi
tracks_nin_nwc_graveyards <- tracks_nin %>% filter_nin_nwc_graveyards
tracks_nin_nwc_hunters <- tracks_nin %>% filter_nin_nwc_hunters
tracks_nin_nwc_lighthouse <- tracks_nin %>% filter_nin_nwc_lighthouse
tracks_nin_nwc_navypier <- tracks_nin %>% filter_nin_nwc_navypier
tracks_nin_nwc_bundegi <- tracks_nin %>% filter_nin_nwc_bundegi

tracks_nin_cr <- tracks_nin %>% filter_nin_cr
tracks_nin_bn <- tracks_nin %>% filter_nin_bn
tracks_nin_cb <- tracks_nin %>% filter_nin_cb
tracks_nin_wr <- tracks_nin %>% filter_nin_wr
tracks_nin_gn <- tracks_nin %>% filter_nin_gn

surveys_nin <- surveys %>% filter_nin() %>% split_site_name()
surveys_nin_prod <- surveys_nin %>% exclude_training_surveys()

dist_pred_nin <- disturbance %>% filter_nin() %>% split_site_name()
dist_nin <- dist_pred_nin %>% filter_disturbance()
dist_nin_mi <- dist_nin %>% filter_nin_mi
dist_nin_nwc <- dist_nin %>% filter_nin_nwc
dist_nin_cr <- dist_nin %>% filter_nin_cr
dist_nin_bn <- dist_nin %>% filter_nin_bn
dist_nin_cb <- dist_nin %>% filter_nin_cb
dist_nin_wr <- dist_nin %>% filter_nin_wr
dist_nin_gn <- dist_nin %>% filter_nin_gn

pred_nin <- dist_pred_nin %>% filter_predation()
pred_nin_mi <- pred_nin %>% filter_nin_mi
pred_nin_nwc <- pred_nin %>% filter_nin_nwc
pred_nin_cr <- pred_nin %>% filter_nin_cr
pred_nin_bn <- pred_nin %>% filter_nin_bn
pred_nin_cb <- pred_nin %>% filter_nin_cb
pred_nin_wr <- pred_nin %>% filter_nin_wr
pred_nin_gn <- pred_nin %>% filter_nin_gn

# no tagged nests
nests_nin <- nests_all %>% filter_nin()
prefix_nin <- "NIN"
placename_nin <- "Ningaloo"

areas_nin <- areas %>% dplyr::filter(area_type=="Site", pk %in% 58:114)
areas_nin_nwc <- areas_nin %>% dplyr::filter(area_type=="Site")
```


# Background
This section will provide context for this document.

## About the NTP
* website
* links to data on DBCA data catalogue
* links to DBCA turtle wiki (which links Google Drive folder with products)

## About this document
This document is the primary reporting mechanism for turtle data from the WA Sea
Turtle Database (WAStD). It tells a story about the data.
This document is a live workbook: it contains code which automates
the retrieval of the data from WAStD, preparation and visualisation of the data, 
export of the data and products into files, and finally uploading of these 
products as well as of this document itself to points of dissemination, such as
the DBCA data catalogue, and Google Drive folders.

## Changes in Season 2019-20
This section summarises changes to this document in Season 2019-20.

* Predation separate from disturbance
* Add filters for sections as well
* For each of NW Cape, Bungelup, Janes Bay, Gnaraloo: species, number false 
  crawls, number of nests (table > CSV, figure > PNG)
* Tour guides want to know highest likelihood of finding hatchlings.
* Nat cons teams want to know where to put predator traps.

We need the following issues to be addressed to be able to use the data:

* Survey dates. Clarify confusion between turtle date and calendar date.
  * Added `calendar_date_awst` in local timezone to tracks.csv, surveys.csv, dist.csv.
  
* Predation versus predator tracks: Clarify predated nests vs disturbed nests vs 
  signs of predator presence and signs of disturbances
  
* Survey effort: 
  * list dates surveyed (calendar date AWST) per division. 
  * list dates surveyed (calendar date AWST) per section. 
  * list dates surveyed (calendar date AWST) per subsection/site.
  * Survey effort per subsection: tally dates surveyed per subsection/site 
    across season. How many times was each site sampled per season?
  * Duplicate surveys: list sites surveyed more than once per day
  * Figure out what to do with duplicate surveys
  * Find orphaned end points: forgot SVS, found no tracks, found no dist, did SVE.

* Reliable records of which subsections were monitored and when. At present the 
  data provided indicate that some subsections were monitored twice on some days 
  and adjacent subsections weren’t monitored at all. This is an artifact of 
  sites being started (site visit start geolocation) incorrectly against training or 
  TSC areas (based on NTP db areas) have incorrect coordinates.

* Gnaraloo Bay. We need to be able to distinguish among Divisions and Gnaraloo Bay from Janes Bay from the North West Cape, etc. Don’t appear to be able to distinguish Gnaraloo Bay at present. FIXED updated filters.
* Training records. Training records appear to still be in the data we’ve been provided. These need to be removed.
* Dead turtles and rescued turtles. We need to be able to distinguish between dead turtles and rescued turtles and at which subsection they were found. FIX: split by health
into dead (stranding) and alive (rescue).
* We have records of some turtle rescues that weren’t recorded with the tablets. Need to have the facility to add these to the database.


2 Those are split up correctly in reports: predated/disturbed nests vs stand-alone signs of pred/dist. will make two separate CSVs.

3 Depends on QA of surveys. Some work needed there from my end. Corrine could help once I have specified work to be done.

4 See report, should be there, but will make sure it's in there

5 Need to review. Shouldn't be hard.

6 Will review and remove.

7 Working on that. NTP have been reporting rescues as strandings but will find a way to distinguish. That's a new CSV I'll add for them.

8 Dani Rob is trained how to enter those.

## Effort saved by going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.
Overall, the Volunteers have saved 
`r round((nrow(tracks_nin) + nrow(surveys_nin)*2 + nrow(dist_nin)) * 2 / 60)` 
hours of mind-numbing data entry and proofreading work for 
`r nrow(tracks_nin)` Tracks, 
`r nrow(dist_nin)` Disturbances, and 
`r nrow(surveys_nin)*2` Site Visit Starts/Ends. 

# Turtle Nesting
This section visualises and summarises turtle nesting.
Data are summarised at different spatial resolutions 
(entire locality of Ningaloo, divisions, sections, subsections) 
and temporal resolutions (season, 4 week core season, weekly, daily).

## Maps
### 2018-19
All  turtle tracks or nests recorded in season 2018-19. 
Click markers to view popup with details
and link to original record on WAStD.
Toggle visibility of individual species in the legend.

```{r map_nin_2018}
tracks_nin %>% filter_2018() %>% map_tracks(cluster = T)
```


```{r map_nin_animated, eval=F, warning=FALSE, message=FALSE}
# Animated map
# A [Google API key](https://developers.google.com/maps/documentation/geocoding/start?csw=1) must be present as R environment variable `GOOGLE_MAPS_APIKEY` for the next step.
tracks_nin_2018 <- tracks_nin %>% filter_2018()
# gganimate_tracks(tracks_nin_2018, placename_nin, prefix_nin)
```

```{r nests_nin, eval=F}
# Tagged nests, currently unused
nests_nin %>%
  filter_2018() %>%
  map_nests(cluster = T) %>%
  mapview::mapshot(file = glue::glue("{prefix_nin}_tagged_nests_2018.png"))
```

## Nesting abundance
### All sites
```{r tracks_all_by_season}
tracks_nin %>% 
  nesting_type_by_season_species() %>% 
  knitr::kable(caption = "Nesting over entire season at Ningaloo")
```

### Divisions

```{r tracks_divisions_by_season}
tracks_nin_nwc %>% 
  nesting_type_by_season_species() %>% 
  kable(caption = "Nesting over entire season at Division North West Cape")
tracks_nin_bn %>% 
  nesting_type_by_season_species() %>% 
  kable(caption = "Nesting over entire season at Division Bundera Ningaloo")
# TODO insert other divisions
```

### Sections

```{r tracks_sections_by_season}
tracks_nin_nwc_bundegi %>% 
  nesting_type_by_season_species() %>% 
  kable(caption = "Nesting over entire season at Section North West Cape - Bundegi")
tracks_nin_nwc_navypier %>% 
  nesting_type_by_season_species() %>% 
  kable(caption = "Nesting over entire season at Section North West Cape - Navy Pier")
tracks_nin_nwc_lighthouse %>% 
  nesting_type_by_season_species() %>% 
  kable(caption = "Nesting over entire season at Section North West Cape - Lighthouse")
tracks_nin_nwc_graveyards %>% 
  nesting_type_by_season_species() %>% 
  kable(caption = "Nesting over entire season at Section North West Cape - Graveyards")
tracks_nin_nwc_hunters %>% 
  nesting_type_by_season_species() %>% 
  kable(caption = "Nesting over entire season at Section North West Cape - Hunters")
tracks_nin_nwc_tantabiddi %>% 
  nesting_type_by_season_species() %>% 
  kable(caption = "Nesting over entire season at Section North West Cape - Tantabiddi")

# TODO insert other sections
```

### Subsections

## Nesting over standard season
The Ningaloo Turtle Program defines its standard season as the four weeks over
the new year's break, which spans calendar weeks 51, 52, 1, and 2, or season
week 25--28.

### All sites
### Divisions
### Sections
### Subsections

## Nesting by week
### All sites
The next table show data from all sites.

```{r tracks_all_beaches_by_week_nin}
tracks_nin %>% nesting_type_by_season_week_species() %>% kable()
```

### Divisions

TODO export as CSV

```{r tracks_all_beaches_by_week_nin_divisions}
tracks_nin_nwc %>% 
  nesting_type_by_season_week_species() %>% 
  kable(caption = "Nesting by week at Division North West Cape")
tracks_nin_bn %>% 
  nesting_type_by_season_week_species() %>% 
  kable(caption = "Nesting by week at Division Bundera Ningaloo")
# TODO other divisions
```


### Sections

```{r tracks_all_beaches_by_week_nin_sections}
tracks_nin_nwc_bundegi %>% 
  nesting_type_by_season_week_species() %>% 
  kable(caption = "Nesting by week at Section North West Cape - Bundegi")
tracks_nin_nwc_navypier %>% 
  nesting_type_by_season_week_species() %>% 
  kable(caption = "Nesting by week at Section North West Cape - Navy Pier")
tracks_nin_nwc_lighthouse %>% 
  nesting_type_by_season_week_species() %>% 
  kable(caption = "Nesting by week at Section North West Cape - Lighthouse")
tracks_nin_nwc_graveyards %>% 
  nesting_type_by_season_week_species() %>% 
  kable(caption = "Nesting by week at Section North West Cape - Graveyards")
tracks_nin_nwc_hunters %>% 
  nesting_type_by_season_week_species() %>% 
  kable(caption = "Nesting by week at Section North West Cape - Hunters")
tracks_nin_nwc_tantabiddi %>% 
  nesting_type_by_season_week_species() %>% 
  kable(caption = "Nesting by week at Section North West Cape - Tantabiddi")

# TODO other sections
```

## Nesting by day

### All sites

This section shows data from all sites.
The data visualised as figure here is also exported as spreadsheet.

```{r nesting_success_nin, warning = F}
tracks_nin_day_site <- tracks_nin %>% 
  dplyr::group_by(division, section, subsection, site_name,
                  season, iso_week, season_week, calendar_date_awst, 
                  species, nest_type) %>%
  dplyr::tally() %>%
  tidyr::pivot_wider(names_from = nest_type, values_from = n) %>% 
  dplyr::arrange(calendar_date_awst, site_name, species)

tracks_ts(tracks_nin, surveys_nin_prod, placename_nin, prefix_nin)
nests <- tracks_nin %>% track_success()
nests %>% ggplot_track_success_by_date("caretta-caretta", placename_nin, prefix_nin)
nests %>% ggplot_track_successrate_by_date("caretta-caretta", placename_nin, prefix_nin)
nests %>% ggplot_track_success_by_date("chelonia-mydas", placename_nin, prefix_nin)
nests %>% ggplot_track_successrate_by_date("chelonia-mydas", placename_nin, prefix_nin)
nests %>% track_success_by_species() %>% DT::datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```

### Divisions
### Sections
### Subsections

## Hatching and emergence success
Not measured in 2018.

```{r hatching_emergence_success_nin}
tracks_nin %>%
  hatching_emergence_success() %>%
  DT::datatable(., 
                caption = "Hatching and emergence success summary", 
                options = list(paging = F))
```

# Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Disturbance or Predation" and appear here as "dist_(place)".

## Predation
Predated nests as well as signs of predator presence on nesting beaches.

```{r predation}
pred_nin %>% DT::datatable(caption="Predation at Ningaloo")

pred_nests_nin_2018 <- tracks_nin %>% filter_2018() %>% 
  filter(disturbance == "present")
pred_nests_nin_2018 %>% map_tracks()
```


## Disturbance
Disturbed nests as well as disturbance on nesting beaches.

```{r disturbance}
# TODO map, table
```


## OLD: Disturbance and predation of nests
Turtle nests with recorded disturbance or predation.

```{r dist_nests_nin}
disturbed_nests_nin_2018 <- tracks_nin %>% filter_2018() %>% filter(disturbance == "present")
disturbed_nests_nin_2018 %>% map_tracks()
```

There were **`r count(disturbed_nests_nin_2018)` reports** in 2018. 
Caveat: QA to exclude possible training records.

Coming soon: details of disturbance to nests.

### OLD General and nest disturbance and predator presence
There were **`r count(dist_nin)` reports** of general disturbance and **`r count(pred_nin)` reports** predator presence.

TODO is the following true:
These counts include both disturbed nests (encounter type "nest") and stand-alone signs
of predator presence or disturbance (encounter type "other").
Those records deemed (by observer or QA) having "negligible" impact or significance are excluded here.

```{r map_dist_nin}
dist_nin %>% filter_2018() %>% group_by(season, disturbance_cause, encounter_type) %>% tally() %>% arrange(-n) %>% kable()
dist_nin %>% filter_2018() %>% group_by(season, site_name, disturbance_cause, encounter_type) %>% tally() %>% arrange(-n) %>% dt
dist_nin %>% filter_2018() %>% map_dist()
```

# Surveys
Curators are to QA 
[all `r placename_nin` surveys on WAStD](https://tsc.dbca.wa.gov.au/admin/observations/survey/)
following the 
[QA instructions for surveys](https://wastd.readthedocs.io/data_curators.html#data-flow-of-surveys).

TODO update QA instructions for duplicate surveys after TSC survey reconstruction.

## Per season effort
Highest level of aggregation: entire season.
Assumption: all training surveys have been marked as such in WAStD.

### Excluding training

```{r survey_season_stats_all}
surveys_nin %>% survey_season_stats %>% dt
```

## Per day effort
Lowest level of aggregation: daily.

```{r survey_count}
surveys_nin %>% survey_count_heatmap(placename_nin, prefix_nin)
surveys_nin %>% list_survey_count(placename_nin)
surveys_nin %>% survey_hours_heatmap(placename_nin, prefix_nin)
surveys_nin %>% list_survey_effort(placename_nin)
```

## Individual surveys
No aggregation: individual surveys.

```{r surveys_list}
surveys_nin %>% survey_show_detail %>% dt()
```


## Survey effort per primary reporter

### Caveat
This list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text.
To correctly report on team member effort, curators have to read the list of 
team members in the Site Visit "comments" field and attach the correct users 
to the Survey team list.
Once the surveys are comprehensively attributed to all volunteers present, this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with
team rosters.

```{r survey_effort_by_person}
personal_effort_nin <- surveys_nin %>% survey_hours_per_person()
personal_effort_nin %>% kable()
```

## Volunteer effort 

### Season 2018-19
Number of surveys per sector times length of site.

In subsection "North West Cape - Bundegi - South to North", 
`r survey_ground_covered(surveys, 60, 2, 2018)` km were walked 
in `r survey_count(surveys, 60, 2018)` surveys.

TODO: table of survey volunteer effort at all subsections (sites).

## Training dates

* TODO insert dates
* Action required: Find duplicate surveys at NIN in TSC and mark as training.
* Update 2019-10: TSC Areas at NIN were built from incorrect data. New NIN areas will be added, surveys and sites will be reconstructed again. 

## Provided username mismatches WAStD users

* TODO: get Chantelle's list of username mismatches > update TSC users (aliases).

```{r name_mismatch_nin}
surveys_nin %>% filter_surveys_requiring_qa() %>% dt()
```


# Data upload
In this section, this report, exported data, and exported figures are
uploaded to the DBCA data catalogue as well as to a Google Drive folder.

```{r data_upload}
tracks_nin %>% dplyr::select(-obs, -photos) %>% 
  readr::write_csv(glue::glue("{prefix_nin}_tracks.csv"))
tracks_nin_day_site %>% 
  readr::write_csv(glue::glue("{prefix_nin}_tracks_by_day.csv"))

surveys_nin %>% readr::write_csv(glue::glue("{prefix_nin}_surveys.csv"))
dist_nin %>% dplyr::select(-photos) %>% 
  readr::write_csv(glue::glue("{prefix_nin}_disturbance.csv"))

# Package all output files into one ZIP archive
products_fn_nin <- glue::glue("{prefix_nin}_products.zip")
if (file.exists(products_fn_nin)) file.remove(products_fn_nin)
products <- list.files(pattern=prefix_nin)
zip(zipfile = products_fn_nin, files = products)

# Create a resource for the ZIP archives
# d <- ckanr::package_show("turtle-tracks")
# r <- resource_create(package_id=d$id, name="Ningaloo Turtle Nesting Census", upload="ningaloo.html")
# r <- resource_create(package_id=d$id, name="Ningaloo outputs", upload=products_fn_nin)

# ckanr::resource_update("0c35791e-c478-451e-94c6-9c82d03c721d", "ningaloo.html")
# ckanr::resource_update("7c141bdb-f3b0-4a89-8465-b6aff1d43679", products_fn_nin)
```

## Google Drive
A copy of this report and all generated outputs is uploaded to 
Google Drive and shared via 
[this link](https://drive.google.com/drive/folders/1ovTwAfUXm8hCK7ct_neemwHCQ3TAHbfF?usp=sharing) 
with DBCA external collaborators.

Every machine has to be authenticated with Google Drive once.
As this process involves browser windows and pasting of approval codes into the 
R Console, it is not run automatically when compiling this workbook.

Running the following chunk manually once per machine will cache the authentication
token locally for future use.

```{r google_drive_auth, eval=F}
googledrive::drive_auth(use_oob = TRUE)
```

The following code refreshes the [shared folder on Google Drive](https://drive.google.com/drive/folders/1ovTwAfUXm8hCK7ct_neemwHCQ3TAHbfF?usp=sharing) 
with this report and a zip archive of all products.

```{r google_drive_upload, eval=F}
googledrive::drive_ls(prefix_nin) %>% googledrive::drive_rm(.)
googledrive::drive_upload("ningaloo.html", path=glue::glue("{prefix_nin}/ningaloo.html"))
products %>% purrr::map(googledrive::drive_upload, path = googledrive::as_dribble(prefix_nin))
```

