---
title: "Care for Hedland Season Review 2017-2018"
author: "Care for Hedland with Marine Turtles WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 9
    fig_height: 5
---

```{r r setup, include=FALSE, message=FALSE, warning=FALSE}
source("tracks_setup.R")
source("tracks_helpers.R")
```

# Data
```{r, eval=T}
source("load_data.R")
tracks_pth <- tracks_all %>% filter_port_hedland_sites() %>% filter_2017()
tracks_pth_cem <- tracks_all %>% filter_port_hedland_cemetery() %>% filter_2017()
tracks_pth_ppo <- tracks_all %>% filter_port_hedland_prettypool() %>% filter_2017()
surveys_pth <- surveys %>% filter_port_hedland_sites() %>% filter_2017()
dist_pth <- disturbance %>% filter_port_hedland_sites() %>% filter_2017()
```

# Season summary

## General notes
Nests without a clearly visible emergence point were not recorded.
This led to under-reporting of hatched nests this season.

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.
This season, the Care for Hedland Volunteers have saved Jo and the Kellies 
`r (nrow(tracks_pth) + nrow(surveys_pth)*2) * 2 / 60` hours of mind-numbing data entry
and proofreading work for `r nrow(tracks_pth)` Track or Treats and 
`r nrow(surveys_pth)*2` Site Visit Starts/Ends. 
This estimate excludes Fox Sakes and Marine Wildlife Incidents, about 2-3h extra!

## Maps
```{r map_overall}
tracks_pth_cem %>% add_nest_labels() %>% map_tracks()
tracks_pth_ppo %>% add_nest_labels() %>% map_tracks()
```

## Nesting abundance
```{r tracks_all_sectors}
tracks_pth %>% species_by_type() %>% kable()
```

## Cemetery Beach
```{r tracks_cem}
tracks_pth_cem %>% species_by_type() %>% kable()
```

## Pretty Pool Beach
```{r tracks_ppo}
tracks_pth_ppo %>% species_by_type() %>% kable()
```

## Nesting success 
```{r nesting_success, fig.height=5, fig.width=7, warning = F}
tracks_pth %>% tracks_ts()
tracks_pth_cem %>% tracks_ts()
tracks_pth_ppo %>% tracks_ts()

nests <- tracks_pth %>% track_success()
label <- "pth"

nests %>% ggplot_track_success_by_date("natator-depressus", label)
nests %>% ggplot_track_success_by_date("chelonia-mydas", label)
nests %>% ggplot_track_successrate_by_date("natator-depressus", label)
nests %>% ggplot_track_successrate_by_date("chelonia-mydas", label)
nests %>% track_success_by_species() %>% datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```

## Hatching and emergence success
```{r hatching_emergence_success}
tracks_pth %>%
  hatching_emergence_success() %>%
  DT::datatable(.,
    caption = "Hatching and emergence success summary",
    options = list(paging = F)
  )
```

## Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Fox Sake" and appear here as "dist_(place)".

### Disturbance and predation of nests
Turtle nests with recorded disturbance or predation.

```{r}
disturbed_nests_pth <- tracks_pth %>% filter(disturbance == "present")
disturbed_nests_pth %>% add_nest_labels() %>% map_tracks()
```

There were **`r count(disturbed_nests_pth)` reports** of disturbed nests. Caveat: QA to exclude possible training records.

Coming soon: details of disturbance to nests.

### General disturbance
Coming soon: map.

There were **`r count(dist_pth)` reports** of general disturbance or predator presence:

```{r}
dist_pth %>% group_by(disturbance_cause) %>% tally() %>% arrange(-n) %>% kable()
```

## Survey count

### Caveat
* Data includes training surveys.

This contributes to over-estimation of effort, but shows actual boots on the 
ground - training is also an effort!

```{r}
place <- "Port Hedland"
surveys_pth %>% plot_survey_count(place)
surveys_pth %>% list_survey_count(place)
```

## Survey effort in hours
Q How many hours did volunteers survey, how many kilometers were walked?

### Caveat
* Surveys with missing end points are auto-closed after 5 hours. 
* Data includes training surveys.

Both factors contribute to over-estimation of effort.

```{r survey_effort}
surveys_pth %>% plot_survey_effort(place)
surveys_pth %>% list_survey_effort(place)
```

## Survey effort per person
Q Which volunteer did the most hours overall on survey?

### Caveat
This list only includes the primary reporter. This leads to under-reporting, as time spent
as team is not reported on.

```{r survey_effort_by_person}
personal_effort <- surveys_pth %>% survey_hours_per_person()
personal_effort %>% kable()
```

## Survey effort in kilometers walked
Number of surveys per sector times length of site.

At Cemetery Beach, `r survey_ground_covered(surveys, 35, 1.6)` km were walked 
in `r survey_count(surveys_pth, 35)` surveys.

At Pretty Pool Beach, `r survey_ground_covered(surveys, 45, 1.8)` km were walked 
in `r survey_count(surveys_pth, 45)` surveys.

# Data upload

```{r data_upload}
# d <- ckanr::package_show("turtle-tracks")
ckanr::resource_update("b35fb19e-e72d-420c-b875-12b4db4afe8c", "pthedland_2017-18.html")
```

