---
title: "WA Turtles Summary and Quality Control"
author: "Florian Mayer, WA Turtles DBCA"
date: "9/20/2018"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("tracks_setup.R")
source("tracks_helpers.R")
source("load_data.R")
```

# Turtle nesting census

## All seasons
```{r}
tracks_all %>%
  exclude_training_species() %>% 
  dplyr::group_by(season, species) %>% 
  dplyr::tally() %>% 
  dplyr::arrange(season, -n) %>% 
  dt0
```

## Nesting 2016-17
```{r maps_2016}
tracks_2016 <- tracks_all %>% 
  filter_2016() %>% 
  exclude_training_species() %>%
  add_nest_labels()

tracks_2016 %>% map_tracks(cluster=TRUE)
tracks_2016 %>% nesting_type_by_season_species()
```


## Nesting 2017-18
```{r maps_2017}
tracks_all %>% 
  filter_2017 %>% 
  exclude_training_species() %>%
  add_nest_labels() %>% 
  map_tracks(cluster=TRUE)

tracks_all %>% 
  filter_2017 %>% 
  nesting_type_by_season_species()
```

## Nesting 2018-19
```{r maps_2018}
tracks_all %>% 
  filter_2018 %>% 
  exclude_training_species() %>%
  add_nest_labels() %>% 
  map_tracks(cluster=TRUE)

tracks_all %>% 
  filter_2018 %>% 
  nesting_type_by_season_species()
```


## Disturbance
```{r dist}
disturbance %>% disturbance_by_season() %>% dplyr::arrange(season, -n)
disturbance %>% map_dist(cluster=T)
```

## Survey effort

### Including training
```{r surv_incl_tr}
surveys %>% survey_hours_heatmap(placename = "Western Australia", prefix = "WA")
```

### Including training
```{r surv_excl_tr}
surveys_prod <- surveys %>% exclude_training_surveys()
surveys_prod %>% survey_season_stats() %>% dt0
surveys_prod %>% survey_hours_heatmap(placename = "Western Australia", prefix = "WA")
```


# QA
This chapter supports various QA tasks.

Which data did you QA?

Surveys S3 https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=24
Surveys S2 https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=23
Surveys S1 https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=22
Turtle nests (all sectors) https://tsc.dbca.wa.gov.au/admin/observations/turtlenestencounter/?area__id__exact=19
Tagged turtle nests (a subset of turtle nests): https://tsc.dbca.wa.gov.au/admin/observations/nesttagobservation/?encounter__area__id__exact=19
Disturbed nests broome: https://tsc.dbca.wa.gov.au/admin/observations/encounter/?area__id__exact=19&encounter_type__exact=other&source__exact=odk

What did you QA for?

Unexpected species e.g. Hawksbill is https://tsc.dbca.wa.gov.au/admin/observations/turtlenestencounter/?area__id__exact=19&species__exact=eretmochelys-imbricata (filter to other unexpected species)
Unidentified turtle species https://tsc.dbca.wa.gov.au/admin/observations/turtlenestencounter/?area__id__exact=19&species__exact=cheloniidae-fam
Missing tag label (e.g. sector 3 https://tsc.dbca.wa.gov.au/admin/observations/nesttagobservation/?encounter__site__id__exact=24&o=-10, same for sectors 2 and 1)
Are training surveys (multiple surveys on one day in the tetris chart is a good indication) on training days marked as "not production"?
    
    

## Missing sites
* If tracks are just outside of known sites: Extend WAStD site polygons to include GPS error.
* Add WAStD sites for new surveyed sites (production or training) and include those sites in existing or new report.
* Change confirmed training records outside training areas to "Hatchback turtles" and mark as "curated" to
  prevent from being overwritten by repeated data loading.
* Note: re-run uncached data import after changing tracks.

## Tracks outside known sites
Review whether there are stray tracks close to known sites.

* WAStD sites missing (new sites) or boundaries set too narrow. 
  Extend sites, re-save stray tracks to auto-repair their site affiliation.
* Training records close to site (but not meant to be within site).
  Change confirmed training records outside training areas to "Hatchback turtles" 
  and mark as "curated" to prevent from being overwritten by repeated data loading.
* Note: re-run uncached data import after changing tracks.

```{r tracks_missing_sites}
tracks_all %>% 
  exclude_training_species() %>%
  filter_missing_site() %>%
  add_nest_labels() %>% 
  map_tracks(cluster=TRUE)
```

## Disturbance outside known sites
Review disturbances close to known sites.

Disturbances can either be legitimate, although opportunistic, records, or training
records.

* Training records: set to "Other, see comments", mention "training" in comments,
  ideally instruct data collectors to photograph their writing board with "training"
  written on it to make absolutely clear that this record is a training record.

```{r dist_missing_sites}
disturbance %>% filter_missing_site() %>% map_dist()
```



## Surveys

### Tracks with missing surveys
```{r missing_surveys}
missing_surveys <- tracks_all %>%
  filter_missing_survey() %>%
  dplyr::group_by(site_id, site_name, turtle_date) %>%
  dplyr::summarise(earliest_record = min(datetime), 
                   latest_record = max(datetime), 
                   survey_start_time = min(datetime) - lubridate::minutes(30),
                   survey_end_time = max(datetime) + lubridate::minutes(30),
                   reporter = first(reporter),
                   no_tracks = n()) 
missing_surveys %>% dt
```


## Surveys missing Site Visit End
Where a "Site Visit End" form was forgotten, `end_source_id` is empty, and the duration is set to 6h. 
These surveys should be closed a few minutes after the last recorded track or disturbance.

Surveys with missing end contribute to an over-estimation of survey effort.

This task is now automated, this section is only for QA.
  
```{r surveys_missing_end}
surveys %>% filter_surveys_missing_end() %>% dt()
```


## Nest tags

### Tagged nests
Nest tags are often malformed. Use the "Update in WAStD" button and extract "tag name" and 
"date laid" into their own top level fields in WAStD.

```{r nests_all}
nests_all %>% map_nests()
```


## Username mismatches
TODO: define QA rules for mismatches of usernames as (possibly mis)typed by data collectors vs 
[WAStD users](https://tsc.dbca.wa.gov.au/admin/users/user/).

```{r name_mismatch}
surveys %>% filter_surveys_requiring_qa() %>% dt()
```

# State-wide season summary
This section contains state-wide season stats.

## Survey effort

For each monitoring program:

* Number of volunteers
* Hours surveyed (with/without training)
* Kms walked (with/without training)
* Number of days surveyed
* Season length in days (days between first and last season survey)

## Nesting

For each monitoring program / for each site:

* Rookery size (# nests per season per species)
* Track density (# nesting emergences per km of coastline)
* Hatching and emergence success with number of excavated nests
* Percentage of nests disturbed
* Number of disturbances recorded
* Density of disturbances per km of coastline
* Disturbance causes: number and category

```{r}
tracks_all %>% exclude_training_species() %>% nesting_type_by_season_species() %>% dt0
tracks_all %>% exclude_training_species() %>% nesting_type_by_area_season_species() %>% dt
tracks_all %>% exclude_training_species() %>% nesting_type_by_site_season_species() %>% dt

tracks_all %>% hatching_emergence_success %>% dt0
tracks_all %>% hatching_emergence_success_area %>% dt
tracks_all %>% hatching_emergence_success_site %>% dt
```

## Disturbance

```{r}
disturbance %>% 
  group_by(area_name, season, disturbance_cause) %>% 
  tally() %>% 
  filter(n>3) %>% 
  filter(!is.null(area_name)) %>% 
  arrange(area_name, season, -n) %>% 
  dt
```

# Data upload
Data are uploaded to the [turtle nest census dataset](https://data.dpaw.wa.gov.au/dataset/turtle-tracks) 
on the departmental data catalogue, accessible from the DBCA intranet only.

```{r data_upload_ckan}
# Create a resource for the QA workbook
# d <- ckanr::package_show("turtle-tracks")
# r <- resource_create(package_id=d$id, name="Turtle Nesting Census Quality Control", upload="QA.html")

# Update resources on data catalogue
ckanr::resource_update("be63cdd8-4aba-4329-b333-8090b076c792", "QA.html")
```
