---
title: "WA Turtles Quality Control"
author: "Florian Mayer, WA Turtles DBCA"
date: "9/20/2018"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("tracks_setup.R")
source("tracks_helpers.R")
source("load_data.R")
```

# Missing sites
* If tracks are just outside of known sites: Extend WAStD site polygons to include GPS error.
* Add WAStD sites for new surveyed sites (production or training) and include those sites in existing or new report.
* Change confirmed training records outside training areas to "Hatchback turtles" and mark as "curated" to
  prevent from being overwritten by repeated data loading.
* Note: re-run uncached data import after changing tracks.

## Tracks outside known sites
```{r tracks_missing_sites}
tracks_all %>% 
  filter_realspecies() %>%
  filter_nosite() %>%
  add_nest_labels() %>% 
  map_tracks()
```

## Disturbance outside known sites
```{r dist_missing_sites}
disturbance %>% filter_nosite() %>% map_dist()
```



## Missing surveys

### Tracks with missing surveys
```{r missing_surveys}
missing_surveys <- tracks_all %>%
  filter_nosurvey() %>%
  dplyr::group_by(site_id, site_name, date) %>%
  dplyr::summarise(earliest_record = min(datetime), 
                   latest_record = max(datetime), 
                   survey_start_time = min(datetime) - lubridate::minutes(30),
                   survey_end_time = max(datetime) + lubridate::minutes(30),
                   reporter = first(reporter),
                   no_tracks = n()) 
missing_surveys %>% dt
```


## Nest tags

### Tagged nests
Nest tags are often malformed. Use the "Update in WAStD" button and extract "tag name" and 
"date laid" into their own top level fields in WAStD.

```{r nests_all}
nests_all %>% map_nests()
```
