---
title: "Season review 2017-2018: Delambre and Thevenard Islands"
author: "Marine Turtles WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 9
    fig_height: 5
---

```{r r setup, include=FALSE, message=FALSE, warning=FALSE}
source("tracks_setup.R")
source("tracks_helpers.R")
```

# Caveat
The data shown are pending QA and reconciliation with paper-based records. 
Until this warning disappears, this dataset may contain duplicates from training runs, effort attributed to the wrong operators, and possibly other issues and biases.

No data returned from Rosemary Islands. Need to deliver better/more/stricter training.

# Data
```{r, eval=T}
source("load_data.R")
tracks_thv <- tracks_all %>% filter_thv() %>% filter_2017()
tracks_di <- tracks_all %>% filter_di() %>% filter_2017()

surveys_thv <- surveys %>% filter_thv() %>% filter_2017()
surveys_di <- surveys %>% filter_di() %>% filter_2017()

dist_thv <- disturbance %>% filter_thv() %>% filter_2017()
dist_di <- disturbance %>% filter_di() %>% filter_2017()
```

# Season summary

## General notes
Nests without a clearly visible emergence point were not recorded.
This led to under-reporting of hatched nests this season.

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.
This season, the Volunteers have saved 
`r (nrow(tracks_thv) + nrow(surveys_thv)*2 + nrow(dist_thv)) * 2 / 60` hours of mind-numbing data entry
and proofreading work for `r nrow(tracks_thv)` Track or Treats and 
`r nrow(surveys_thv)*2` Site Visit Starts/Ends. 

Time savings Delambre: `r (nrow(tracks_di) + nrow(surveys_di)*2 + nrow(dist_di)) * 2 / 60` hours for `r nrow(tracks_di)` Track or Treats and 
`r nrow(surveys_di)*2` Site Visit Starts/Ends. 

## Maps
```{r map_overall}
tracks_thv %>% add_nest_labels() %>% map_tracks()
tracks_di %>% add_nest_labels() %>% map_tracks()
```

## How many nests did we have all together across all sectors?

### Thevenard
```{r tracks_all_sectors_thv}
tracks_thv %>% species_by_type() %>% kable()
```

```{r nesting_success_thv}
tracks_thv %>% tracks_ts()
```

### Delambre
```{r tracks_all_sectors_di}
tracks_di %>% species_by_type() %>% kable()
```

```{r nesting_success_di}
tracks_di %>% tracks_ts()
```


## Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Fox Sake" and appear here as "dist_(place)".

### Disturbance and predation of nests Thevenard
Turtle nests with recorded disturbance or predation.

```{r}
disturbed_nests_thv <- tracks_thv %>% filter(disturbance=="present")
disturbed_nests_thv %>% add_nest_labels %>% map_tracks
```

There were **`r count(disturbed_nests_thv)` reports** of disturbed nests. Caveat: QA to exclude possible training records.


### Disturbance and predation of nests Delambre
Turtle nests with recorded disturbance or predation.

```{r}
disturbed_nests_di <- tracks_di %>% filter(disturbance=="present")
disturbed_nests_di %>% add_nest_labels %>% map_tracks
```

There were **`r count(disturbed_nests_di)` reports** of disturbed nests. Caveat: QA to exclude possible training records.

Coming soon: details of disturbance to nests.

### General disturbance
Coming soon: map.

There were **`r count(dist_thv)` reports** of general disturbance or predator presence on Thevenard Is, and **`r count(dist_di)` reports** on Delambre Is.

```{r}
dist_thv %>% group_by(disturbance_cause) %>% tally %>% arrange(-n) %>% kable
dist_thv %>% map_dist

dist_di %>% group_by(disturbance_cause) %>% tally %>% arrange(-n) %>% kable
dist_di %>% map_dist
```


## Survey count

### Caveat
* Data includes training surveys.

This contributes to over-estimation of effort, but shows actual boots on the 
ground - training is also an effort!

```{r survey_count}
place <- "Thevenard Is"
surveys_thv %>% plot_survey_count(place)
surveys_thv %>% list_survey_count(place)

place <- "Delambre Is"
surveys_di %>% plot_survey_count(place)
surveys_di %>% list_survey_count(place)
```

## Survey effort in hours
Q How many hours did volunteers survey, how many kilometers were walked?

### Caveat
* Surveys with missing end points are auto-closed after 5 hours. 
* Data includes training surveys.

Both factors contribute to over-estimation of effort.

```{r survey_effort}
surveys_thv %>% plot_survey_effort("Thevenard Is")
surveys_thv %>% list_survey_effort("Thevenard Is")

surveys_di %>% plot_survey_effort("Delambre Is")
surveys_di %>% list_survey_effort("Delambre Is")
```

## Survey effort per primary reporter
Q Which volunteer did the most hours overall on survey?

### Caveat
This list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text.
To correctly report on team member effort, curators have to read the list of 
team members in the Site Visit "comments" field and attach the correct users 
to the Survey team list.
Once the surveys are comprehensively attributed to all volunteers present, this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with
team rosters.

```{r survey_effort_by_person}
personal_effort_thv <- surveys_thv %>% survey_hours_per_person()
personal_effort_thv %>% kable()

personal_effort_di <- surveys_di %>% survey_hours_per_person()
personal_effort_di %>% kable()
```

## Survey effort in kilometers walked
Number of surveys per sector times length of site.

At Thevenard North Beach, assuming the trip was 6 km,
`r survey_ground_covered(surveys, 28, 6)` km were walked 
in `r survey_count(surveys_thv, 28)` surveys.

At Thevenard South Beach, assuming the trip was 6 km,
`r survey_ground_covered(surveys, 29, 6)` km were walked 
in `r survey_count(surveys_thv, 29)` surveys.

At Thevenard Nesting Beach, assuming the trip was 1 km,
`r survey_ground_covered(surveys, 20, 1)` km were walked 
in `r survey_count(surveys_thv, 20)` surveys.

At Delambre, assuming the trip was 4 km,
`r survey_ground_covered(surveys, 39, 4)` km were walked 
in `r survey_count(surveys_thv, 39)` surveys.

# Data upload

```{r data_upload}
# d <- ckanr::package_show("turtle-tracks")
# ckanr::resource_update("b35fb19e-e72d-420c-b875-12b4db4afe8c", "pthedland_2017-18.html")
```

```{r data_export}
tracks_thv %>% select(-obs, -photos) %>% readr::write_csv("../data/tracks_thv.csv")
surveys_thv %>% readr::write_csv("../data/surveys_thv.csv")
dist_thv %>% select(-photos) %>% readr::write_csv("../data/dist_thv.csv")

tracks_di %>% select(-obs, -photos) %>% readr::write_csv("../data/tracks_di.csv")
surveys_di %>% readr::write_csv("../data/surveys_di.csv")
dist_di %>% select(-photos) %>% readr::write_csv("../data/dist_di.csv")
```

