---
title: "Thevenard Island Turtle Nesting Census"
author: "Marine Turtles WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: show
    theme: lumen
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
source("tracks_setup.R")
source("tracks_helpers.R")
source("load_data.R")
```

# Caveat
Particular to Thevenard Island is that all training occurs at the accommodation 
area, which is registered  as a separate site, and explicitly not included in the data analysed here.

Training and test data from Delambre Island are not yet separated out.

Data from Rosemary Islands was recovered from unfinalised forms on devices.

# Data
```{r load_data}
tracks_thv <- tracks_all %>% filter_thv() %>% exclude_training_species() %>% add_nest_labels()
surveys_thv <- surveys %>% filter_thv()
# surveys prod: training is done in separate area and already excluded.
dist_thv <- disturbance %>% filter_thv()
nests_thv <- nests_all %>% filter_thv()
prefix_thv <- "THV"
placename_thv <- "Thevenard Is"
```

# QA
Most QA is done across all WA records and programs in the separate QA workbook.

This section identifies data requiring editing.

## Provided username mismatches WAStD users
```{r name_mismatch_thv}
surveys_thv %>% filter_surveys_requiring_qa() %>% dt()
```

# Background
Source: Milani Chaloupka, presentation to DBCA / NWS committee 8/5/2018.

Demographic focus is essential for diagnosing population status and trend.
Survival rates, conditional breeding rates, hatchling production rates, recruitment rates and population abundance.

Three major flatback pops in Pilbara: Delambre 3.5k nesting/y, 
Mundabullangana 2.5k nesting/y, Barrow Is 2k/y.
Biggest pop in GBR is 700 nests/y.

Milani presented: 

* Bayesian gompertz state-space model trend
* Random effects forest plot with rookery specific posterior distribution

# Season summary

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.
This season, the Volunteers at Thevenard have saved 
`r (nrow(tracks_thv) + nrow(surveys_thv)*2 + nrow(dist_thv)) * 2 / 60` hours of mind-numbing data entry
and proofreading work for `r nrow(tracks_thv)` Track or Treats and 
`r nrow(surveys_thv)*2` Site Visit Starts/Ends. 

# Thevenard Nesting
## Maps
```{r map_thv}
tracks_thv %>% filter_2017() %>% map_tracks()
tracks_thv %>% filter_2018() %>% map_tracks()
```

### Animated map
A [Google API key](https://developers.google.com/maps/documentation/geocoding/start?csw=1) must
be present as R environment variable `GOOGLE_MAPS_APIKEY` for the next step.

```{r map_thv_animated, warning=FALSE, eval=FALSE}
tracks_thv_2017 <- tracks_thv %>% filter_2017()
# gganimate_tracks(tracks_thv_2017, placename_thv, prefix_thv)
```

### Tagged nests
The map is saved as a PNG file `r glue::glue(prefix_thv, "_tagged_nests.png")`.

```{r nests_thv}
nests_thv %>%
  filter_2017() %>%
  map_nests() %T>%
  mapview::mapshot(file = glue::glue("{prefix_thv}_tagged_nests_2017.png"))

# nests_thv %>%
#   filter_2018() %>%
#   map_nests() %T>%
#   mapview::mapshot(file = glue::glue("{prefix_thv}_tagged_nests_2018.png"))
```

## Nesting abundance
### All beaches
```{r tracks_all_sectors_thv}
tracks_thv %>% nesting_type_by_season_species() %>% kable()
```

## Nesting by week
### All beaches
The next table show data from all sites.

```{r tracks_all_beaches_by_week_thv}
tracks_thv %>% nesting_type_by_season_week_species() %>% kable()
```

## Nesting by day
This section shows data from all sites.

```{r nesting_success_thv, message=FALSE}
tracks_ts(tracks_thv, surveys_thv, placename_thv, prefix_thv)
nests <- tracks_thv %>% track_success()
nests %>% ggplot_track_success_by_date("natator-depressus", placename_thv, prefix_thv)
nests %>% ggplot_track_successrate_by_date("natator-depressus", placename_thv, prefix_thv)
# nests %>% ggplot_track_success_by_date("chelonia-mydas", placename_thv, prefix_thv)
# nests %>% ggplot_track_successrate_by_date("chelonia-mydas", placename_thv, prefix_thv)
nests %>% track_success_by_species() %>% DT::datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```

## Hatching and emergence success
Values in `clutch_size_fresh` indicate the inclusion of records from tagging/nesting observations 
where eggs were counted and the "Track or Treat" turtle track count form was used to record the number of eggs.
Only rows with empty `clutch_size_fresh` show trustworthy data.

```{r hatching_emergence_success_thv}
tracks_thv %>%
  hatching_emergence_success() %>%
  DT::datatable(., caption = "Hatching and emergence success summary", options = list(paging = F))
```

## Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Fox Sake" and appear here as "dist_(place)".

### Disturbance and predation of nests
Turtle nests with recorded disturbance or predation.

```{r dist_nests_thv}
disturbed_nests_thv <- tracks_thv %>% filter_2018() %>% filter(disturbance == "present")
disturbed_nests_thv %>% map_tracks()
```

There were **`r count(disturbed_nests_thv)` reports** of disturbed nests. 
Caveat: QA to exclude possible training records.

Coming soon: details of disturbance to nests.

### General disturbance
There were **`r count(dist_thv %>% filter_2017())` reports** of general disturbance or predator presence in 2017-18.
There were **`r count(dist_thv %>% filter_2018())` reports** of general disturbance or predator presence in 2018-19.

```{r map_dist_thv}
dist_thv %>% filter_2017() %>% group_by(disturbance_cause) %>% tally() %>% arrange(-n) %>% kable()
dist_thv %>% map_dist()

dist_thv %>% filter_2018() %>% group_by(disturbance_cause) %>% tally() %>% arrange(-n) %>% kable()
dist_thv %>% map_dist()
```

# Surveys
Curators are to QA 
[all `r placename_thv` surveys on WAStD](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__in=20,28,29),
following the 
[QA instructions for surveys](https://wastd.readthedocs.io/data_curators.html#data-flow-of-surveys).
Split up per site:

* [Thevenard Island North Beach](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=28)
* [Thevenard Island South Beach](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=29)
* [Thevenard Island Tagging Area](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=20)

Note, we might need to re-define the beaches into North and South, dissolving the tagging area.

Season 2018-19:

* Wed 14 Nov mornings: crossed off tracks, recorded a few for training but should be non-production run 
* Wed 14 Nov afternoon: corrine did some training at the cabin with volunteers. 
  Forms should have been deleted but may have been sent by mistake. Also non-production.
* Wed 14 Nov night: shouldn't be any records for that night
* Thu 15 Nov morning: normal morning walks with proper records

### Per season effort
Highest level of aggregation: entire season.
Assumption: all training surveys have been marked as such in WAStD.

#### Excluding training

```{r survey_season_stats_all}
surveys_thv %>% survey_season_stats %>% dt
```

### Per day effort
Lowest level of aggregation: daily.

```{r survey_count}
surveys_thv %>% survey_count_heatmap(placename_thv, prefix_thv)
surveys_thv %>% list_survey_count(placename_thv)
surveys_thv %>% survey_hours_heatmap(placename_thv, prefix_thv)
surveys_thv %>% list_survey_effort(placename_thv)
```

### Individual surveys
No aggregation: individual surveys.

```{r surveys_list}
surveys_thv %>% survey_show_detail %>% dt()
```

## Survey effort per primary reporter

### Caveat
This list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text.
To correctly report on team member effort, curators have to read the list of 
team members in the Site Visit "comments" field and attach the correct users 
to the Survey team list.
Once the surveys are comprehensively attributed to all volunteers present, this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with
team rosters.

```{r survey_effort_by_person}
personal_effort_thv <- surveys_thv %>% survey_hours_per_person()
personal_effort_thv %>% kable()
```

## Survey effort
### Season 2017-18
Number of surveys per sector times length of site.

At Thevenard North Beach, assuming the trip was 6 km,
`r survey_ground_covered(surveys, 28, 6, 2017)` km were walked 
in `r survey_count(surveys, 28, 2017)` surveys.

At Thevenard South Beach, assuming the trip was 6 km,
`r survey_ground_covered(surveys, 29, 6, 2017)` km were walked 
in `r survey_count(surveys, 29, 2017)` surveys.

At Thevenard Nesting Beach, assuming the trip was 1 km,
`r survey_ground_covered(surveys, 20, 1, 2017)` km were walked 
in `r survey_count(surveys, 20, 2017)` surveys.

### Season 2018-19
Number of surveys per sector times length of site.

At Thevenard North Beach, assuming the trip was 6 km,
`r survey_ground_covered(surveys, 28, 6, 2018)` km were walked 
in `r survey_count(surveys, 28, 2018)` surveys.

At Thevenard South Beach, assuming the trip was 6 km,
`r survey_ground_covered(surveys, 29, 6, 2018)` km were walked 
in `r survey_count(surveys, 29, 2018)` surveys.

At Thevenard Nesting Beach, assuming the trip was 1 km,
`r survey_ground_covered(surveys, 20, 1, 2018)` km were walked 
in `r survey_count(surveys, 20, 2018)` surveys.

# Data upload
## Raw data
* Raw data are exported into CSV spreadsheets.
* Key figures are exported into .png raster files.
* This report is rendered into a single HTML page.

```{r data_export_csv}
tracks_thv %>%
  select(-obs, -photos) %>%
  readr::write_csv(glue::glue("{prefix_thv}_tracks.csv"))
surveys_thv %>% readr::write_csv(glue::glue("{prefix_thv}_surveys.csv"))
dist_thv %>%
  select(-photos) %>%
  readr::write_csv(glue::glue("{prefix_thv}_disturbance.csv"))
```

## Data catalogue
Data are uploaded to the [turtle nest census dataset](https://data.dpaw.wa.gov.au/dataset/turtle-tracks) 
on the departmental data catalogue, accessible from the DBCA intranet only.

```{r data_upload}
# Package all output files into one ZIP archive
products_fn_thv <- glue::glue("{prefix_thv}_products.zip")
if (file.exists(products_fn_thv)) file.remove(products_fn_thv)
zip(zipfile = products_fn_thv, files = list.files(pattern = prefix_thv))

# Create a resource for the ZIP archives
# d <- ckanr::package_show("turtle-tracks")
# r <- resource_create(package_id=d$id, name="Thevenard outputs", upload=products_fn_thv)

# Update resources on data catalogue
ckanr::resource_update("19294b29-a70f-45e3-b06b-fc1eefe858d2", "thevenard.html")
ckanr::resource_update("25c887fa-0561-495a-944a-734b01050e8a", products_fn_thv)
```
