---
title: "Season review 2017-2018"
author: "West Pilbara Turtle Program Marine Turtles WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 9
    fig_height: 5
---

```{r r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(wastdr)
library(dplyr)
library(tidyr)
library(magrittr)
library(skimr)
library(leaflet)
library(lubridate)
library(listviewer)
library(DT)
library(ggplot2)
library(ckanr)

# Configure ckanr to data.dpaw.wa.gov.au
if (file.exists("~/.Rprofile")) source("~/.Rprofile")
ckanr::ckanr_setup(url = Sys.getenv("CKAN_URL"), key = Sys.getenv("CKAN_API_KEY"))
```

# Caveat
The data shown are pending QA and reconciliation with paper-based records. 
Until this warning disappears, this dataset may contain duplicates from training runs,
effort attributed to the wrong operators, and possibly other issues and biases.

# Data
## Load data
```{r get_data}
if (file.exists("../data/tracks.Rda")) {
  load("../data/tracks.Rda")
} else {
  q <- list(taxon = "Cheloniidae", format = "json")
  animal_records <- wastdr::wastd_GET("animal-encounters", query = q)
  animals <- wastdr::parse_animal_encounters(animal_records)
  track_records <- wastdr::wastd_GET("turtle-nest-encounters")
  tracks_all <- wastdr::parse_turtle_nest_encounters(track_records)
  dist_records <- wastdr::wastd_GET("disturbance-observations")
  disturbance <- wastdr::parse_disturbance_observations(dist_records)
  surveys <- wastd_GET("surveys") %>% wastdr::parse_surveys()
  save(animal_records, animals, tracks_all, track_records, 
       dist_records, disturbance, surveys, file = "../data/tracks.Rda"
  )
}
```


## Helpers
```{r helpers}
filter_2017 <- . %>% dplyr::filter(date > dmy("01/10/2017"))

filter_wp <- . %>% dplyr::filter(site_id %in% c(25, 26, 27))
filter_cw <- . %>% dplyr::filter(site_name == "Cooling Water Beach") # id 26
filter_bb <- . %>% dplyr::filter(site_name == "Bells Beach") # id 25
filter_yc <- . %>% dplyr::filter(site_name == "Wickam Yacht Club Beach") # id 27

daily_species_by_type <- . %>%
  filter(nest_age == "fresh") %>%
  group_by(date, species, nest_type) %>%
  tally() %>%
  ungroup()

daily_summary <- . %>%
  daily_species_by_type() %>%
  tidyr::spread(nest_type, n, fill = 0) %>%
  DT::datatable(.)

species_by_type <- . %>%
  filter(nest_age == "fresh") %>%
  group_by(species, nest_type) %>%
  tally() %>%
  ungroup() %>%
  tidyr::spread(nest_type, n, fill = 0)

tracks_ts <- . %>%
  daily_species_by_type() %>%
  {
    ggplot2::ggplot(data = ., aes(x = date, y = n, colour = nest_type)) +
      ggplot2::geom_point() +
      ggplot2::geom_smooth(method = "auto") +
      # ggplot2::geom_line() +
      ggplot2::scale_x_date(
        breaks = scales::pretty_breaks(),
        labels = scales::date_format("%d %b %Y")
      ) +
      ggplot2::scale_y_continuous(limits = c(0, NA)) +
      ggplot2::xlab("Date") +
      ggplot2::ylab("Number counted per day") +
      ggplot2::ggtitle("Nesting activity") +
      ggplot2::theme_light()
  }
```

## Filter data

```{r, eval=T}
tracks_wp <- tracks_all %>% filter_wp() %>% filter_2017()
# tracks_cw <- tracks_all %>% filter_cw %>% filter_2017
tracks_bb <- tracks_all %>% filter_bb() %>% filter_2017()
tracks_yc <- tracks_all %>% filter_yc() %>% filter_2017()

surveys_wp <- surveys %>% filter_wp() %>% filter_2017()
# surveys_cw <- surveys %>% filter_cw %>% filter_2017
surveys_bb <- surveys %>% filter_bb() %>% filter_2017()
surveys_yc <- surveys %>% filter_yc() %>% filter_2017()

dist_wp <- disturbance %>% filter_wp() %>% filter_2017()
```

# Season summary

## General notes
Nests without a clearly visible emergence point were not recorded.
This led to under-reporting of hatched nests this season.

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.
This season, the Volunteers have saved 
`r (nrow(tracks_wp) + nrow(surveys_wp)*2) * 2 / 60` hours of mind-numbing data entry
and proofreading work for `r nrow(tracks_wp)` Track or Treats and 
`r nrow(surveys_wp)*2` Site Visit Starts/Ends. 
This estimate excludes Fox Sakes and Marine Wildlife Incidents, about 2-3h extra!

## Maps
```{r map_overall}
tracks_wp %>% add_nest_labels() %>% map_tracks()
tracks_bb %>% add_nest_labels() %>% map_tracks()
tracks_yc %>% add_nest_labels() %>% map_tracks()
```

## How many nests did we have all together across all sectors?
```{r tracks_all_sectors}
tracks_wp %>% species_by_type() %>% kable()
```

```{r tracks_bb}
tracks_bb %>% species_by_type() %>% kable()
```

```{r tracks_yc}
tracks_yc %>% species_by_type() %>% kable()
```


```{r nesting_success}
tracks_wp %>% tracks_ts()
```

## How many nests were predated?
Coming soon

## What types of predation were recorded?
Coming soon

## Survey count

### Caveat
* Data includes training surveys.

This contributes to over-estimation of effort, but shows actual boots on the 
ground - training is also an effort!

```{r survey_count}
place <- "West Pilbara"
surveys_wp %>% plot_survey_count(place)
surveys_wp %>% list_survey_count(place)

place <- "West Pilbara Bells Beach"
surveys_bb %>% plot_survey_count(place)
surveys_bb %>% list_survey_count(place)

place <- "West Pilbara Yacht Club Beach"
surveys_yc %>% plot_survey_count(place)
surveys_yc %>% list_survey_count(place)
```

## Survey effort in hours
Q How many hours did volunteers survey, how many kilometers were walked?

### Caveat
* Surveys with missing end points are auto-closed after 5 hours. 
* Data includes training surveys.

Both factors contribute to over-estimation of effort.

```{r survey_effort}
surveys_wp %>% plot_survey_effort("West Pilbara")
surveys_wp %>% list_survey_effort("West Pilbara")

surveys_bb %>% plot_survey_effort("West Pilbara Bells Beach")
surveys_bb %>% list_survey_effort("West Pilbara Bells Beach")

surveys_yc %>% plot_survey_effort("West Pilbara Yacht Club Beach")
surveys_yc %>% list_survey_effort("West Pilbara Yacht Club Beach")
```

## Survey effort per primary reporter
Q Which volunteer did the most hours overall on survey?

### Caveat
This list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text.
To correctly report on team member effort, curators have to read the list of 
team members in the Site Visit "comments" field and attach the correct users 
to the Survey team list.
Once the surveys are comprehensively attributed to all volunteers present, this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with
team rosters.

```{r survey_effort_by_person}
personal_effort <- surveys_wp %>% survey_hours_per_person()
personal_effort %>% kable()
```

## Survey effort in kilometers walked
Number of surveys per sector times length of site.

```{r}
survey_count <- function(surveys, site_id) {
  surveys %>% filter(site_id == site_id) %>% nrow()
}
survey_ground_covered <- function(surveys, site_id, km_per_survey) {
  survey_count(surveys, site_id) * km_per_survey
}
```

At Bells Beach, assuming the round trip was 2 km,
`r survey_ground_covered(surveys, 25, 2)` km were walked 
in `r survey_count(surveys_wp, 25)` surveys.

At Yacht Club Beach, assuming the round trip was 1 km,
`r survey_ground_covered(surveys, 27, 1)` km were walked 
in `r survey_count(surveys_wp, 27)` surveys.

# Data upload

```{r data_upload}
# d <- ckanr::package_show("turtle-tracks")
# ckanr::resource_update("b35fb19e-e72d-420c-b875-12b4db4afe8c", "pthedland_2017-18.html")
```

```{r data_export}
tracks_wp %>% select(-obs, -photos) %>% readr::write_csv("../data/tracks_wp.csv")
surveys_wp %>% readr::write_csv("../data/surveys_wp.csv")
dist_wp %>% select(-photos) %>% readr::write_csv("../data/dist_wp.csv")
```

