---
title: "West Pilbara Season Review 2017-2018"
author: "West Pilbara Turtle Program with Marine Turtles WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
source("tracks_setup.R")
source("tracks_helpers.R")
source("load_data.R")
```

# Caveat
The data shown are currently updated by the curators through Quality Assurance (QA).
Update September 2018:

* Some records were uploaded but not ingested. This happened because
  the older forms used still accepted submissions, but were already
  retired from data ingest into WAStD. Re-enabling the ingest for those
  older forms managed to import the late uploads.
* Most tablets did have unfinalised, unsent forms. These required manual
  review (which is the purpose of leaving forms unfinalised), re-enabling
  of those old form versions both on the server and on the data ingest.
  The option to leave unfinished records on the tablets has now been removed.
* Missing surveys are automatically reconstructed.
* Coming soon: surveys missing an end point (default length 6h) will be adjusted
  to latest record plus buffer time. This will reduce over-estimation of survey effort.
  
The data shown here now includes these two batches of observations.
In situ tests in Sept 2018 have shown that 100% of the data collected digitally arrive in the database.

Remaining QA tasks:

* Reconcile reporter and team names from data as submitted (typos may lead to wrong match)
  and team roster.

# Data
```{r load_data, eval=T}
tracks_wp <- tracks_all %>% filter_wp() %>% exclude_training_species() %>% add_nest_labels()
# tracks_cw <- tracks_wp %>% filter_wp_coolingwater() # none yet
tracks_bb <- tracks_wp %>% filter_wp_bells()
tracks_yc <- tracks_wp %>% filter_wp_yachtblub()
# tracks_bt <- tracks_wp %>% filter_wp_boat() # none yet
# tracks_cl <- tracks_wp %>% filter_wp_cleaverville() # none yet

surveys_wp <- surveys %>% filter_wp()
surveys_wp_prod <- surveys_wp %>% exclude_training_surveys()

dist_wp <- disturbance %>% filter_wp()

# No nests yet
# nests_wp <- nests_all %>% filter_wp()
# nests_cw <- nests_all %>% filter_wp_coolingwater()
# nests_bb <- nests_all %>% filter_wp_bells()
# nests_yc <- nests_all %>% filter_wp_yachtblub()
# nests_bt <- nests_all %>% filter_wp_boat()
# nests_cl <- nests_all %>% filter_wp_cleaverville()

prefix <- "WPTP"
placename <- "Cape Lambert"
```

Note: Jason reported the following counts for manual entry (not done yet, so not showing up in this report):

```{r new_data, eval=F}
# 16/11:
# 5 false crawls, 3 nests:
# -20.61127   117.15553 below dune
# -20.61067 117.15619 above high water
# -20.61138 117.15530 right on the high water mark
#
# 22/11:
# 11 false crawls, 4 nests:
# -20.61144 117.15526 above high water
# -20.61031 117.15677 below dune
# -20.61119 117.15564 above high water
# -20.61129 117.15547 above high water
#
# 23/11:
# 2 false crawls, 2 nests:
# -20.61149 117.15523 Above high water
# -20.61143 117.15528 Above high water
```

# QA
Most QA is done across all WA records and programs in the separate QA workbook.


This section identifies data requiring editing.

## Username mismatches
TODO: define QA rules for mismatches of usernames as (possibly mis)typed by data collectors vs 
[WAStD users](https://tsc.dbca.wa.gov.au/admin/users/user/).

```{r name_mismatch}
surveys_wp %>% filter_surveys_requiring_qa() %>% dt()
```

# Season summary

## General notes
Nests without a clearly visible emergence point were not recorded.
This led to under-reporting of hatched nests this season.

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.
This season, the Volunteers have saved 
`r (nrow(tracks_wp) + nrow(surveys_wp)*2 + nrow(dist_wp)) * 2 / 60` 
hours of mind-numbing data entry and proofreading work for 
`r nrow(tracks_wp)` Tracks, 
`r nrow(dist_wp)` Disturbances, and 
`r nrow(surveys_wp)*2` Site Visit Starts/Ends. 

## Maps
```{r map_current_season}
tracks_wp %>% filter_2017() %>% map_tracks()
# tracks_wp %>% filter_2018() %>% map_tracks()
```

### Animated map
A [Google API key](https://developers.google.com/maps/documentation/geocoding/start?csw=1) must
be present as R environment variable `GOOGLE_MAPS_APIKEY` for the next step.

```{r map_animated, warning=FALSE}
gganimate_tracks(tracks_wp, placename, prefix)
# gganimate_tracks(tracks_cl, placename, prefix)
```


### Tagged nests
2017-18: no tagged nests.

The map (once enabled) is saved as a PNG file `r glue::glue(prefix, "_tagged_nests.png")`.\
This map can be printed and used to navigate back to the nests.

```{r nests_all, eval=F}
nests_wp %>%
  filter_2018() %>%
  map_nests() %>%
  mapview::mapshot(file = glue::glue("{prefix}_tagged_nests.png"))
```

## Nesting by season
### All beaches
The next table includes all tracks from all sites combined.

```{r tracks_all_sectors}
tracks_wp %>% nesting_type_by_season_species() %>% kable()
```

### Bells Beach
Only Bells Beach tracks.

```{r tracks_bb}
tracks_bb %>% nesting_type_by_season_species() %>% kable()
```

### Yacht Club Beach
Only Yacht Club Beach tracks.

```{r tracks_yc}
tracks_yc %>% nesting_type_by_season_species() %>% kable()
```

## Boat Beach
Coming 2018.

## Cleaverville Beach 1
Coming 2018.

## Nesting by week
### All beaches
The next table includes all tracks from Bells and Yacht Club beach combined.

```{r tracks_all_beaches_by_week}
tracks_wp %>% nesting_type_by_season_week_species() %>% kable()
```


## Nesting by day
This section shows data from all sites.

```{r nesting_success_all_beaches, warning = F}
tracks_wp %>% tracks_ts(placename = placename, prefix = prefix)
nests <- tracks_wp %>% track_success()
nests %>% ggplot_track_success_by_date("natator-depressus", placename, prefix)
# nests %>% ggplot_track_success_by_date("chelonia-mydas", placename)
nests %>% ggplot_track_successrate_by_date("natator-depressus", placename, prefix)
# nests %>% ggplot_track_successrate_by_date("chelonia-mydas", placename)
nests %>% track_success_by_species() %>% datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```

## Hatching and emergence success
Coming 2018?

```{r hatching_emergence_success}
tracks_wp %>%
  hatching_emergence_success() %>%
  DT::datatable(.,
    caption = "Hatching and emergence success summary",
    options = list(paging = F)
  )
```

## Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Fox Sake" and appear here as "dist_(place)".

### Disturbance and predation of nests
Turtle nests with recorded disturbance or predation. All beaches.

```{r dist_nests}
disturbed_nests_wp <- tracks_wp %>% filter_2017() %>% dplyr::filter(disturbance == "present")
disturbed_nests_wp %>% map_tracks()
```

There were **`r count(disturbed_nests_wp)` reports** of disturbed nests. 
Caveat: QA to exclude possible training records.

Coming soon: details of disturbance to nests.

### General disturbance
There were **`r count(dist_wp)` reports** of general disturbance or predator presence:

```{r dist_general}
dist_wp %>% disturbance_by_season() %>% dt()
dist_wp %>% filter_2017() %>% map_dist()
```


## Surveys
Curators are to QA 
[all West Pilbara surveys on WAStD](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__in=25,26,27,46,47) 
following the 
[QA instructions for surveys](https://wastd.readthedocs.io/data_curators.html#data-flow-of-surveys).

Split up per site:

* [Cooling Water Beach](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=26)
* [Bells Beach](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=25)
* [Yacht Club Beach](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=27)
* [Boat Beach](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=46)
* [Cleaverville Beach 1](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=47)

### Caveat 
* Survey data includes training surveys. (Not production, but also effort)
* Surveys with missing end points are auto-closed after 5 hours (over-estimation of effort).
* Data excludes commute to survey site.

### Per season effort
Highest level of aggregation: entire season.

* Excluding training, there were **`r nrow(surveys_wp_prod)` surveys** to all beaches
  of the WPTP (Bells, Yacht Club, coming 2018: Cleaverville, not included: cooling water)
  for a total duration of **`r sum(surveys_wp_prod$duration_hours)` hours**  over a period of **XXX days**.
* Caveat: all training surveys must be marked as "not production" manually in WAStD.
  Above data currently may include training effort.

### Per day effort
Lowest level of aggregation: daily.

```{r survey_count}
surveys_wp %>% survey_count_heatmap(placename, prefix)
surveys_wp %>% list_survey_count(placename)
surveys_wp %>% survey_hours_heatmap(placename, prefix)
surveys_wp %>% list_survey_effort(placename)
```

### Individual surveys
No aggregation: individual surveys.

Note, a direct link to update each survey in WAStD is at the end of the table.

```{r surveys_list}
surveys_wp %>%
  dplyr::select(
    -id, -site_type, -site_id, -reporter_username, -absolute_admin_url
  ) %>%
  dt()
```


## Volunteer effort
* Surveys including training runs represent volunteer effort.
* There were **`r surveys_wp %>% filter_2017() %>% nrow` surveys** 
  including training for a total duration
  of **`r sum(filter_2017(surveys_wp_prod)$duration_hours)` hours**.
* At Bells Beach, assuming the round trip was 2 km,
  **`r survey_ground_covered(surveys, 25, 2, season=2017)` km** were walked 
  in **`r survey_count(surveys, 25, season=2017)` surveys** including training.

* At Yacht Club Beach, assuming the round trip was 1 km,
  **`r survey_ground_covered(surveys, 27, 1, season=2017)` km** were walked 
  in **`r survey_count(surveys_wp, 27, season=2017)` surveys ** including training.

### Per person
The following list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text
(surveys > start_comments), and from 2018, in the Site Visit Start form field "team".

To correctly report on team member effort, curators have to read the list of 
team members in the Survey "start_comments" field and attach the correct users 
to the Survey's team list. From 2018, this will be automated but still requires QA.

Once the surveys are comprehensively attributed to all volunteers present, the code underneath this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with team rosters.

```{r survey_effort_by_person}
personal_effort <- surveys_wp %>% survey_hours_per_person()
personal_effort %>% kable()
```

# Data upload

## Raw data
* Raw data are exported into CSV spreadsheets.
* Key figures are exported into .png raster files.
* This report is rendered into a single HTML page.

```{r data_export_csv}
tracks_wp %>%
  dplyr::select(-obs, -photos) %>%
  readr::write_csv(glue::glue("{prefix}_tracks.csv"))
surveys_wp %>% readr::write_csv(glue::glue("{prefix}_surveys.csv"))
dist_wp %>%
  dplyr::select(-photos) %>%
  readr::write_csv(glue::glue("{prefix}_disturbance.csv"))
```

## Data catalogue
Data are uploaded to the [turtle nest census dataset](https://data.dpaw.wa.gov.au/dataset/turtle-tracks) 
on the departmental data catalogue, accessible from the DBCA intranet only.

```{r data_upload_ckan}
# Package all output files into one ZIP archive.
products_fn <- glue::glue("{prefix}_products.zip")
if (file.exists(products_fn)) file.remove(products_fn)
products <- list.files(pattern = prefix)
zip(zipfile = products_fn, files = products)

# Create a resource for the ZIP archive
# d <- ckanr::package_show("turtle-tracks")
# r <- resource_create(package_id=d$id, name="West Pilbara Turtle Program outputs", upload=products_fn)

# Update resources on data catalogue
ckanr::resource_update("1c57e04a-c29c-464c-b82d-a69e4ee0b199", "westpilbara.html")
ckanr::resource_update("53fcc7fc-a974-4f26-a4d7-2d39c9111211", products_fn)
```

## Google Drive
A copy of this report and all generated outputs is uploaded to Google Drive
and shared via link with DBCA external collaborators.

Every machine has to be authenticated with Google Drive once.
As this process involves browser windows and pasting of approval codes into the 
R Console, it is not run automatically when compiling this workbook.

Running the following chunk manually once per machine will cache the authentication
token locally for future use.

```{r google_drive_auth, eval=F}
googledrive::drive_auth(reset = TRUE, use_oob = TRUE)
```

```{r data_upload_google}
googledrive::drive_ls("WPTP") %>% googledrive::drive_rm(.)
googledrive::drive_upload("westpilbara.html", path = "WPTP/westpilbara.html")
products %>% purrr::map(googledrive::drive_upload, path = as_dribble("WPTP"))
```


