---
title: "West Pilbara Season Review 2017-2018"
author: "West Pilbara Turtle Program with Marine Turtles WA"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 9
    fig_height: 5
---

```{r r setup, include=FALSE, message=FALSE, warning=FALSE}
source("tracks_setup.R")
source("tracks_helpers.R")
```

# Caveat
The data shown are pending QA and reconciliation with paper-based records. 
Until this warning disappears, this dataset may contain duplicates from training runs, effort attributed to the wrong operators, and possibly other issues and biases.

Update September 2018:

* Some records were uploaded but not ingested. This happened because
  the older forms used still accepted submissions, but were already
  retired from data ingest into WAStD. Re-enabling the ingest for those
  older forms managed to import the late uploads.
* Most tablets did have unfinalised, unsent forms. These required manual
  review (which is the purpose of leaving forms unfinalised), re-enabling
  of those old form versions both on the server and on the data ingest.
  
The data shown here now includes these two batches of observations.
In situ tests in Sept 2018 have shown that 100% of the data collected digitally arrive in the database.

# Data
```{r, eval=T}
source("load_data.R")
tracks_wp <- tracks_all %>% filter_wp() %>% filter_2017()
# tracks_cw <- tracks_all %>% filter_cw %>% filter_2017() # none yet
tracks_bb <- tracks_all %>% filter_bb() %>% filter_2017()
tracks_yc <- tracks_all %>% filter_yc() %>% filter_2017()

surveys_wp <- surveys %>% filter_wp() %>% filter_2017()
# surveys_cw <- surveys %>% filter_cw %>% filter_2017()
surveys_bb <- surveys %>% filter_bb() %>% filter_2017()
surveys_yc <- surveys %>% filter_yc() %>% filter_2017()

dist_wp <- disturbance %>% filter_wp() %>% filter_2017()
```

Note: Jason reported the following counts for manual entry (not done yet, so not showing up in this report):

```{r, eval=F}
# 16/11:
# 5 false crawls, 3 nests:
# -20.61127   117.15553 below dune
# -20.61067 117.15619 above high water
# -20.61138 117.15530 right on the high water mark
#  
# 22/11:
# 11 false crawls, 4 nests:
# -20.61144 117.15526 above high water
# -20.61031 117.15677 below dune
# -20.61119 117.15564 above high water
# -20.61129 117.15547 above high water
#  
# 23/11:
# 2 false crawls, 2 nests:
# -20.61149 117.15523 Above high water
# -20.61143 117.15528 Above high water
```

# QA
This section identifies tracks and disturbances outside known sites, missing surveys, and surveys requiring editing.
We show data from all regions as not to miss any stray records - ignore data not relevant to this analysis.

## Tracks outside known sites
* If tracks are close to sites: Extend WAStD site polygons to include GPS error.
* Add WAStD sites for new surveyed sites (production or training) and include those sites in existing or new report.
* Change confirmed training records outside training areas to "Hatchback turtles" and mark as "curated" to
  prevent from being overwritten by repeated data loading.
* Note: re-run uncached data import after changing tracks.

```{r tracks_missing_sites}
tracks_all %>% 
  filter_nosite() %>%
  add_nest_labels() %>% 
  map_tracks()
```

```{r dist_missing_sites}
disturbance %>% filter_nosite() %>% map_dist()
```

## Tracks at known sites with missing surveys
Actions:

* Filter datatable to site of interest by typing the site name into the "Search" box.
* Create missing surveys, one for each row in the table. (This will be automated early 2019.)
* Re-run uncached data import after adding surveys, as data for surveys, 
  tracks and disturbance will change.
  
TODO: find earliest observation and reporter on given dates. (group by > summarize)

```{r missing_surveys}
tracks_all %>%
  filter_2017 %>%
  filter_nosurvey() %>%
  dplyr::group_by(site_name, date) %>%
  dplyr::summarise(earliest_record = min(datetime), 
                   no_tracks = n(), 
                   reporter = first(reporter)) %>%
  DT::datatable(.)
```
### QA: Surveys
Curators are to QA [all West Pilbara surveys on WAStD](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__in=25,26,27,46,47) following the 
[QA instructions for surveys](https://wastd.readthedocs.io/data_curators.html#data-flow-of-surveys).

Split up per site:

* [Cooling Water Beach](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=26)
* [Bells Beach](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=25)
* [Yacht Club Beach](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=27)
* [Boat Beach](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=46)
* [Cleaverville Beach 1](https://tsc.dbca.wa.gov.au/admin/observations/survey/?site__id__exact=47)

### QA: Surveys missing Site Visit End
Where a "Site visit end" was forgotten, "end source id" is empty, and the duration is set to 6h. 
These surveys should be closed a few minutes after the last recorded track or disturbance.

Surveys with missing end contribute to an over-estimation of survey effort.

TODO: list datetime of last record.
  
```{r}
surveys_wp %>% 
  dplyr::filter(is.na(end_source_id)) %>%
  dplyr::select(-site_type, -site_id, -reporter_username, -reporter_id, -id, -absolute_admin_url) %>% 
  DT::datatable(escape = FALSE, rownames = FALSE)
```

### QA: Provided username mismatches WAStD users
TODO: define QA rules for mismatches of usernames as (possibly mis)typed by data collectors vs 
[WAStD users](https://tsc.dbca.wa.gov.au/admin/users/user/).
```{r}
surveys_wp %>% 
  dplyr::filter(grepl("QA", start_comments) | grepl("QA", end_comments)) %>%
  dplyr::select(site_name, reporter, date, start_time, end_time, start_comments, end_comments, change_url) %>% 
  DT::datatable(escape = FALSE, rownames = FALSE)
```

# Season summary

## General notes
Nests without a clearly visible emergence point were not recorded.
This led to under-reporting of hatched nests this season.

## Going digital
Entering and proofreading a record takes about 2 minutes with full concentration.
By going digital, data entry and proofreading are fully automated.
This season, the Volunteers have saved 
`r (nrow(tracks_wp) + nrow(surveys_wp)*2 + nrow(dist_wp)) * 2 / 60` 
hours of mind-numbing data entry and proofreading work for 
`r nrow(tracks_wp)` Tracks, 
`r nrow(dist_wp)` Disturbances, and 
`r nrow(surveys_wp)*2` Site Visit Starts/Ends. 

## Maps
```{r map_overall}
tracks_wp %>% add_nest_labels() %>% map_tracks()
# tracks_bb %>% add_nest_labels() %>% map_tracks()
# tracks_yc %>% add_nest_labels() %>% map_tracks()
```

## Nesting abundance
### All sectors
```{r tracks_all_sectors}
tracks_wp %>% species_by_type() %>% kable()
```
## Bell's Beach
```{r tracks_bb}
tracks_bb %>% species_by_type() %>% kable()
```
### Yacht Club Beach
```{r tracks_yc}
tracks_yc %>% species_by_type() %>% kable()
```

## Nesting success 
```{r nesting_success, fig.height=5, fig.width=7, warning = F}
label <- "wp"
tracks_wp %>% tracks_ts(placename = label)
nests <- tracks_wp %>% track_success()
nests %>% ggplot_track_success_by_date("natator-depressus", label)
# nests %>% ggplot_track_success_by_date("chelonia-mydas", label)
nests %>% ggplot_track_successrate_by_date("natator-depressus", label)
# nests %>% ggplot_track_successrate_by_date("chelonia-mydas", label)
nests %>% track_success_by_species() %>% datatable(.,
  caption = "Nesting success of fresh tracks (excl. nests without tracks and tagging)"
)
```

## Hatching and emergence success
```{r hatching_emergence_success}
tracks_wp %>%
  hatching_emergence_success() %>%
  DT::datatable(.,
    caption = "Hatching and emergence success summary",
    options = list(paging = F)
  )
```

## Disturbance and predation
Disturbed nests are captured through form "Track or Treat" and appear here
as "tracks" with "disturbance" recorded as "present".

General signs of disturbance or predator presence are recorded through form 
"Fox Sake" and appear here as "dist_(place)".

### Disturbance and predation of nests
Turtle nests with recorded disturbance or predation.

```{r}
disturbed_nests_wp <- tracks_wp %>% filter(disturbance == "present")
disturbed_nests_wp %>% add_nest_labels() %>% map_tracks()
```

There were **`r count(disturbed_nests_wp)` reports** of disturbed nests. Caveat: QA to exclude possible training records.

Coming soon: details of disturbance to nests.

### General disturbance
There were **`r count(dist_wp)` reports** of general disturbance or predator presence:

```{r}
dist_wp %>% dplyr::group_by(disturbance_cause) %>% dplyr::tally() %>% dplyr::arrange(-n) %>% kable
dist_wp %>% map_dist
```


## Surveys
### Caveat 
* Survey data includes training surveys. (Not production, but also effort)
* Surveys with missing end points are auto-closed after 5 hours (over-estimation of effort).
* Data includes training surveys. This is desired for an estimate of volunteer contribution, but
  not desired for an estimate of "production"" survey effort.
* Data excludes commute to survey site.

### All Surveys
A direct link to update each survey is at the end of the table.

```{r}
surveys_wp %>% 
  dplyr::select(-site_type, -site_id, -reporter_username, -id, -absolute_admin_url) %>% 
  DT::datatable(escape = F, rownames = FALSE)
```

## Survey count
```{r survey_count}
surveys_wp %>% plot_survey_count("West Pilbara")
surveys_wp %>% list_survey_count("West Pilbara")

surveys_bb %>% plot_survey_count("West Pilbara Bells Beach")
surveys_bb %>% list_survey_count("West Pilbara Bells Beach")

surveys_yc %>% plot_survey_count("West Pilbara Yacht Club Beach")
surveys_yc %>% list_survey_count("West Pilbara Yacht Club Beach")
```

## Survey effort in hours
```{r survey_effort}
surveys_wp %>% plot_survey_effort("West Pilbara")
surveys_wp %>% list_survey_effort("West Pilbara")

surveys_bb %>% plot_survey_effort("West Pilbara Bells Beach")
surveys_bb %>% list_survey_effort("West Pilbara Bells Beach")

surveys_yc %>% plot_survey_effort("West Pilbara Yacht Club Beach")
surveys_yc %>% list_survey_effort("West Pilbara Yacht Club Beach")
```

## Survey effort per primary reporter

### Caveat
This list only includes the primary reporter. 
This leads to under-reporting of time spent by other volunteers attending the survey.
Their names are listed in the Site Visit Start form field "comments" as free text, and from 2018, in the Site Visit Start form field "team".
To correctly report on team member effort, curators have to read the list of 
team members in the Site Visit "comments" field and attach the correct users 
to the Survey team list. From 2018, this will be automated but still requires QA.
Once the surveys are comprehensively attributed to all volunteers present, the code underneath this
workbook has to be updated to include the Survey team's effort.

If the primary reporter forgot to update the username in the tablet,
their survey effort on that day is incorrectly attributed to the previous reporter.
The only way to reconcile attribution is to manually compare survey usernames with team rosters.

```{r survey_effort_by_person}
personal_effort <- surveys_wp %>% survey_hours_per_person()
personal_effort %>% kable()
```

## Survey effort in kilometers walked
Number of surveys per sector times length of site.
This includes local knowledge about distances on site.

At Bells Beach, assuming the round trip was 2 km,
`r survey_ground_covered(surveys, 25, 2)` km were walked 
in `r survey_count(surveys_wp, 25)` surveys.

At Yacht Club Beach, assuming the round trip was 1 km,
`r survey_ground_covered(surveys, 27, 1)` km were walked 
in `r survey_count(surveys_wp, 27)` surveys.

# Data upload

```{r data_upload}
# d <- ckanr::package_show("turtle-tracks")
ckanr::resource_update("1c57e04a-c29c-464c-b82d-a69e4ee0b199", "westpilbara_2017-18.html")
```

```{r data_export}
tracks_wp %>% dplyr::select(-obs, -photos) %>% 
  readr::write_csv(here::here("data", "tracks_wp.csv"))
surveys_wp %>% readr::write_csv(here::here("data", "surveys_wp.csv"))
dist_wp %>% dplyr::select(-photos) %>% 
  readr::write_csv(here::here("data", "dist_wp.csv"))
```

TODO: zip CSV and PDF/PNG exports, upload to a secure file share.

