---
title: "Aerial track census comparison"
author: "Marine Turtles WA"
date: "03/08/2017"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 7
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/projects/turtle-scripts/aerial")
if (file.exists("../config/setup.R")) source("../config/setup.R")
# if (file.exists("../config/load.R")) source("../config/load.R")
library(geojsonio)
library(magrittr)
library(jsonlite)
library(tidyverse)
library(purrr)
library(tibble)
library(leaflet)
library(DT)

map_chr_hack <- function(.x, .f, ...) {
    map(.x, .f, ...) %>%
        purrr::map_if(is.null, ~ NA_character_) %>%
        purrr::flatten_chr(.)
}

#' Convert a GeoJSON of aerial track counts to a tibble
gj_attrs <- function(geojson_list){
  geojson_list$features %>% {
    tibble::tibble(
      observer = purrr::map_chr(., c("properties", "observer")),
      longitude = purrr::map(., c("geometry", "coordinates")) %>% purrr::map_dbl(magrittr::extract2, 1),
      latitude = purrr::map(., c("geometry", "coordinates")) %>% purrr::map_dbl(magrittr::extract2, 2),
      species = purrr::map_chr(., c("properties", "species")),
      nest_age = purrr::map_chr(., c("properties", "nest_age")),
      wastd_id = map_chr_hack(., c("properties", "WAStD_ID"))
    )
  }
}

wastd_attrs <- function(geojson_list){
  geojson_list$features %>% {
    tibble::tibble(
      observer = "WAStD",
      longitude = purrr::map(., c("geometry", "coordinates")) %>% purrr::map_dbl(extract2, 1),
      latitude = purrr::map(., c("geometry", "coordinates")) %>% purrr::map_dbl(extract2, 2),
      species = purrr::map_chr(., c("properties", "species")),
      nest_age = purrr::map_chr(., c("properties", "nest_age")),
      wastd_id = map_chr_hack(., c("properties", "pk"))
    )
  }
}
```

# Data
Background on location, project, data acquisition.
Three operators (Sabrina, Tony, Scott) analyse aerial imagery.
One team on the ground at the time of the aerial survey counted the tracks.

```{r read_geojson}
wastd <- geojsonio::geojson_read("data/wastd_tracks.geojson") %>% wastd_attrs
sab <- geojsonio::geojson_read("data/2017-03-01_sabrinafo.geojson") %>% gj_attrs
tony <- geojsonio::geojson_read("data/2017-08-02_tonyt.geojson") %>% gj_attrs
scott <- geojsonio::geojson_read("data/2017-07-01_scottwh.geojson") %>% gj_attrs
```

# Map
```{r view_map, echo=FALSE}
#' makeAwesomeIcon factory
mkicon <- function(ico, col) leaflet::makeAwesomeIcon(icon = ico, markerColor = col)

trackIcons <- leaflet::awesomeIconList(
  "cheloniidae-fam" = mkicon('align-center', 'black'),
  "chelonia-mydas" = mkicon('align-center', 'green'),
  "eretmochelys-imbricata" = mkicon('align-center', 'blue'),
  "natator-depressus" = mkicon('align-center', 'red')
  )

leaflet(wastd) %>% 
  addProviderTiles(
    "Esri.WorldImagery", group = "Aerial") %>%
  addProviderTiles(
    "OpenStreetMap.Mapnik", group = "Place names") %>%
  setView(
    lng=115.0, lat=-21.45, zoom=12) %>%
  addAwesomeMarkers(
    ~longitude, ~latitude,
    data = wastd, 
    icon = ~trackIcons[species],
    label = ~paste(observer, wastd_id, nest_age, species),
    group = "WAStD") %>%
  addAwesomeMarkers(
    ~longitude, ~latitude,
    data = sab, 
    icon = ~trackIcons[species],
    label = ~paste(observer, wastd_id, nest_age, species),
    group = "Sabrina") %>%
  addAwesomeMarkers(
    ~longitude, ~latitude,
    data = scott, 
    icon = ~trackIcons[species],
    label = ~paste(observer, wastd_id, nest_age, species),
    group = "Scott") %>%
  addAwesomeMarkers(
    ~longitude, ~latitude,
    data = tony, 
    icon = ~trackIcons[species],
    label = ~paste(observer, wastd_id, nest_age, species),
    group = "Tony") %>%
  addLayersControl(
    baseGroups = c("Aerial", "Place names"),
    overlayGroups = c("WAStD", "Sabrina", "Scott", "Tony"))
```

# Comparison

## Tracks in WAStD (all) and seen by observers (unless missed)

* Rename columns in `wastd` to be distinct from user columns.
* Define and use helper function to create comparison table for agreements, 
  where the operator found tracks which also exist in WAStD.
* Define and use helper function to create comparison table for disagreements, 
  where the operator found tracks not in WAStD.
  
```{r comparison_table}
wastd_new <- wastd %>% 
  dplyr::transmute(
    wastd_id = wastd_id,
    w_lon = longitude,
    w_lat = latitude,
    w_species = species,
    w_nest_age = nest_age
  )

#' Join all records of wastd_tbl with matching records from user_tbl
compare_to_wastd <- function(wastd_tbl, user_tbl){
  wastd_tbl %>% 
    dplyr::left_join(
      dplyr::select(user_tbl, species, nest_age, wastd_id), 
      by = "wastd_id") %>%
    dplyr::mutate(
      agree_species = ifelse(w_species == species, 1, 0),
      agree_nest_age = ifelse(w_nest_age == nest_age, 1, 0)
    )
}

#' Join all records of user_tbl which are not in wastd_tbl
not_in_wastd <- function(wastd_tbl, user_tbl){
  user_tbl %>%
    dplyr::anti_join(wastd_tbl, by = "wastd_id")
}

falsely_fresh_nw <- . %>% filter(is.na(wastd_id)) %>% tally %>% as.integer

wastd_and_sab <- compare_to_wastd(wastd_new, sab)
sab_not_wastd <- not_in_wastd(wastd_new, sab)

agree_species_wastd_vs_sab <- wastd_and_sab %>% group_by(agree_species) %>% tally %>% ungroup
agree_nest_age_wastd_vs_sab <- wastd_and_sab %>% group_by(agree_nest_age) %>% tally %>% ungroup
correct_age <- filter(agree_nest_age_wastd_vs_sab, agree_nest_age==1)$n
falsely_old <- filter(agree_nest_age_wastd_vs_sab, agree_nest_age==0)$n
falsely_fresh <- filter(agree_nest_age_wastd_vs_sab, is.na(agree_nest_age))$n
agree_species <- filter(agree_species_wastd_vs_sab, agree_species==1)$n
disagree_species <- filter(agree_species_wastd_vs_sab, agree_species==0)$n
falsely_fresh_not_wastd <- sab_not_wastd %>% falsely_fresh_nw

summary_sab <- tibble(
  observer = "sabrina",
  compared_to = "wastd",
  correct_age,
  falsely_old = 0,
  falsely_fresh,
  falsely_fresh_not_wastd,
  agree_species,
  disagree_species
)

wastd_and_scott <- compare_to_wastd(wastd_new, scott)
scott_not_wastd <- not_in_wastd(wastd_new, scott)

agree_species_wastd_vs_scott <- wastd_and_scott %>% group_by(agree_species) %>% tally %>% ungroup
agree_nest_age_wastd_vs_scott <- wastd_and_scott %>% group_by(agree_nest_age) %>% tally %>% ungroup
correct_age <- filter(agree_nest_age_wastd_vs_scott, agree_nest_age==1)$n
falsely_old <- filter(agree_nest_age_wastd_vs_scott, agree_nest_age==0)$n
falsely_fresh <- filter(agree_nest_age_wastd_vs_scott, is.na(agree_nest_age))$n
agree_species <- filter(agree_species_wastd_vs_scott, agree_species==1)$n
disagree_species <- filter(agree_species_wastd_vs_scott, agree_species==0)$n
falsely_fresh_not_wastd <- scott_not_wastd %>% falsely_fresh_nw

summary_scott <- tibble(
  observer = "scott",
  compared_to = "wastd",
  correct_age,
  falsely_old = 0,
  falsely_fresh,
  falsely_fresh_not_wastd,
  agree_species,
  disagree_species
)


wastd_and_tony <- compare_to_wastd(wastd_new, tony)
tony_not_wastd <- not_in_wastd(wastd_new, tony)

agree_species_wastd_vs_tony <- wastd_and_tony %>% group_by(agree_species) %>% tally %>% ungroup
agree_nest_age_wastd_vs_tony <- wastd_and_tony %>% group_by(agree_nest_age) %>% tally %>% ungroup
correct_age <- filter(agree_nest_age_wastd_vs_tony, agree_nest_age==1)$n
falsely_old <- filter(agree_nest_age_wastd_vs_tony, agree_nest_age==0)$n
falsely_fresh <- filter(agree_nest_age_wastd_vs_tony, is.na(agree_nest_age))$n
agree_species <- filter(agree_species_wastd_vs_tony, agree_species==1)$n
disagree_species <- filter(agree_species_wastd_vs_tony, agree_species==0)$n
falsely_fresh_not_wastd <- tony_not_wastd %>% falsely_fresh_nw

summary_tony <- tibble(
  observer = "tony",
  compared_to = "wastd",
  correct_age,
  falsely_old,
  falsely_fresh,
  falsely_fresh_not_wastd,
  agree_species,
  disagree_species
)

smr <- rbind(summary_sab, 
             summary_scott, 
             summary_tony)
```


* NA: fresh tracks falsely counted as old (and therefore not recorded), 
  or missed altogether by aerial observer
* 1: agree ground and aerial
* 0: disagree ground and aerial: aerial observer recorded track falsely as old


### Sabrina vs WAStD

```{r dt_sab}
DT::datatable(wastd_and_sab, caption = "Tracks seen by Sabrina and WAStD")
```


### Scott vs WAStD

```{r dt_scott}
DT::datatable(wastd_and_scott, caption = "Tracks seen by Scott and WAStD")
```

### Tony vs WAStD

```{r dt_tony}
DT::datatable(wastd_and_tony, caption = "Tracks seen by Tony and WAStD")
```


## All tracks not in WAStD

* tracks not in wastd, where wastd id exists: track counted differently on ground (tagging) - not an error
* tracks not in wastd, where wastd id does not exist: old track falsely counted as fresh by aerial observer

### Sabrina not in WAStD

```{r sab_not_on_wastd}
DT::datatable(sab_not_wastd, caption = "Tracks seen by Sabrina not in WAStD")
```

### Scott not in WAStD

```{r scott_not_on_wastd}
DT::datatable(scott_not_wastd, caption = "Tracks seen by Scott not in WAStD")
```

### Tony not in WAStD

```{r tony_not_on_wastd}
DT::datatable(tony_not_wastd, caption = "Tracks seen by Tony not in WAStD")
```

### Map of tracks not in WAStD

```{r map_false_positives}
leaflet(tony_not_wastd) %>% 
  addProviderTiles(
    "Esri.WorldImagery", group = "Aerial") %>%
  addProviderTiles(
    "OpenStreetMap.Mapnik", group = "Place names") %>%
  setView(
    lng=115.0, lat=-21.45, zoom=12) %>%
  addAwesomeMarkers(
    ~longitude, ~latitude,
    data = wastd, 
    icon = ~trackIcons[species],
    label = ~paste(observer, wastd_id, nest_age, species),
    group = "WAStD") %>%
  addAwesomeMarkers(
    ~longitude, ~latitude,
    data = sab_not_wastd,
    icon = ~trackIcons[species],
    label = ~paste(observer, wastd_id, nest_age, species),
    group = "Sabrina only") %>%
  addAwesomeMarkers(
    ~longitude, ~latitude,
    data = scott_not_wastd,
    icon = ~trackIcons[species],
    label = ~paste(observer, wastd_id, nest_age, species),
    group = "Scott only") %>%
  addAwesomeMarkers(
    ~longitude, ~latitude,
    icon = ~trackIcons[species],
    label = ~paste(observer, wastd_id, nest_age, species),
    group = "Tony only") %>%
  addLayersControl(
    baseGroups = c("Aerial", "Place names"),
    overlayGroups = c("WAStD", "Sabrina only", "Scott only", "Tony only"))
```


# Discussion

## Summary

```{r}
DT::datatable(smr)
```

## Notes on data collection
* WAStD is the point of truth for track and nest age, but not for species.
* Scott only counted tracks he thought were fresh ones.
* Scott disregarded tracks he tought to be old.
* Scott analysed the tagging area as well, for which no tracks were counted on the 
  ground (and therefore are not in WAStD as tracks, but as taggings).
* Tony recorded some old tracks (against protocol), while Scott and Sabrina did
  not record tracks they detected but judged to be old.

* WAStD tracks omit tagging area, as presence of turtles leaving tracks there was
  already recorded as a "tagging" event.
* Tagged turtles leave tracks visible on aerial imagery.

To link recorded tracks, the WAStD id for each track was manually added to 
the observers' recorded data while visually comparing both the aerial imagery,
WAStD tracks and the observers' tracks.

`wastd_id` is set to 5555xx for tracks of tagged turtles visible on aerial imagery,
but counted as taggings in WAStD.

