---
title: "Turtle Nesting Census Previews 2019-20"
author: "Flatbacks WA, DBCA"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
  pdf_document:
    latex_engine: xelatex
  word_document: default
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

# All packages used in this example
library(ckanr)
library(dplyr)
library(DT)
library(fs)
library(googledrive)
library(knitr)
library(leaflet)
library(readr)
library(skimr)
library(ruODK)

tz <- "Australia/Perth"
loc <- fs::path("media")
fs::dir_create(loc)
```

# Download data

```{r svc_get}
pl <- ruODK::project_list()
pl

fl <- ruODK::form_list(pid = 1)
fl

# SV start
ruODK::ru_setup(pid=1, fid=fl$fid[[5]])
message(glue::glue("Downloading {ruODK::get_default_fid()}"))
ft <- ruODK::odata_service_get()
ft %>% knitr::kable(.)
svs <- ruODK::odata_submission_get(table = ft$url[[1]], verbose = T)

# SV end
ruODK::ru_setup(pid=1, fid=fl$fid[[4]])
message(glue::glue("Downloading {ruODK::get_default_fid()}"))
ft <- ruODK::odata_service_get()
ft %>% knitr::kable(.)
sve <- ruODK::odata_submission_get(table = ft$url[[1]], verbose = T)

# MWI
ruODK::ru_setup(pid=1, fid=fl$fid[[2]])
message(glue::glue("Downloading {ruODK::get_default_fid()}"))
ft <- ruODK::odata_service_get()
ft %>% knitr::kable(.)
mwi <- ruODK::odata_submission_get(table = ft$url[[1]], verbose = T)
mwi_dmg <- ft$url[[2]] %>% 
  ruODK::odata_submission_get(table = ., verbose = T) %>%
  dplyr::left_join(mwi, by = c("submissions_id" = "id"))
mwi_tag <- ft$url[[3]] %>% 
  ruODK::odata_submission_get(table = ., verbose = T) %>%
  dplyr::left_join(mwi, by = c("submissions_id" = "id"))

# Dist
message(glue::glue("Downloading {ruODK::get_default_fid()}"))
ruODK::ru_setup(pid=1, fid=fl$fid[[3]])
ft <- ruODK::odata_service_get()
ft %>% knitr::kable(.)
dist <- ruODK::odata_submission_get(table = ft$url[[1]], verbose = T)

# Tracks
ruODK::ru_setup(pid=1, fid=fl$fid[[8]])
message(glue::glue("Downloading {ruODK::get_default_fid()}"))
ft <- ruODK::odata_service_get()
ft %>% knitr::kable(.)
tracks <- ft$url[1] %>% 
  ruODK::odata_submission_get(table = ., verbose = T) %>% 
  dplyr::rename(
    # Adjust coordinate colnames
    longitude = x14,
    latitude = x15,
    altitude = x16
  )

tracks_dist <- ft$url[2] %>% 
  ruODK::odata_submission_get(table = ., verbose = T) %>% 
  dplyr::left_join(tracks, by = c("submissions_id" = "id"))

tracks_egg <- ft$url[3] %>%
  ruODK::odata_submission_get(table = ., verbose = T) #%>%
  # dplyr::left_join(tracks, by = c("submissions_id" = "id"))

tracks_log <- ft$url[4] %>% 
  ruODK::odata_submission_get(table = ., verbose = T) #%>%
  # dplyr::left_join(tracks, by = c("submissions_id" = "id"))

tracks_hatch <- ft$url[5] %>% 
  ruODK::odata_submission_get(table = ., verbose = T) #%>%
  # dplyr::left_join(tracks, by = c("submissions_id" = "id"))

data_fan_outlier <- ft$url[6] %>% 
  ruODK::odata_submission_get(table = ., verbose = T) #%>%
  # dplyr::left_join(tracks, by = c("submissions_id" = "id"))

data_light <- ft$url[7] %>% 
  ruODK::odata_submission_get(table = ., verbose = T) #%>%
  # dplyr::left_join(tracks, by = c("submissions_id" = "id"))
```

# Analyse data
## Tracks
### Nests
### Disturbed nests

## Pred and dist (non-nest)

## Strandings

## Surveys (start and end)

<!-- Summarise, analyse, visualise the tibble(s) data (, data_sub1, data_sub2) -->
```{r data_vis}
skimr::skim(data)
dplyr::glimpse(data)
DT::datatable(data)
```

```{r data_map, eval=F}
leaflet::leaflet(width = 800, height = 600) %>%
  leaflet::addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
  leaflet::addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  leaflet::clearBounds() %>%
  leaflet::addAwesomeMarkers(
    data = data,
    lng = ~longitude, lat = ~latitude,
    icon = leaflet::makeAwesomeIcon(text = "Q", markerColor = "red"),
    label = ~ glue::glue("{species} {nest_type}"),
    popup = ~ glue::glue(
      "<h3>{nest_age} {species} {nest_type}</h3>",
      "Seen on {observation_start_time}</br>",
      "Seen by {reporter}</br>",
      "Device {device_id}</br>",
      "<h5>Track</h5>",
      '<div><img src="{photo_track_1}"',
      ' height="150px" alt="Track photo 1"></img></div>',
      '<div><img src="{photo_track_2}"',
      ' height="150px" alt="Track photo 2"></img></div>'
    ),
    clusterOptions = leaflet::markerClusterOptions()
  ) %>%
  leaflet::addLayersControl(
    baseGroups = c("Place names", "Aerial"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```

```

# Analyse data
<!-- Summarise, analyse, visualise the tibble(s) data (, data_sub1, data_sub2) -->
```{r data_vis, eval=F}
skimr::skim(data)
dplyr::glimpse(data)
DT::datatable(head(data))
```

```{r data_map, eval=F}
leaflet::leaflet(width = 800, height = 600) %>%
  leaflet::addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
  leaflet::addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  leaflet::clearBounds() %>%
  leaflet::addAwesomeMarkers(
    data = data,
    lng = ~longitude, lat = ~latitude,
    icon = leaflet::makeAwesomeIcon(text = "Q", markerColor = "red"),
    # 
    # # With your own field names
    # 
    # label = ~ glue::glue("{area_name} {encounter_start_datetime}"),
    # popup = ~ glue::glue(
    #   "<h3>{area_name}</h3>",
    #   "Survey start {encounter_start_datetime}</br>",
    #   "Reporter {reporter}</br>",
    #   "Device {device_id}</br>",
    #   "<h5>Site</h5>",
    #   '<div><img src="{my_photo}"',
    #   ' height="150px" alt="My photo"></img></div>'
    # ),
    clusterOptions = leaflet::markerClusterOptions()
  ) %>%
  leaflet::addLayersControl(
    baseGroups = c("Place names", "Aerial"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```

# Export
<!--
The form submissions are now extracted and visualised. What's next:

* Save data to local files (e.g. CSV).
* Compile report (e.g. to HTML).
* Compress all outputs as ZIP.
* Upload these artifacts to a CKAN data catalogue.
* Upload same artifacts to Google Drive.

Notes: 

* Generate the HTML report once off without the next chunk (`eval=F`), 
as the chunk refers to the rendered output file (HTML) before the file is 
created initially.
* Run report always twice to generate (run 1) and upload (run 2) the latest HTML.
-->
```{r data_export, eval=FALSE}
#------------------------------------------------------------------------------#
# Prepare report and products as local files
#
rep_fn <- "my_report.html" # The file name you save this template under
data_fn <- here::here(loc, "data.csv") %>% as.character()           # Main data
data_sub1_fn <- here::here(loc, "data_sub1.csv") %>% as.character() # Nested table 1
data_sub2_fn <- here::here(loc, "data_sub2.csv") %>% as.character() # Nested table 2
zip_fn <- "products.zip" # Attachments as one zip file (top level)

# Write data tbls to CSV files
readr::write_csv(data, path = data_fn)
readr::write_csv(data_sub1, path = data_sub1_fn)
readr::write_csv(data_sub2, path = data_sub2_fn)

# Compress everything into `zip_fn`, retain relative path to `loc`
zip(zipfile = zip_fn, files = fs::dir_ls(loc))

#------------------------------------------------------------------------------#
# CKAN
#
# Upload to a CKAN data catalogue (need url and API key of a write permitted user)
# See ROpenSci package ckanr
ckanr::ckanr_setup(url = Sys.getenv("CKAN_URL"), key = Sys.getenv("CKAN_KEY"))
ckan_ds_name <- "my-ckan-dataset-slug"

# Run once to create resources on an existing dataset, then comment out
d <- ckanr::package_show(ckan_ds_name)
res_data_main <- ckanr::resource_create(
  package_id = d$id, name="Main data", upload = data_fn)
res_data_sub1 <- ckanr::resource_create(
  package_id = d$id, name="Nested data table 1", upload = data_sub1_fn)
res_data_sub2 <- ckanr::resource_create(
  package_id = d$id, name="Nested data table 2", upload = data_sub2_fn)
res_report <- ckanr::resource_create(
  package_id = d$id, name="Data report", upload = rep_fn)
res_zip <- ckanr::resource_create(
  package_id = d$id, name="All data and attachments", upload = zip_fn)

# Paste res_data_main$id over RID and keep here, repeat for each resource
r <- ckanr::resource_update(res_data_main$id, path = data_fn)
r <- ckanr::resource_update(res_data_sub1$id, path = data_sub1_fn)
r <- ckanr::resource_update(res_data_sub2$id, path = data_sub2_fn)
r <- ckanr::resource_update(res_report$id, path = res_report)
r <- ckanr::resource_update(res_zip$id, path = zip_fn)

#------------------------------------------------------------------------------#
# Google Drive
#
# Run once per machine, then comment out:
googledrive::drive_auth(reset=TRUE, use_oob = TRUE)

# Upload to Google Drive
gd_fn <- "My Google Drive folder name"
googledrive::drive_ls(gd_fn) %>% googledrive::drive_rm(.)  # Wipe older outputs
googledrive::drive_upload(rep_fn, path=rep_fn)             # Report as HTML
googledrive::drive_upload(data_fn, path=data_fn)           # Main data as CSV
googledrive::drive_upload(data_sub1_fn, path=data_sub1_fn) # Nested table 1 as CSV
googledrive::drive_upload(data_sub2_fn, path=data_sub2_fn) # Nested table 2 as CSV
googledrive::drive_upload(zip_fn, path=zip_fn)             # All outputs as ZIP
```
